### 2017-06-01 paper/paper.tex
Start

% vim: set spell spelllang=en tw=100 :

\documentclass{llncs}
\usepackage[pass]{geometry}

\usepackage{complexity}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cleveref}
\usepackage{tikz}

\usetikzlibrary{decorations, decorations.pathreplacing, calc, backgrounds}

\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}

\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
\newcommand{\citep}[1]{\cite{#1}}

% cref style
\crefname{figure}{Figure}{Figures}
\Crefname{figure}{Figure}{Figures}

% http://tex.stackexchange.com/questions/22100/the-bar-and-overline-commands
\newcommand{\shortoverline}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}

\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

\title{Observations from Parallelising Three Maximum Common Subgraph Algorithms}

\author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/K503058/1 and EP/M508056/1]} \and Patrick Prosser \and James Trimble\samethanks[1]}

\institute{University of Glasgow, Glasgow, Scotland,\\
\email{c.mccreesh.1@research.gla.ac.uk}}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

\bibliographystyle{splncs03}
\bibliography{dblp}

\end{document}

### 2017-06-01 paper/paper.tex
Lots of authors

@@
 
-\title{Observations from Parallelising Three Maximum Common Subgraph Algorithms}
+\title{Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms}
 
-\author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/K503058/1 and EP/M508056/1]} \and Patrick Prosser \and James Trimble\samethanks[1]}
+\author{Ruth Hoffmann \inst{1} \and Ciaran McCreesh\thanks{This work was supported by the
+Engineering and Physical Sciences Research Council [grant numbers EP/K503058/1, EP/M508056/1, and EP/M508056/1]}
+\inst{2} \and Patrick Prosser \inst{2} \and Samba Ndojh Ndiaye\thanks{This work
+was supported by the ANR project SoLStiCe (ANR-13-BS02-0002-01)} \inst{3}
+\and Craig Reilly \samethanks[1] \inst{2}
+\and Christine Solnon\samethanks[2]\inst{4} \and James Trimble\samethanks[1] \inst{2}}
 
-\institute{University of Glasgow, Glasgow, Scotland,\\
+\institute{
+    University of St Andrews, St Andrews, Scotland
+    \and University of Glasgow, Glasgow, Scotland
+    \and Universit\'e Lyon 1, LIRIS, UMR5205, F-69621, France
+    \and INSA-Lyon, LIRIS, UMR5205, F-69621, France \\
     \email{c.mccreesh.1@research.gla.ac.uk}}
@@
 
+\section{Sequential Algorithms}
+
+\subsection{A Basic Constraint Model}
+
+\subsection{Ensuring Connectedness}
+
+\subsection{The Splitting Algorithm}
+
+\subsection{The Clique Encoding}
+
+\subsection{The $k$-less Subgraph Isomorphism Algorithm}
+
+\section{Parallel Search}
+
+\section{Experiments}
+
+\section{Conclusion}
+
 \bibliographystyle{splncs03}

### 2017-06-01 paper/paper.tex
I know my ABCs

@@
 
-\author{Ruth Hoffmann \inst{1} \and Ciaran McCreesh\thanks{This work was supported by the
-Engineering and Physical Sciences Research Council [grant numbers EP/K503058/1, EP/M508056/1, and EP/M508056/1]}
-\inst{2} \and Patrick Prosser \inst{2} \and Samba Ndojh Ndiaye\thanks{This work
-was supported by the ANR project SoLStiCe (ANR-13-BS02-0002-01)} \inst{3}
+\author{
+    Ruth Hoffmann \inst{1}
+    \and Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/K503058/1, EP/M508056/1, and EP/M508056/1]} \inst{2}
+    \and Samba Ndojh Ndiaye\thanks{This work was supported by the ANR project SoLStiCe (ANR-13-BS02-0002-01)} \inst{3}
+    \and Patrick Prosser \inst{2}
     \and Craig Reilly \samethanks[1] \inst{2}

### 2017-06-01 paper/paper.tex
Just dump in thesis plots and text for now

@@
 \documentclass{llncs}
-\usepackage[pass]{geometry}
+\usepackage[pass,showframe]{geometry}
 
@@
 
+\begin{figure}[p]
+    \includegraphics*[scale=0.7]{plots/33v-par-cumulative.pdf}
+    \hfill
+    \includegraphics*[scale=0.7]{plots/plain-par-cumulative.pdf}
+
+    \vspace*{1em}
+    \includegraphics*[scale=0.7]{plots/33ved-par-cumulative.pdf}
+    \hfill
+    \includegraphics*[scale=0.7]{plots/plain-connected-par-cumulative.pdf}
+
+    \vspace*{1em}
+    \includegraphics*[scale=0.7]{plots/33ve-connected-par-cumulative.pdf}
+    \hfill
+    \includegraphics*[scale=0.7]{plots/sip-par-cumulative.pdf}
+
+    \caption{The cumulative number of instances solved in under a certain time, comparing sequential
+    (solid lines) and parallel (dotted lines) algorithms, for six different families of maximum
+    common (connected) subgraph problems. All experiments use 32 threads on a 16 core,
+    hyper-threaded system.}
+    \label{figure:mcs-par-unconnected-cumulative}
+\end{figure}
+
+Cumulative performance plots comparing each parallel algorithm to its sequential version are given
+in \cref{figure:mcs-par-unconnected-cumulative}, using 32 threads on 16 core hyper-threaded Intel Xeon
+E5-2640 v2 systems. The results are consistent: in each case the parallel algorithm is clearly and
+substantially better than its sequential version, except when the sequential runtime is below 100
+milliseconds.
+
+\begin{figure}[p]
+    \centering
+    \hspace*{0.5em}\includegraphics*[scale=0.7]{plots/plain-par-heatmap-clique.pdf}\hfill
+    \includegraphics*[scale=0.7]{plots/plain-par-heatmap-kdown.pdf}\hfill
+    \includegraphics*[scale=0.7]{plots/plain-par-heatmap-james.pdf}\hspace*{0.5em}
+
+    \vspace*{2em}
+    \centering
+    \hspace*{0.5em}\includegraphics*[scale=0.7]{plots/33ved-par-heatmap-clique.pdf}\hfill
+    \includegraphics*[scale=0.7]{plots/33ved-par-heatmap-james.pdf}\hspace*{0.5em}
+
+    \vspace*{2em}
+    \centering
+    \hspace*{0.5em}\includegraphics*[scale=0.7]{plots/sip-par-heatmap-kdown.pdf}\hfill
+    \includegraphics*[scale=0.7]{plots/sip-par-heatmap-james.pdf}\hspace*{0.5em}
+
+    \caption{The benefits of parallelism, for different sets of problem instances and for different
+    algorithms. All experiments use 32 threads on a 16 core, hyper-threaded system; points below the
+    top diagonal line indicate a speedup, and the lower diagonal line is a sixteen times speedup.}
+    \label{figure:mcs-par-speedups}
+\end{figure}
+
+\Cref{figure:mcs-par-speedups} gives instance-by-instance comparisons for a representative subset of
+these algorithm and problem combinations. For the clique algorithm on unlabelled graphs (top left),
+the results show fairly consistent speedups for harder instances: there is only a single instance
+whose sequential runtime is over two seconds where we obtain a speedup of less than ten. We also see
+a significant number of superlinear speedups. This is because there are relatively long proofs of
+optimality for most of the harder instances, but strong solutions are usually found quickly (and
+where they are not, the additional diversity helps). The results are similarly positive on
+labelled graphs (centre left), with speedups of greater than ten being the norm for harder
+instances, although there is more variety in the results.
+
+For $k{\downarrow}$ on the unlabelled maximum common subgraph instances (top centre), the typical speedup
+is between two and ten, and occasionally there are speedups of between ten and one hundred. A few
+instances show small slowdowns: remember that when running 32 threads on a 16 core, hyper-threaded
+system, each individual thread runs somewhat slower than it would if only a single thread were
+running---in other words, these slight slowdowns are due to hardware effects, not changes to the
+search tree.
+
+On the larger subgraph isomorphism instances (bottom left), we see a concentration of speedups
+of around ten. We also see strongly superlinear speedups in some cases, with eight
+instances finishing in under one second which timed out sequentially at one thousand seconds.
+Examining these instances closely shows that in each case, they are satisfiable with either $k = 0$
+or very low values of $k$, and that the superlinear speedups are due to the parallel search
+introducing early diversity, offsetting an incorrect initial value-ordering heuristic choice. We
+also see more instances with slight absolute slowdowns: these are instances where the additional
+threads contribute no helpful work to the solution, and where the slowdown due to not having 32
+times as many resources when using 32 threads is particularly pronounced (memory contention is much
+more of a problem for these larger graphs). Although not ideal, in aggregate these results are rare
+and are more than offset by the superlinear speedups for other instances.
+
+\begin{figure}[p]
+    \centering
+    \hspace*{1em}\includegraphics*[scale=0.7]{plots/plain-clique-par-repeat-heatmap.pdf}\hfill
+    \includegraphics*[scale=0.7]{plots/plain-clique-par-8to32-heatmap.pdf}\hspace*{1em}
+
+    \vspace*{1em}
+    \centering
+    \hspace*{1em}\includegraphics*[scale=0.7]{plots/plain-kdown-par-repeat-heatmap.pdf}\hfill
+    \includegraphics*[scale=0.7]{plots/plain-kdown-par-8to32-heatmap.pdf}\hspace*{1em}
+
+    \vspace*{1em}
+    \centering
+    \hspace*{1em}\includegraphics*[scale=0.7]{plots/plain-james-par-repeat-heatmap.pdf}\hfill
+    \includegraphics*[scale=0.7]{plots/plain-james-par-8to32-heatmap.pdf}\hspace*{1em}
+
+    \caption{Parallel search is reproducible and scalable, as well as risk-free and beneficial. In the
+    left-hand column, we compare each of the three parallel algorithms against a repeated run with
+    the same parameters, on the unconnected, unlabelled instances (points on the diagonal line show
+    reproducibility of runtimes). In the right-hand column, we compare going from 8 threads to 32
+    threads (points below the diagonal line are an improvement). All experiments are on a 16 core,
+    hyper-threaded system.}
+    \label{figure:mcs-par-repeats}
+\end{figure}
+
+What about the splitting algorithm? On unlabelled instances (top right), our speedups are typically
+between those obtained from the clique algorithm and those from $k{\downarrow}$, being mainly a
+little below ten times. However, in some cases we see absolute slowdowns of up to four times: in
+these cases, parallelism does little, and is not enough to offset the costs we paid when switching
+from an in-place data structure to a data structure which requires copying. Interestingly, we also
+see many more superlinear speedups for this algorithm, which suggests that its value ordering
+heuristics are particularly poor early-on in search. (This last point is not particularly
+surprising: compared to $k{\downarrow}$, the splitting algorithm does not try to reduce domains at
+all at the top of search. The clique encoding, meanwhile, is able to capture even richer information
+about variable-value relationships. The splitting algorithm is fast, but comparatively stupid.)
+
+Similar trends occur with the subgraph isomorphism instances (bottom right), and the absolute
+slowdowns go as high as a factor of ten---this is on instances with particularly large
+target graphs, where copying is most expensive, and where there is a high branching factor at the
+root of the search tree. Again, superlinear speedups are common.
+
+On labelled instances (centre right), the parallel performance is less erratic. Although we no
+longer see absolute slowdowns (excluding on very easy instances), we also see fewer superlinear
+speedups, with most speedups being between one and ten. In many of these instances, parallelism
+does very little: restricting splitting to a depth of five in these cases gives work balance
+problems, but does help us avoid the substantial overheads that come with deeper splitting. This is
+because the search trees for these instances tend to have a very low branching factor---in
+constraint programming terms, the labels make each domain contain only a few values.
+
+\Cref{figure:mcs-par-unconnected-cumulative,figure:mcs-par-speedups} show that for all three
+algorithms, parallel search is beneficial, and also (reasonably) risk-free: although in a few cases the
+constant factor slowdowns due to hardware limitations and having to make changes to data structures
+are unpleasant, we do not encounter any exponential slowdowns due to search tree changes. But what
+about reproducibility and scalability? Further experiments in \cref{figure:mcs-par-repeats}
+establish that both of these properties also hold. In the left-hand column, we plot each algorithm
+run against itself, using 32 threads in both cases. For the clique and $k{\downarrow}$ approaches,
+every instance is on the central diagonal line, showing an extremely high level of reproducibility.
+For the splitting algorithm, there is a little more variability, particularly for easier instances:
+inspecting the results suggests that this is not down to different amounts of search work being
+performed, but rather whether or not helper threads end up triggering additional copying.
+
+In the right-hand column of \cref{figure:mcs-par-repeats}, we plot the effects of moving from 8
+threads (on the same system) to 32 threads. Points below the diagonal line indicate an improvement.
+For the clique approach, increasing the number of threads is unequivocally beneficial. For the
+$k{\downarrow}$ algorithm, the benefits are slim, and sometimes performance worsens slightly due to
+memory contention (not work done) but we do not see any exponential slowdowns. Finally, for the
+splitting algorithm, the results are usually strictly better, although in a few cases the overheads
+increase further giving a slowdown; we also see an increase in the occurrence of strongly
+superlinear speedups.
+
+It is likely that the parallel results for the splitting algorithm can be improved further. In the
+labelled case, we are often not obtaining a good work balance. However, deeper splitting is
+expensive, particularly with larger graphs. One option to consider would be recomputing parts of the
+search tree to avoid the costs of potentially allowing a copy in the future. This would be extremely
+unpleasant from an implementation perspective, but may be necessary to obtain more even results.
+However, even without this change, the parallel algorithm is clearly preferable to the sequential
+one.
+
 \section{Conclusion}

### 2017-06-02 clique-code/clique.cc
Sequential code

@@ -1,0 +1,373 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "clique.hh"
+#include "bit_graph.hh"
+#include "template_voodoo.hh"
+
+#include <algorithm>
+#include <limits>
+#include <iostream>
+#include <atomic>
+#include <mutex>
+#include <numeric>
+#include <chrono>
+
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+using std::chrono::steady_clock;
+
+namespace
+{
+    using Association = std::map<std::pair<unsigned, unsigned>, unsigned>;
+    using AssociatedEdges = std::vector<std::pair<unsigned, unsigned> >;
+
+    auto modular_product(const VFGraph & g1, const VFGraph & g2) -> std::pair<Association, AssociatedEdges>
+    {
+        unsigned next_vertex = 0;
+        Association association;
+        AssociatedEdges edges;
+
+        for (unsigned l = 0 ; l < std::min(g1.vertices_by_label.size(), g2.vertices_by_label.size()) ; ++l)
+            for (unsigned v1 : g1.vertices_by_label[l])
+                for (unsigned v2 : g2.vertices_by_label[l])
+                    for (unsigned m = 0 ; m < std::min(g1.vertices_by_label.size(), g2.vertices_by_label.size()) ; ++m)
+                        for (unsigned w1 : g1.vertices_by_label[m])
+                            for (unsigned w2 : g2.vertices_by_label[m])
+                                if (v1 != w1 && v2 != w2
+                                        && (g1.edges[v1][v1] == g2.edges[v2][v2])
+                                        && (g1.edges[w1][w1] == g2.edges[w2][w2])
+                                        && (g1.edges[v1][w1] == g2.edges[v2][w2])
+                                        && (g1.edges[w1][v1] == g2.edges[w2][v2])) {
+
+                                    if (! association.count({v1, v2}))
+                                        association.insert({{v1, v2}, next_vertex++});
+                                    if (! association.count({w1, w2}))
+                                        association.insert({{w1, w2}, next_vertex++});
+
+                                    edges.push_back({association.find({v1, v2})->second, association.find({w1, w2})->second});
+                                }
+
+        return { association, edges };
+    }
+
+    auto unproduct(const Association & association, unsigned v) -> std::pair<unsigned, unsigned>
+    {
+        for (auto & a : association)
+            if (a.second == v)
[…]

### 2017-06-02 clique-code/solve_max_common_subgraph.cc

@@ -1,0 +1,358 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "clique.hh"
+
+#include <boost/program_options.hpp>
+
+#include <iostream>
+#include <fstream>
+#include <exception>
+#include <cstdlib>
+#include <chrono>
+#include <thread>
+#include <mutex>
+#include <condition_variable>
+#include <algorithm>
+
+namespace po = boost::program_options;
+
+using std::chrono::steady_clock;
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+
+namespace
+{
+    class GraphFileError :
+        public std::exception
+    {
+        private:
+            std::string _what;
+
+        public:
+            GraphFileError(const std::string & filename, const std::string & message) throw ();
+
+            auto what() const throw () -> const char *;
+    };
+
+    GraphFileError::GraphFileError(const std::string & filename, const std::string & message) throw () :
+        _what("Error reading graph file '" + filename + "': " + message)
+    {
+    }
+
+    auto GraphFileError::what() const throw () -> const char *
+    {
+        return _what.c_str();
+    }
+
+    auto read_word(std::ifstream & infile) -> unsigned
+    {
+        unsigned char a, b;
+        a = static_cast<unsigned char>(infile.get());
+        b = static_cast<unsigned char>(infile.get());
+        return unsigned(a) | (unsigned(b) << 8);
+    }
+
+    auto read_vf(const std::string & filename, bool unlabelled, bool no_edge_labels, bool undirected) -> VFGraph
+    {
+        VFGraph result;
+
+        std::ifstream infile{ filename };
+        if (! infile)
+            throw GraphFileError{ filename, "unable to open file" };
+
+        result.size = read_word(infile);
+        if (! infile)
+            throw GraphFileError{ filename, "error reading size" };
+
+        // to be like the CP 2011 labelling scheme...
+        int m = result.size * 33 / 100;
+        int p = 1, k1 = 0, k2 = 0;
+        while (p < m && k1 < 16) {
+            p *= 2;
+            k1 = k2;
+            k2++;
+        }
+
+        if (unlabelled)
+            result.vertices_by_label.resize(1);
+        else
+            result.vertices_by_label.resize(0x10000 >> (16 - k1));
+
+        result.vertex_labels.resize(result.size);
+        result.edges.resize(result.size);
+        for (auto & e : result.edges)
+            e.resize(result.size);
+
[…]

### 2017-06-02 parallel-clique-code/clique.cc
Add parallel clique code

@@ -1,0 +1,819 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "clique.hh"
+#include "bit_graph.hh"
+#include "template_voodoo.hh"
+
+#include <algorithm>
+#include <limits>
+#include <iostream>
+#include <atomic>
+#include <mutex>
+#include <condition_variable>
+#include <numeric>
+#include <chrono>
+#include <thread>
+
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+using std::chrono::steady_clock;
+
+namespace
+{
+    struct AtomicIncumbent
+    {
+        std::atomic<unsigned> value;
+
+        AtomicIncumbent()
+        {
+            value.store(0, std::memory_order_seq_cst);
+        }
+
+        bool update(unsigned v)
+        {
+            while (true) {
+                unsigned cur_v = value.load(std::memory_order_seq_cst);
+                if (v > cur_v) {
+                    if (value.compare_exchange_strong(cur_v, v, std::memory_order_seq_cst))
+                        return true;
+                }
+                else
+                    return false;
+            }
+        }
+
+        bool beaten_by(unsigned v) const
+        {
+            return v > value.load(std::memory_order_seq_cst);
+        }
+
+        unsigned get() const
+        {
+            return value.load(std::memory_order_relaxed);
+        }
+    };
+
+    template <typename Item_>
+    class Queue
+    {
+        private:
+            const bool _donations_possible;
+
+            std::atomic<bool> _want_producer;
+
+            /* These protect and watch all subsequent private members. */
+            std::mutex _mutex;
+            std::condition_variable _cond;
+
+            std::list<Item_> _list;
+            bool _initial_producer_done;
+            std::atomic<bool> _want_donations; /* Readable without holding _mutex */
+
+            /* We keep track of how many consumers are busy, to decide whether
+             * donations might be produced or whether consumers can exit. */
+            unsigned _number_busy;
+            std::atomic<unsigned> _number_idle;
+
+        public:
+            /**
+             * We need to know how many consumers we will have. We also allow
+             * donations to be disabled. */
+            Queue(unsigned number_of_dequeuers, bool donations_possible) :
+                _donations_possible(donations_possible),
+                _want_producer(true),
+                _initial_producer_done(false),
+                _number_busy(number_of_dequeuers),
+                _number_idle(0)
+            {
+                /* Initially we don't want donations */
[…]

### 2017-06-02 parallel-clique-code/solve_max_common_subgraph.cc

@@ -1,0 +1,364 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "clique.hh"
+
+#include <boost/program_options.hpp>
+
+#include <iostream>
+#include <fstream>
+#include <exception>
+#include <cstdlib>
+#include <chrono>
+#include <thread>
+#include <mutex>
+#include <condition_variable>
+#include <algorithm>
+
+namespace po = boost::program_options;
+
+using std::chrono::steady_clock;
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+
+namespace
+{
+    class GraphFileError :
+        public std::exception
+    {
+        private:
+            std::string _what;
+
+        public:
+            GraphFileError(const std::string & filename, const std::string & message) throw ();
+
+            auto what() const throw () -> const char *;
+    };
+
+    GraphFileError::GraphFileError(const std::string & filename, const std::string & message) throw () :
+        _what("Error reading graph file '" + filename + "': " + message)
+    {
+    }
+
+    auto GraphFileError::what() const throw () -> const char *
+    {
+        return _what.c_str();
+    }
+
+    auto read_word(std::ifstream & infile) -> unsigned
+    {
+        unsigned char a, b;
+        a = static_cast<unsigned char>(infile.get());
+        b = static_cast<unsigned char>(infile.get());
+        return unsigned(a) | (unsigned(b) << 8);
+    }
+
+    auto read_vf(const std::string & filename, bool unlabelled, bool no_edge_labels, bool undirected) -> VFGraph
+    {
+        VFGraph result;
+
+        std::ifstream infile{ filename };
+        if (! infile)
+            throw GraphFileError{ filename, "unable to open file" };
+
+        result.size = read_word(infile);
+        if (! infile)
+            throw GraphFileError{ filename, "error reading size" };
+
+        // to be like the CP 2011 labelling scheme...
+        int m = result.size * 33 / 100;
+        int p = 1, k1 = 0, k2 = 0;
+        while (p < m && k1 < 16) {
+            p *= 2;
+            k1 = k2;
+            k2++;
+        }
+
+        if (unlabelled)
+            result.vertices_by_label.resize(1);
+        else
+            result.vertices_by_label.resize(0x10000 >> (16 - k1));
+
+        result.vertex_labels.resize(result.size);
+        result.edges.resize(result.size);
+        for (auto & e : result.edges)
+            e.resize(result.size);
+
[…]

### 2017-06-09 cilk-kdown-code/parallel.cc
Cilk kdown

@@ -1,0 +1,903 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "parallel.hh"
+
+#include <algorithm>
+#include <array>
+#include <atomic>
+#include <chrono>
+#include <condition_variable>
+#include <functional>
+#include <limits>
+#include <list>
+#include <map>
+#include <numeric>
+#include <random>
+#include <thread>
+#include <tuple>
+#include <utility>
+#include <vector>
+
+#include <iostream>
+
+#include <boost/dynamic_bitset.hpp>
+
+#include <cilk/cilk.h>
+#include <cilk/reducer_opadd.h>
+
+using std::array;
+using std::atomic;
+using std::condition_variable;
+using std::fill;
+using std::find_if;
+using std::function;
+using std::get;
+using std::greater;
+using std::list;
+using std::make_pair;
+using std::map;
+using std::mutex;
+using std::mt19937;
+using std::numeric_limits;
+using std::pair;
+using std::thread;
+using std::to_string;
+using std::tuple;
+using std::uniform_int_distribution;
+using std::unique;
+using std::unique_lock;
+using std::vector;
+
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+using std::chrono::steady_clock;
+
+using std::cerr;
+using std::endl;
+
+using cilk::reducer;
+using cilk::op_add;
+
+namespace
+{
+    /// We'll use an array of unsigned long longs to represent our bits.
+    using BitWord = unsigned long long;
+
+    /// Number of bits per word.
+    static const constexpr int bits_per_word = sizeof(BitWord) * 8;
+
+    /**
+     * A bitset with a fixed maximum size. This only provides the operations
+     * we actually use in the bitset algorithms: it's more readable this way
+     * than doing all the bit voodoo inline.
+     *
+     * Indices start at 0.
+     */
+    template <unsigned words_>
+    class FixedBitSet
+    {
+        private:
+            using Bits = array<BitWord, words_>;
+
+            Bits _bits = {{ }};
+
+        public:
+            FixedBitSet() = default;
+
+            FixedBitSet(unsigned size)
+            {
+                assert(size <= words_ * bits_per_word);
+            }
+
+            FixedBitSet(const FixedBitSet &) = default;
+
+            FixedBitSet & operator= (const FixedBitSet &) = default;
+
+            /**
+             * Set a given bit 'on'.
+             */
+            auto set(int a) -> void
+            {
[…]

### 2017-06-09 cilk-kdown-code/solve_subgraph_isomorphism.cc

@@ -1,0 +1,285 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "lad.hh"
+#include "vf.hh"
+#include "dimacs.hh"
+#include "parallel.hh"
+
+#include <boost/program_options.hpp>
+
+#include <iostream>
+#include <exception>
+#include <cstdlib>
+#include <chrono>
+#include <thread>
+#include <mutex>
+#include <condition_variable>
+
+namespace po = boost::program_options;
+
+using std::chrono::steady_clock;
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+
+/* Helper: return a function that runs the specified algorithm, dealing
+ * with timing information and timeouts. */
+template <typename Result_, typename Params_, typename Data_>
+auto run_this_wrapped(const std::function<Result_ (const Data_ &, const Params_ &)> & func)
+    -> std::function<Result_ (const Data_ &, Params_ &, bool &, int)>
+{
+    return [func] (const Data_ & data, Params_ & params, bool & aborted, int timeout) -> Result_ {
+        /* For a timeout, we use a thread and a timed CV. We also wake the
+         * CV up if we're done, so the timeout thread can terminate. */
+        std::thread timeout_thread;
+        std::mutex timeout_mutex;
+        std::condition_variable timeout_cv;
+        std::atomic<bool> abort;
+        abort.store(false);
+        params.abort = &abort;
+        if (0 != timeout) {
+            timeout_thread = std::thread([&] {
+                    auto abort_time = std::chrono::steady_clock::now() + std::chrono::seconds(timeout);
+                    {
+                        /* Sleep until either we've reached the time limit,
+                         * or we've finished all the work. */
+                        std::unique_lock<std::mutex> guard(timeout_mutex);
+                        while (! abort.load()) {
+                            if (std::cv_status::timeout == timeout_cv.wait_until(guard, abort_time)) {
+                                /* We've woken up, and it's due to a timeout. */
+                                aborted = true;
+                                break;
+                            }
+                        }
+                    }
+                    abort.store(true);
+                    });
+        }
+
+        /* Start the clock */
+        params.start_time = std::chrono::steady_clock::now();
+
+        try {
+            auto result = func(data, params);
+
+            /* Clean up the timeout thread */
+            if (timeout_thread.joinable()) {
+                {
+                    std::unique_lock<std::mutex> guard(timeout_mutex);
+                    abort.store(true);
+                    timeout_cv.notify_all();
+                }
[…]

### 2017-06-09 paper/paper.tex
Text from James

@@
 \subsection{The Splitting Algorithm}
+
+Due to the special structure of the maximum common subgraph problem, the following property holds
+throughout the search process using the CP model: any two variables either have domains with no
+values in common (with the possible exception of \bot), or have identical domains. The algorithm
+McSplit exploits this property, while exploring essentially the same search tree as the CP model.
+Rather than storing a domain for each vertex in V(G) -- which results in redundant storage in cases
+where domains are identical -- equivalence classes of vertices in V(G) are stored, and a set of
+candidate vertices in H is stored for each equivalence class. This enables fast propagation of the
+constraints and smaller memory requirements. In addition, the data structure enables good branching
+heuristics to be calculated cheaply.
 

### 2017-06-09 paper/paper.tex
Muck around

@@
 McSplit exploits this property, while exploring essentially the same search tree as the CP model.
-Rather than storing a domain for each vertex in V(G) -- which results in redundant storage in cases
-where domains are identical -- equivalence classes of vertices in V(G) are stored, and a set of
-candidate vertices in H is stored for each equivalence class. This enables fast propagation of the
-constraints and smaller memory requirements. In addition, the data structure enables good branching
-heuristics to be calculated cheaply.
+Rather than storing a domain for each vertex in $\operatorname{V}(G)$, which results in redundant
+storage in cases where domains are identical, equivalence classes of vertices in
+$\operatorname{V}(G)$ are stored, and a set of candidate vertices in $H$ is maintained for each
+equivalence class. This enables fast propagation of the constraints and smaller memory requirements.
+In addition, the data structure enables good branching heuristics to be calculated cheaply.
 

### 2017-06-09 paper/paper.tex
Need more dollars

@@
 throughout the search process using the CP model: any two variables either have domains with no
-values in common (with the possible exception of \bot), or have identical domains. The algorithm
+values in common (with the possible exception of $\bot$), or have identical domains. The algorithm
 McSplit exploits this property, while exploring essentially the same search tree as the CP model.

### 2017-06-10 cilk-clique-code/clique.cc
Cilk clique

@@ -1,0 +1,426 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "clique.hh"
+#include "bit_graph.hh"
+#include "template_voodoo.hh"
+
+#include <algorithm>
+#include <numeric>
+#include <limits>
+#include <iostream>
+#include <atomic>
+#include <mutex>
+#include <chrono>
+#include <map>
+
+#include <cilk/cilk.h>
+#include <cilk/reducer_opadd.h>
+
+using std::chrono::steady_clock;
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+
+namespace
+{
+    using Association = std::map<std::pair<unsigned, unsigned>, unsigned>;
+    using AssociatedEdges = std::vector<std::pair<unsigned, unsigned> >;
+
+    auto modular_product(const VFGraph & g1, const VFGraph & g2) -> std::pair<Association, AssociatedEdges>
+    {
+        unsigned next_vertex = 0;
+        Association association;
+        AssociatedEdges edges;
+
+        for (unsigned l = 0 ; l < std::min(g1.vertices_by_label.size(), g2.vertices_by_label.size()) ; ++l)
+            for (unsigned v1 : g1.vertices_by_label[l])
+                for (unsigned v2 : g2.vertices_by_label[l])
+                    for (unsigned m = 0 ; m < std::min(g1.vertices_by_label.size(), g2.vertices_by_label.size()) ; ++m)
+                        for (unsigned w1 : g1.vertices_by_label[m])
+                            for (unsigned w2 : g2.vertices_by_label[m])
+                                if (v1 != w1 && v2 != w2
+                                        && (g1.edges[v1][v1] == g2.edges[v2][v2])
+                                        && (g1.edges[w1][w1] == g2.edges[w2][w2])
+                                        && (g1.edges[v1][w1] == g2.edges[v2][w2])
+                                        && (g1.edges[w1][v1] == g2.edges[w2][v2])) {
+
+                                    if (! association.count({v1, v2}))
+                                        association.insert({{v1, v2}, next_vertex++});
+                                    if (! association.count({w1, w2}))
+                                        association.insert({{w1, w2}, next_vertex++});
+
+                                    edges.push_back({association.find({v1, v2})->second, association.find({w1, w2})->second});
+                                }
+
+        return { association, edges };
+    }
+
[…]

### 2017-06-10 cilk-clique-code/heatmapify.rb

@@ -1,0 +1,33 @@
+#!/usr/bin/ruby
+# vim: set sw=4 sts=4 et tw=80 :
+
+buckets = Hash.new(0)
+granularity = 101
+maxbx = 0
+
+IO.readlines(ARGV[0]).each_with_index do | line, index |
+    if index == 0 then next end
+    words = line.split(' ')
+
+    bx = words[0].to_i
+
+    y = words[1].sub('NA', '0')
+    by = ((1.0 + y.to_f) * (granularity / 2.0)).round
+
+    buckets[[bx, by]] += 1
+    maxbx = [bx, maxbx].max
+end
+
+ysums = [0] * (maxbx + 1)
+0.upto maxbx do | x |
+    0.upto granularity do | y |
+        ysums[x] += (buckets[[x, y]] || 0)
+    end
+end
+
+0.upto granularity do | y |
+    0.upto maxbx do | x |
+    print((buckets[[x, y]] != 0 ? buckets[[x, y]] / ysums[x].to_f : 0).to_s + " ")
+    end
+    puts
+end

### 2017-06-10 cilk-clique-code/solve_max_common_subgraph.cc

@@ -1,0 +1,327 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "clique.hh"
+
+#include <boost/program_options.hpp>
+
+#include <iostream>
+#include <fstream>
+#include <exception>
+#include <cstdlib>
+#include <chrono>
+#include <thread>
+#include <mutex>
+#include <condition_variable>
+#include <algorithm>
+
+namespace po = boost::program_options;
+
+using std::chrono::steady_clock;
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+
+namespace
+{
+    class GraphFileError :
+        public std::exception
+    {
+        private:
+            std::string _what;
+
+        public:
+            GraphFileError(const std::string & filename, const std::string & message) throw ();
+
+            auto what() const throw () -> const char *;
+    };
+
+    GraphFileError::GraphFileError(const std::string & filename, const std::string & message) throw () :
+        _what("Error reading graph file '" + filename + "': " + message)
+    {
+    }
+
+    auto GraphFileError::what() const throw () -> const char *
+    {
+        return _what.c_str();
+    }
+
+    auto read_word(std::ifstream & infile) -> unsigned
+    {
+        unsigned char a, b;
+        a = static_cast<unsigned char>(infile.get());
+        b = static_cast<unsigned char>(infile.get());
+        return unsigned(a) | (unsigned(b) << 8);
+    }
+
+    auto read_vf(const std::string & filename, bool unlabelled, bool no_edge_labels, bool undirected) -> VFGraph
+    {
+        VFGraph result;
+
+        std::ifstream infile{ filename };
+        if (! infile)
+            throw GraphFileError{ filename, "unable to open file" };
+
+        result.size = read_word(infile);
+        if (! infile)
+            throw GraphFileError{ filename, "error reading size" };
+
+        // to be like the CP 2011 labelling scheme...
+        int m = result.size * 33 / 100;
+        int p = 1, k1 = 0, k2 = 0;
+        while (p < m && k1 < 16) {
+            p *= 2;
+            k1 = k2;
+            k2++;
+        }
+
+        if (unlabelled)
+            result.vertices_by_label.resize(1);
+        else
+            result.vertices_by_label.resize(0x10000 >> (16 - k1));
+
+        result.vertex_labels.resize(result.size);
+        result.edges.resize(result.size);
+        for (auto & e : result.edges)
+            e.resize(result.size);
+
[…]

### 2017-06-11 paper/paper.tex
Plotting

@@
 
-\section{Parallel Search}
+\section{Experiments with Parallel Search}
 
-\section{Experiments}
+?? Hardware.
+
+?? Datasets.
+
+\subsection{Parallel Search is Better Overall}
+
+\begin{figure}[p]
+    \includegraphics*{gen-graph-cumulative-plain.pdf}
+    \hfill
+    \includegraphics*{gen-graph-cumulative-33v.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-cumulative-33ved.pdf}
+    \hfill
+
+\end{figure}
+
+\begin{figure}[p]
+    \includegraphics*{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-clique-vs-clique-cilk-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-plain-kdown-vs-kdown-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-kdown-vs-kdown-cilk-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-plain-mcsplit-vs-mcsplit-par-t32.pdf}
+
+\end{figure}
+
+\begin{figure}[p]
+    \includegraphics*{gen-graph-scatter-plain-clique-par-t32r-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-clique-cilk-t32r-vs-clique-cilk-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-plain-kdown-par-t32r-vs-kdown-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-kdown-cilk-t32r-vs-kdown-cilk-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t32r-vs-mcsplit-par-t32.pdf}
+
+\end{figure}
+
+\begin{figure}[p]
+    \includegraphics*{gen-graph-scatter-33v-clique-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33v-clique-vs-clique-cilk-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-33v-mcsplit-vs-mcsplit-par-t32.pdf}
+\end{figure}
+
+\begin{figure}[p]
+    \includegraphics*{gen-graph-scatter-33v-clique-par-t32r-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33v-clique-cilk-t32r-vs-clique-cilk-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-33v-mcsplit-par-t32r-vs-mcsplit-par-t32.pdf}
+
+\end{figure}
+
+\begin{figure}[p]
+    \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-cilk-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-33ved-mcsplit-vs-mcsplit-par-t32.pdf}
+\end{figure}
+
+\begin{figure}[p]
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t32r-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t32r-vs-clique-cilk-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-33ved-mcsplit-par-t32r-vs-mcsplit-par-t32.pdf}
+
+\end{figure}
 

### 2017-06-11 paper/paper.tex
added a few lines about kdown.

--- 2017-06-09 paper/paper.tex
@@
 \subsection{The $k$-less Subgraph Isomorphism Algorithm}
+
+The algorithm as described by Hoffmann, McCreesh, and Reilly \cite{DBLP:conf/aaai/HoffmannMR17} approaches the subgraph isomorphism problem by asking the question ``if a pattern graph cannot be found in the target, how much of the pattern graph can be found?''.
+The $k{\downarrow}$ algorithm tries to solve the subgraph isomorphism problem first for $k=0$ (the whole pattern graph can be found in the target), should that be not satisfiable, it tries to solve the problem for $k=1$ (one vertex cannot be matched) should that also not be satisfiable, it iterates $k$ further by a small number to keep the shrinking pattern graph within a reasonable size.
+By unmatching some vertices  invariants are weakened but kept effective.
+Specifically, invariants based around paths and degrees of vertices hold.
+
+This technique bridges the subgraph isomorphism problem with the maximal common subgraph problem.
+The iteration of the decision problem (iterating over $k$) leads to an algorithm with rivalling runtimes.
 

### 2017-06-12 paper/paper.tex
Connected plots
More plotting
Plot all the things

--- 2017-06-11 paper/paper.tex
@@
 
-\section{Parallel Search}
+\section{Experiments with Parallel Search}
 
-\section{Experiments}
+?? Hardware.
+
+?? Datasets.
+
+\subsection{Parallel Search is Better Overall}
 
 \begin{figure}[p]
+    \includegraphics*{gen-graph-cumulative-plain.pdf}
+    \hfill
+    \includegraphics*{gen-graph-cumulative-33v.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-cumulative-33ved.pdf}
+    \hfill
+    \includegraphics*{gen-graph-cumulative-plain-connected.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-cumulative-33ve-connected.pdf}
+    \hfill
+    \includegraphics*{gen-graph-cumulative-sip.pdf}
+
+    \caption{Cumulative performance.}
+
+\end{figure}
+
+\begin{figure}[p]
+    \includegraphics*{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-kdown-vs-kdown-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-mcsplit-vs-mcsplit-par-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-histogram-plain-clique-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-histogram-plain-kdown-vs-kdown-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-histogram-plain-mcsplit-vs-mcsplit-par-t32.pdf}
+
+    \caption{Per-instance speedups for unlabelled, undirected graphs.}
+\end{figure}
+
+\begin{figure}[t]
+    \includegraphics*{gen-graph-scatter-33v-clique-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33v-mcsplit-vs-mcsplit-par-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-histogram-33v-clique-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-histogram-33v-mcsplit-vs-mcsplit-par-t32.pdf}
+
+    \caption{Per-instance speedups for vertex-labelled graphs.}
+\end{figure}
+
+\begin{figure}[t]
+    \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-mcsplit-vs-mcsplit-par-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-histogram-33ved-clique-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-histogram-33ved-mcsplit-vs-mcsplit-par-t32.pdf}
+
+    \caption{Per-instance speedups for vertex-and edge-labelled, directed graphs.}
+\end{figure}
+
+\begin{figure}[t]
     \includegraphics*[scale=0.7]{plots/33v-par-cumulative.pdf}

### 2017-06-12 paper/paper.tex
Nuke old plots

@@
 
-\begin{figure}[t]
-    \includegraphics*[scale=0.7]{plots/33v-par-cumulative.pdf}
-    \hfill
-    \includegraphics*[scale=0.7]{plots/plain-par-cumulative.pdf}
-
-    \vspace*{1em}
-    \includegraphics*[scale=0.7]{plots/33ved-par-cumulative.pdf}
-    \hfill
-    \includegraphics*[scale=0.7]{plots/plain-connected-par-cumulative.pdf}
-
-    \vspace*{1em}
-    \includegraphics*[scale=0.7]{plots/33ve-connected-par-cumulative.pdf}
-    \hfill
-    \includegraphics*[scale=0.7]{plots/sip-par-cumulative.pdf}
-
-    \caption{The cumulative number of instances solved in under a certain time, comparing sequential
-    (solid lines) and parallel (dotted lines) algorithms, for six different families of maximum
-    common (connected) subgraph problems. All experiments use 32 threads on a 16 core,
-    hyper-threaded system.}
-    \label{figure:mcs-par-unconnected-cumulative}
-\end{figure}
-
 Cumulative performance plots comparing each parallel algorithm to its sequential version are given
-in \cref{figure:mcs-par-unconnected-cumulative}, using 32 threads on 16 core hyper-threaded Intel Xeon
+in ??, using 32 threads on 16 core hyper-threaded Intel Xeon
 E5-2640 v2 systems. The results are consistent: in each case the parallel algorithm is clearly and
@@
 
-\begin{figure}[p]
-    \centering
-    \hspace*{0.5em}\includegraphics*[scale=0.7]{plots/plain-par-heatmap-clique.pdf}\hfill
-    \includegraphics*[scale=0.7]{plots/plain-par-heatmap-kdown.pdf}\hfill
-    \includegraphics*[scale=0.7]{plots/plain-par-heatmap-james.pdf}\hspace*{0.5em}
-
-    \vspace*{2em}
-    \centering
-    \hspace*{0.5em}\includegraphics*[scale=0.7]{plots/33ved-par-heatmap-clique.pdf}\hfill
-    \includegraphics*[scale=0.7]{plots/33ved-par-heatmap-james.pdf}\hspace*{0.5em}
-
-    \vspace*{2em}
-    \centering
-    \hspace*{0.5em}\includegraphics*[scale=0.7]{plots/sip-par-heatmap-kdown.pdf}\hfill
-    \includegraphics*[scale=0.7]{plots/sip-par-heatmap-james.pdf}\hspace*{0.5em}
-
-    \caption{The benefits of parallelism, for different sets of problem instances and for different
-    algorithms. All experiments use 32 threads on a 16 core, hyper-threaded system; points below the
-    top diagonal line indicate a speedup, and the lower diagonal line is a sixteen times speedup.}
-    \label{figure:mcs-par-speedups}
-\end{figure}
-
-\Cref{figure:mcs-par-speedups} gives instance-by-instance comparisons for a representative subset of
+?? gives instance-by-instance comparisons for a representative subset of
 these algorithm and problem combinations. For the clique algorithm on unlabelled graphs (top left),
@@
 
-\begin{figure}[p]
-    \centering
-    \hspace*{1em}\includegraphics*[scale=0.7]{plots/plain-clique-par-repeat-heatmap.pdf}\hfill
-    \includegraphics*[scale=0.7]{plots/plain-clique-par-8to32-heatmap.pdf}\hspace*{1em}
-
-    \vspace*{1em}
-    \centering
-    \hspace*{1em}\includegraphics*[scale=0.7]{plots/plain-kdown-par-repeat-heatmap.pdf}\hfill
-    \includegraphics*[scale=0.7]{plots/plain-kdown-par-8to32-heatmap.pdf}\hspace*{1em}
-
-    \vspace*{1em}
-    \centering
-    \hspace*{1em}\includegraphics*[scale=0.7]{plots/plain-james-par-repeat-heatmap.pdf}\hfill
-    \includegraphics*[scale=0.7]{plots/plain-james-par-8to32-heatmap.pdf}\hspace*{1em}
-
-    \caption{Parallel search is reproducible and scalable, as well as risk-free and beneficial. In the
-    left-hand column, we compare each of the three parallel algorithms against a repeated run with
-    the same parameters, on the unconnected, unlabelled instances (points on the diagonal line show
-    reproducibility of runtimes). In the right-hand column, we compare going from 8 threads to 32
-    threads (points below the diagonal line are an improvement). All experiments are on a 16 core,
-    hyper-threaded system.}
-    \label{figure:mcs-par-repeats}
-\end{figure}
-
 What about the splitting algorithm? On unlabelled instances (top right), our speedups are typically
@@
 
-\Cref{figure:mcs-par-unconnected-cumulative,figure:mcs-par-speedups} show that for all three
+?? show that for all three
 algorithms, parallel search is beneficial, and also (reasonably) risk-free: although in a few cases the
@@
 are unpleasant, we do not encounter any exponential slowdowns due to search tree changes. But what
-about reproducibility and scalability? Further experiments in \cref{figure:mcs-par-repeats}
+about reproducibility and scalability? Further experiments in ??
 establish that both of these properties also hold. In the left-hand column, we plot each algorithm
@@
 
-In the right-hand column of \cref{figure:mcs-par-repeats}, we plot the effects of moving from 8
+In the right-hand column of ??, we plot the effects of moving from 8
 threads (on the same system) to 32 threads. Points below the diagonal line indicate an improvement.

### 2017-06-12 clique-code/clique.cc
Prime

@@ -111,6 +111,7 @@ namespace
             vertices_adjacent_to_by_g1(g1.size)
         {
             incumbent.start_time = params.start_time;
+            incumbent.value = params.prime;
 
             // populate our order with every vertex initially
             std::iota(order.begin(), order.end(), 0);

### 2017-06-12 clique-code/solve_max_common_subgraph.cc

@@ -206,6 +206,7 @@ auto main(int argc, char * argv[]) -> int
             ("no-edge-labels",                        "Get rid of edge labels, but keep vertex labels")
             ("undirected",                            "Make the graph undirected")
             ("connected",                             "Only find connected subgraphs")
+            ("prime",              po::value<int>(),  "Set initial incumbent size")
             ;
 
         po::options_description all_options{ "All options" };
@@ -247,6 +248,9 @@ auto main(int argc, char * argv[]) -> int
         Params params;
 
         params.connected = options_vars.count("connected");
+
+        if (options_vars.count("prime"))
+            params.prime = options_vars["prime"].as<int>();
 
         if (params.connected && ! options_vars.count("undirected")) {
             std::cerr << "Currently --connected requires --undirected" << std::endl;

### 2017-06-12 james-cpp-parallel/mcsp.cc

@@ -54,6 +54,7 @@ static struct argp_option options[] = {
     {"vertex-labelled-only", 'x', 0, 0, "Use vertex labels, but not edge labels"},
     {"big-first", 'b', 0, 0, "First try to find an induced subgraph isomorphism, then decrement the target size"},
     {"timeout", 't', "timeout", 0, "Specify a timeout (seconds)"},
+    {"prime", 'p', "prime", 0, "Specify initial incumbent size"},
     {"threads", 'T', "threads", 0, "Specify how many threads to use"},
     { 0 }
 };
@@ -72,6 +73,7 @@ static struct {
     char *filename1;
     char *filename2;
     int timeout;
+    int prime;
     int threads;
     int arg_num;
 } arguments;
@@ -91,6 +93,7 @@ void set_default_arguments() {
     arguments.filename1 = NULL;
     arguments.filename2 = NULL;
     arguments.timeout = 0;
+    arguments.prime = 0;
     arguments.threads = std::thread::hardware_concurrency();
     arguments.arg_num = 0;
 }
@@ -139,6 +142,9 @@ static error_t parse_opt (int key, char *arg, struct argp_state *state) {
             break;
         case 't':
             arguments.timeout = std::stoi(arg);
+            break;
+        case 'p':
+            arguments.prime = std::stoi(arg);
             break;
         case 'T':
             arguments.threads = std::stoi(arg);
@@ -792,6 +798,7 @@ std::pair<vector<VtxPair>, unsigned long long> mcs(const Graph & g0, const Graph
     }
 
     AtomicIncumbent global_incumbent;
+    global_incumbent.value = arguments.prime;
     vector<VtxPair> incumbent;
     unsigned long long global_nodes = 0;
 

### 2017-06-12 james-cpp/mcsp.c

@@ -47,6 +47,7 @@ static struct argp_option options[] = {
     {"vertex-labelled-only", 'x', 0, 0, "Use vertex labels, but not edge labels"},
     {"big-first", 'b', 0, 0, "First try to find an induced subgraph isomorphism, then decrement the target size"},
     {"timeout", 't', "timeout", 0, "Specify a timeout (seconds)"},
+    {"prime", 'p', "prime", 0, "Specify initial incumbent size"},
     { 0 }
 };
 
@@ -64,6 +65,7 @@ static struct {
     char *filename1;
     char *filename2;
     int timeout;
+    int prime;
     int arg_num;
 } arguments;
 
@@ -82,6 +84,7 @@ void set_default_arguments() {
     arguments.filename1 = NULL;
     arguments.filename2 = NULL;
     arguments.timeout = 0;
+    arguments.prime = 0;
     arguments.arg_num = 0;
 }
 
@@ -129,6 +132,9 @@ static error_t parse_opt (int key, char *arg, struct argp_state *state) {
             break;
         case 't':
             arguments.timeout = std::stoi(arg);
+            break;
+        case 'p':
+            arguments.prime = std::stoi(arg);
             break;
         case ARGP_KEY_ARG:
             if (arguments.arg_num == 0) {
@@ -456,7 +462,7 @@ vector<VtxPair> mcs(const Graph & g0, const Graph & g1) {
         domains.push_back({start_l, start_r, left_len, right_len, false});
     }
 
-    vector<VtxPair> incumbent;
+    vector<VtxPair> incumbent(arguments.prime, {-1, -1});
 
     if (arguments.big_first) {
         for (int k=0; k<g0.n; k++) {

### 2017-06-12 parallel-clique-code/clique.cc

@@ -394,6 +394,8 @@ namespace
             nodes(0),
             vertices_adjacent_to_by_g1(g1.size)
         {
+            best_anywhere.value = params.prime;
+
             // populate our order with every vertex initially
             std::iota(order.begin(), order.end(), 0);
 

### 2017-06-12 parallel-clique-code/solve_max_common_subgraph.cc

@@ -207,6 +207,7 @@ auto main(int argc, char * argv[]) -> int
             ("undirected",                            "Make the graph undirected")
             ("connected",                             "Only find connected subgraphs")
             ("threads",            po::value<int>(),  "Number of threads to use")
+            ("prime",              po::value<int>(),  "Set initial incumbent size")
             ;
 
         po::options_description all_options{ "All options" };
@@ -248,6 +249,9 @@ auto main(int argc, char * argv[]) -> int
         Params params;
 
         params.connected = options_vars.count("connected");
+
+        if (options_vars.count("prime"))
+            params.prime = options_vars["prime"].as<int>();
 
         if (params.connected && ! options_vars.count("undirected")) {
             std::cerr << "Currently --connected requires --undirected" << std::endl;

### 2017-06-12 paper/paper.tex
Plottery
Plotting

@@
 
+\begin{figure}[t]
+    \includegraphics*{gen-graph-scatter-sip-kdown-vs-kdown-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-sip-mcsplit-vs-mcsplit-par-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-histogram-sip-kdown-vs-kdown-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-histogram-sip-mcsplit-vs-mcsplit-par-t32.pdf}
+
+    \caption{Per-instance speedups for vertex-and edge-labelled, directed graphs.}
+\end{figure}
+
+\begin{figure}[p]
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t32r-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t2.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t2-vs-clique-par-t4.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t4-vs-clique-par-t8.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t8-vs-clique-par-t16.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t16-vs-clique-par-t32.pdf}
+
+    \caption{Scalability and reproducibility, clique algorithm}
+\end{figure}
+
+\begin{figure}[p]
+    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t32r-vs-mcsplit-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-mcsplit-vs-mcsplit-par-t2.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t2-vs-mcsplit-par-t4.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t4-vs-mcsplit-par-t8.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t8-vs-mcsplit-par-t16.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t16-vs-mcsplit-par-t32.pdf}
+
+    \caption{Scalability and reproducibility, McSplit algorithm}
+\end{figure}
+
 Cumulative performance plots comparing each parallel algorithm to its sequential version are given

### 2017-06-14 paper/paper.tex
Consistency

@@
 \documentclass{llncs}
-\usepackage[pass,showframe]{geometry}
+\usepackage[pass]{geometry}
 
@@
 
-    \caption{Cumulative performance.}
+    \caption{Cumulative performance. The parallel versions, dotted, are always better in aggregate
+    than the sequential versions, solid.}
 
@@
 
-    \caption{Per-instance speedups for unlabelled, undirected graphs.}
+    \caption{Per-instance speedups for unlabelled, undirected graphs. The histograms show the
+    distribution of speedups for instances whose sequential runtime is at least $50$ milliseconds
+    and no more than $100$ seconds (and so they under-represent superlinear speedups slightly).}
 \end{figure}
@@
 
-    \caption{Scalability and reproducibility, clique algorithm}
+    \caption{Scalability and reproducibility, clique algorithm.}
 \end{figure}
@@
 
-    \caption{Scalability and reproducibility, McSplit algorithm}
+    \caption{Scalability and reproducibility, McSplit algorithm.}
 \end{figure}

### 2017-06-14 paper/paper.tex
Abstract

@@
 
-\documentclass{llncs}
-\usepackage[pass]{geometry}
+\documentclass[sigconf]{acmart}
 
-\usepackage{complexity}
+\usepackage{showframe}
+
 \usepackage{microtype}
@@
 
-\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
-\newcommand{\citep}[1]{\cite{#1}}
-
 % cref style
@@
 
-\title{Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms}
+% Copyright
+%\setcopyright{none}
+%\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+\setcopyright{rightsretained}
+%\setcopyright{usgov}
+%\setcopyright{usgovmixed}
+%\setcopyright{cagov}
+%\setcopyright{cagovmixed}
 
-\author{
-    Ruth Hoffmann \inst{1}
-    \and Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/K503058/1, EP/M508056/1, and EP/M508056/1]} \inst{2}
-    \and Samba Ndojh Ndiaye\thanks{This work was supported by the ANR project SoLStiCe (ANR-13-BS02-0002-01)} \inst{3}
-    \and Patrick Prosser \inst{2}
-    \and Craig Reilly \samethanks[1] \inst{2}
-    \and Christine Solnon\samethanks[2]\inst{4} \and James Trimble\samethanks[1] \inst{2}}
+% DOI
+\acmDOI{10.475/123_4}
 
-\institute{
-    University of St Andrews, St Andrews, Scotland
-    \and University of Glasgow, Glasgow, Scotland
-    \and Universit\'e Lyon 1, LIRIS, UMR5205, F-69621, France
-    \and INSA-Lyon, LIRIS, UMR5205, F-69621, France \\
-    \email{c.mccreesh.1@research.gla.ac.uk}}
+% ISBN
+\acmISBN{123-4567-24-567/08/06}
+
+%Conference
+\acmConference[$\textnormal{IA}^3$ 2017]{Seventh Workshop on Irregular Applications: Architectures and Algorithms}{November 2017}{Denver, Colorado USA}
+\acmYear{2017}
+\copyrightyear{2017}
+
+\acmPrice{666.00}
 
 \begin{document}
+\title[Observations from Parallelising Three Subgraph Algorithms]{Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms}
+\titlenote{This work was supported by the Engineering and Physical Sciences
+    Research Council [grant numbers EP/K503058/1, EP/M508056, and EP/P026842/1] and the ANR project
+SoLStiCe (ANR-13-BS02-0002-01)}
+
+\author{Ruth Hoffmann}
+\affiliation{%
+    \institution{University of St Andrews}
+    \city{St Andrews, Scotland}
+}
+
+\author{Ciaran McCreesh}
+\affiliation{%
+    \institution{University of Glasgow}
+    \city{Glasgow, Scotland}
+}
+\email{ciaran.mccreesh@glasgow.ac.uk}
+
+\author{Samba Ndojh Ndiaye}
+\affiliation{
+    \institution{Universit\'e Lyon 1, LIRIS, UMR5205}
+    \postcode{F-69621, France}
+}
+
+\author{Patrick Prosser}
+\affiliation{%
+    \institution{University of Glasgow}
+    \city{Glasgow, Scotland}
+}
+
+\author{Craig Reilly}
+\affiliation{%
+    \institution{University of Glasgow}
+    \city{Glasgow, Scotland}
+}
+
+\author{Christine Solnon}
+\affiliation{
+    \institution{INSA-Lyon, LIRIS, UMR5205}
+    \city{F-69621, France}
+}
+
+\author{James Trimble}
+\affiliation{%
+    \institution{University of Glasgow}
+    \city{Glasgow, Scotland}
+}
+
+% The default list of authors is too long for headers}
+\renewcommand{\shortauthors}{R. Hoffmann et al.}
+
+\begin{abstract}
+    We discuss our experiences adapting three recent algorithms for maximum common (connected) subgraph
+    problems to exploit multi-core parallelism. We consider the effects of different work
+    distribution strategies and their interaction with value-ordering heuristics, and propose a
+    search method which is risk-free, reproducible, and scalable, as well as beneficial. Our results
+    show that each algorithm can be parallelised successfully using this strategy, with the
+    multi-core algorithms we create being clearly better than the sequential ones. We then look in
+    more detail at the results, on an instance by instance basis, and discuss how we should be
+    measuring speedups for these kinds of algorithm.
+\end{abstract}
+
+%
+% The code below should be generated by the tool at
+% http://dl.acm.org/ccs.cfm
+% Please copy and paste the code instead of the example below. 
+%
+\begin{CCSXML}
+\end{CCSXML}
+
+% \ccsdesc[500]{Computer systems organization~Embedded systems}
+% \ccsdesc[300]{Computer systems organization~Redundancy}
+% \ccsdesc{Computer systems organization~Robotics}
+% \ccsdesc[100]{Networks~Network reliability}
+
+% \keywords{ACM proceedings, \LaTeX, text tagging}
 
 \maketitle
-
-\begin{abstract}
-
-\end{abstract}
 
@@
 
-\begin{figure}[p]
+\begin{figure*}[p]
     \includegraphics*{gen-graph-cumulative-plain.pdf}
@@
 
-\end{figure}
+\end{figure*}
 
-\begin{figure}[p]
+\begin{figure*}[p]
     \includegraphics*{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}
@@
     and no more than $100$ seconds (and so they under-represent superlinear speedups slightly).}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[t]
+\begin{figure*}[t]
     \includegraphics*{gen-graph-scatter-33v-clique-vs-clique-par-t32.pdf}
@@
     \caption{Per-instance speedups for vertex-labelled graphs.}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[t]
+\begin{figure*}[t]
     \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t32.pdf}
@@
     \caption{Per-instance speedups for vertex-and edge-labelled, directed graphs.}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[t]
+\begin{figure*}[t]
     \includegraphics*{gen-graph-scatter-sip-kdown-vs-kdown-par-t32.pdf}
@@
     \caption{Per-instance speedups for vertex-and edge-labelled, directed graphs.}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[p]
+\begin{figure*}[p]
     \includegraphics*{gen-graph-scatter-33ved-clique-par-t32r-vs-clique-par-t32.pdf}
@@
     \caption{Scalability and reproducibility, clique algorithm.}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[p]
+\begin{figure*}[p]
     \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t32r-vs-mcsplit-par-t32.pdf}
@@
     \caption{Scalability and reproducibility, McSplit algorithm.}
-\end{figure}
+\end{figure*}
 
@@
 
-\bibliographystyle{splncs03}
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{dblp}

### 2017-07-31 paper/paper.tex
Convert from PASEO to IA3

@@
 \begin{abstract}
-    We discuss our experiences adapting three recent algorithms for maximum common (connected) subgraph
-    problems to exploit multi-core parallelism. We consider the effects of different work
-    distribution strategies and their interaction with value-ordering heuristics, and propose a
-    search method which is risk-free, reproducible, and scalable, as well as beneficial. Our results
-    show that each algorithm can be parallelised successfully using this strategy, with the
-    multi-core algorithms we create being clearly better than the sequential ones. We then look in
-    more detail at the results, on an instance by instance basis, and discuss how we should be
-    measuring speedups for these kinds of algorithm.
 \end{abstract}

### 2017-07-31 paper/paper.tex
Fixy fixy

--- 2017-06-14 paper/paper.tex
@@
 \begin{figure*}[p]
+    \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t32r-vs-clique-cilk-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-cilk-t2.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t2-vs-clique-cilk-t4.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t4-vs-clique-cilk-t8.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t8-vs-clique-cilk-t16.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t16-vs-clique-cilk-t32.pdf}
+
+    \caption{Scalability and reproducibility, clique algorithm, Cilk.}
+\end{figure*}
+
+\begin{figure*}[p]
     \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t32r-vs-mcsplit-par-t32.pdf}

### 2017-08-01 paper/paper.tex
Start resizing things

@@
 
-\begin{figure*}[p]
+\begin{figure*}[tb]
     \includegraphics*{gen-graph-cumulative-plain.pdf}
@@
     \includegraphics*{gen-graph-cumulative-33v.pdf}
+    \hfill
+    \includegraphics*{gen-graph-cumulative-33ved.pdf}
 
@@
 
-    \includegraphics*{gen-graph-cumulative-33ved.pdf}
+    \includegraphics*{gen-graph-cumulative-plain-connected.pdf}
     \hfill
-    \includegraphics*{gen-graph-cumulative-plain-connected.pdf}
-
-    \vspace*{1em}
-
     \includegraphics*{gen-graph-cumulative-33ve-connected.pdf}
@@
     than the sequential versions, solid.}
-
 \end{figure*}
 
-\begin{figure*}[p]
+\begin{figure*}[tb]
     \includegraphics*{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}
@@
     \includegraphics*{gen-graph-scatter-plain-mcsplit-vs-mcsplit-par-t32.pdf}
-
-    \vspace*{1em}
-
-    \includegraphics*{gen-graph-histogram-plain-clique-vs-clique-par-t32.pdf}
     \hfill
-    \includegraphics*{gen-graph-histogram-plain-kdown-vs-kdown-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-histogram-plain-mcsplit-vs-mcsplit-par-t32.pdf}
-
-    \caption{Per-instance speedups for unlabelled, undirected graphs. The histograms show the
-    distribution of speedups for instances whose sequential runtime is at least $50$ milliseconds
-    and no more than $100$ seconds (and so they under-represent superlinear speedups slightly).}
-\end{figure*}
-
-\begin{figure*}[t]
     \includegraphics*{gen-graph-scatter-33v-clique-vs-clique-par-t32.pdf}
@@
 
-    \includegraphics*{gen-graph-histogram-33v-clique-vs-clique-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-histogram-33v-mcsplit-vs-mcsplit-par-t32.pdf}
-
-    \caption{Per-instance speedups for vertex-labelled graphs.}
-\end{figure*}
-
-\begin{figure*}[t]
     \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t32.pdf}
@@
     \includegraphics*{gen-graph-scatter-33ved-mcsplit-vs-mcsplit-par-t32.pdf}
-
-    \vspace*{1em}
-
-    \includegraphics*{gen-graph-histogram-33ved-clique-vs-clique-par-t32.pdf}
     \hfill
-    \includegraphics*{gen-graph-histogram-33ved-mcsplit-vs-mcsplit-par-t32.pdf}
-
-    \caption{Per-instance speedups for vertex-and edge-labelled, directed graphs.}
-\end{figure*}
-
-\begin{figure*}[t]
     \includegraphics*{gen-graph-scatter-sip-kdown-vs-kdown-par-t32.pdf}
@@
 
-    \vspace*{1em}
-
-    \includegraphics*{gen-graph-histogram-sip-kdown-vs-kdown-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-histogram-sip-mcsplit-vs-mcsplit-par-t32.pdf}
-
-    \caption{Per-instance speedups for vertex-and edge-labelled, directed graphs.}
+    \caption{Per-instance speedups. The $x$-axis is sequential performance and the $y$-axis is 32
+    threaded performance, so points below the diagonal line represent a speedup.}
 \end{figure*}
 
-\begin{figure*}[p]
+\begin{figure}[tb]
     \includegraphics*{gen-graph-scatter-33ved-clique-par-t32r-vs-clique-par-t32.pdf}
@@
     \caption{Scalability and reproducibility, clique algorithm.}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[p]
+\begin{figure}[tb]
     \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t32r-vs-clique-cilk-t32.pdf}
@@
     \caption{Scalability and reproducibility, clique algorithm, Cilk.}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[p]
+\begin{figure}[tb]
     \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t32r-vs-mcsplit-par-t32.pdf}
@@
     \caption{Scalability and reproducibility, McSplit algorithm.}
-\end{figure*}
+\end{figure}
 

### 2017-08-07 paper/paper.tex
Address
Write a tiny bit

@@
 
-\usepackage{microtype}
 \usepackage{amsmath}
@@
 \Crefname{figure}{Figure}{Figures}
+
+\newcommand{\citepos}[1]{\citet{#1}'s}
 
@@
     \institution{Universit\'e Lyon 1, LIRIS, UMR5205}
-    \postcode{F-69621, France}
+    \city{F-69621, France}
 }
@@
 
-%
-% The code below should be generated by the tool at
-% http://dl.acm.org/ccs.cfm
-% Please copy and paste the code instead of the example below. 
-%
 \begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10003752.10003809.10003716.10011136</concept_id>
+<concept_desc>Theory of computation~Discrete optimization</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10003752.10003809.10010170</concept_id>
+<concept_desc>Theory of computation~Parallel algorithms</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10003752.10003809.10011254.10011256</concept_id>
+<concept_desc>Theory of computation~Branch-and-bound</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10003752.10003790.10003795</concept_id>
+<concept_desc>Theory of computation~Constraint and logic programming</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
 \end{CCSXML}
 
-% \ccsdesc[500]{Computer systems organization~Embedded systems}
-% \ccsdesc[300]{Computer systems organization~Redundancy}
-% \ccsdesc{Computer systems organization~Robotics}
-% \ccsdesc[100]{Networks~Network reliability}
+\ccsdesc[500]{Theory of computation~Discrete optimization}
+\ccsdesc[500]{Theory of computation~Parallel algorithms}
+\ccsdesc[500]{Theory of computation~Branch-and-bound}
+\ccsdesc[300]{Theory of computation~Constraint and logic programming}
 
-% \keywords{ACM proceedings, \LaTeX, text tagging}
+\keywords{Maximum common subgraph, Parallel branch and bound, Constraint programming}
 
@@
 
+?? Explain induced, give examples.
+
+Finding a maximum common subgraph is the key step in measuring the similarity or difference between
+two graphs \citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}: to determine
+the difference between two graphs, we find what they have in common, and then take everything left
+over. Because of this, maximum common subgraph problems frequently arise in biology and chemistry
+\citep{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
+represent molecules, but also in applications including computer vision
+\citep{DBLP:journals/jair/CookH94,DBLP:conf/gbrpr/CombierDS13}, computer-aided manufacturing
+\citep{o:LuoWSN17}, crisis management \citep{o:DelavalladeFLL16}, deanonymising datasets
+\citep{o:SharadD13}, the analysis of source code \citep{DBLP:journals/tkde/DjokoCH97}, binary
+programs \citep{DBLP:conf/icics/GaoRS08}, and circuit designs \citep{DBLP:journals/jair/CookH94}, in
+malware detection \citep{DBLP:journals/compsec/ParkRS13}, social network analysis
+\citep{DBLP:journals/tkde/FangYZZ15}, graph database query explanations
+\citep{DBLP:journals/jcss/VasilyevaTBL16}, and in character recognition problems
+\citep{DBLP:journals/pr/LuRS91}.
+
+?? Connected, which occurs in many applications
+\citep{DBLP:journals/tcs/Koch01,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
+Adding the connectedness requirement makes certain special cases solvable in polynomial time,
+including outerplanar graphs of bounded degree \citep{DBLP:journals/algorithms/AkutsuT13} and trees
+\citep{DBLP:conf/mfcs/DroschinskyKM16}, but the general case remains \NP-hard. As illustrated in
+\cref{figure:mcsexample}, the maximum common connected subgraph cannot be deduced from the maximum
+common subgraph.
+
+This paper looks at three recent branch and bound algorithms, each of which is the state of the art
+for certain classes of instance.
+
+The ultimate aim of this chapter is to parallelise a state-of-the-art algorithm for maximum common
+subgraph problems, and to see whether we agree with \citepos{DBLP:conf/ictai/MinotNS15} conclusion
+that doing so is unusually challenging.
+
 \section{Sequential Algorithms}
 
-\subsection{A Basic Constraint Model}
+The maximum common subgraph problem may be reformulated as a constraint optimisation problem, as
+follows. We pick whichever input graph has fewer vertices, and call it the \emph{pattern}; the other
+graph is called the \emph{target}. The model is then to construct an injective, partial mapping from
+the pattern to the target, preserving both adjacency and non-adjacency. Thus, for each
+vertex in the pattern, we create a variable, whose domain ranges over each vertex in the target
+graph, plus an additional value $\bot$. We then have three sets of constraints. The first set says
+that for each pair of adjacent vertices in the pattern (that is, for each edge in the pattern),
+if neither of these vertices are mapped to $\bot$ then these vertices must be mapped to an adjacent
+pair of target vertices. The second set is similar, but looks at non-adjacent pairs (or non-edges).
+Finally, the third set ensures injectivity, by enforcing that the variables must be all different
+except when using $\bot$.
 
-\subsection{Ensuring Connectedness}
+?? Either by binary constraints between all pairs, or a specialised propagator.
+
+?? Objective.
+
+?? One approach is branch and bound search over this model. Pick a variable, ... . cite.
+
+?? Ensuring connectedness
 
@@
 
-\subsection{The Clique Encoding}
-
 \subsection{The $k$-less Subgraph Isomorphism Algorithm}
@@
 The iteration of the decision problem (iterating over $k$) leads to an algorithm with rivalling runtimes.
+
+\subsection{The Clique Encoding}
 
@@
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{dblp}
+\bibliography{combined}
 

### 2017-08-07 paper/paper.tex
More things

@@
 including outerplanar graphs of bounded degree \citep{DBLP:journals/algorithms/AkutsuT13} and trees
-\citep{DBLP:conf/mfcs/DroschinskyKM16}, but the general case remains \NP-hard. As illustrated in
+\citep{DBLP:conf/mfcs/DroschinskyKM16}, but the general case remains NP-hard. As illustrated in
 \cref{figure:mcsexample}, the maximum common connected subgraph cannot be deduced from the maximum
@@
 
-\subsection{The Splitting Algorithm}
+\subsection{Domain Splitting}
 
@@
 
-\subsection{The $k$-less Subgraph Isomorphism Algorithm}
+\subsection{$k$-less Subgraph Isomorphism}
 
@@
 
-\subsection{The Clique Encoding}
+\subsection{Encoding as Maximum Clique}
 
@@
 \begin{figure*}[tb]
-    \includegraphics*{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}
+    \includegraphics*{gen-graph-as-plain.pdf}
     \hfill
-    \includegraphics*{gen-graph-scatter-plain-kdown-vs-kdown-par-t32.pdf}
+    \includegraphics*{gen-graph-as-33v.pdf}
     \hfill
-    \includegraphics*{gen-graph-scatter-plain-mcsplit-vs-mcsplit-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33v-clique-vs-clique-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33v-mcsplit-vs-mcsplit-par-t32.pdf}
+    \includegraphics*{gen-graph-as-33ved.pdf}
 
@@
 
-    \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t32.pdf}
+    \includegraphics*{gen-graph-as-plain-connected.pdf}
     \hfill
-    \includegraphics*{gen-graph-scatter-33ved-mcsplit-vs-mcsplit-par-t32.pdf}
+    \includegraphics*{gen-graph-as-33ve-connected.pdf}
     \hfill
-    \includegraphics*{gen-graph-scatter-sip-kdown-vs-kdown-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-sip-mcsplit-vs-mcsplit-par-t32.pdf}
-
-    \caption{Per-instance speedups. The $x$-axis is sequential performance and the $y$-axis is 32
-    threaded performance, so points below the diagonal line represent a speedup.}
+    \includegraphics*{gen-graph-as-sip.pdf}
 \end{figure*}
 
-\begin{figure}[tb]
-    \includegraphics*{gen-graph-scatter-33ved-clique-par-t32r-vs-clique-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t2.pdf}
-
-    \vspace*{1em}
-
-    \includegraphics*{gen-graph-scatter-33ved-clique-par-t2-vs-clique-par-t4.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33ved-clique-par-t4-vs-clique-par-t8.pdf}
-
-    \vspace*{1em}
-
-    \includegraphics*{gen-graph-scatter-33ved-clique-par-t8-vs-clique-par-t16.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33ved-clique-par-t16-vs-clique-par-t32.pdf}
-
-    \caption{Scalability and reproducibility, clique algorithm.}
-\end{figure}
-
-\begin{figure}[tb]
-    \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t32r-vs-clique-cilk-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-cilk-t2.pdf}
-
-    \vspace*{1em}
-
-    \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t2-vs-clique-cilk-t4.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t4-vs-clique-cilk-t8.pdf}
-
-    \vspace*{1em}
-
-    \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t8-vs-clique-cilk-t16.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t16-vs-clique-cilk-t32.pdf}
-
-    \caption{Scalability and reproducibility, clique algorithm, Cilk.}
-\end{figure}
-
-\begin{figure}[tb]
-    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t32r-vs-mcsplit-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-plain-mcsplit-vs-mcsplit-par-t2.pdf}
-
-    \vspace*{1em}
-
-    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t2-vs-mcsplit-par-t4.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t4-vs-mcsplit-par-t8.pdf}
-
-    \vspace*{1em}
-
-    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t8-vs-mcsplit-par-t16.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t16-vs-mcsplit-par-t32.pdf}
-
-    \caption{Scalability and reproducibility, McSplit algorithm.}
-\end{figure}
+% \begin{figure*}[tb]
+%     \includegraphics*{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-plain-kdown-vs-kdown-par-t32.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-plain-mcsplit-vs-mcsplit-par-t32.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-33v-clique-vs-clique-par-t32.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-33v-mcsplit-vs-mcsplit-par-t32.pdf}
+% 
+%     \vspace*{1em}
+% 
+%     \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t32.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-33ved-mcsplit-vs-mcsplit-par-t32.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-sip-kdown-vs-kdown-par-t32.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-sip-mcsplit-vs-mcsplit-par-t32.pdf}
+% 
+%     \caption{Per-instance speedups. The $x$-axis is sequential performance and the $y$-axis is 32
+%     threaded performance, so points below the diagonal line represent a speedup.}
+% \end{figure*}
+% 
+% \begin{figure}[tb]
+%     \includegraphics*{gen-graph-scatter-33ved-clique-par-t32r-vs-clique-par-t32.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t2.pdf}
+% 
+%     \vspace*{1em}
+% 
+%     \includegraphics*{gen-graph-scatter-33ved-clique-par-t2-vs-clique-par-t4.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-33ved-clique-par-t4-vs-clique-par-t8.pdf}
+% 
+%     \vspace*{1em}
+% 
+%     \includegraphics*{gen-graph-scatter-33ved-clique-par-t8-vs-clique-par-t16.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-33ved-clique-par-t16-vs-clique-par-t32.pdf}
+% 
+%     \caption{Scalability and reproducibility, clique algorithm.}
+% \end{figure}
+% 
+% \begin{figure}[tb]
+%     \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t32r-vs-clique-cilk-t32.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-cilk-t2.pdf}
+% 
+%     \vspace*{1em}
+% 
+%     \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t2-vs-clique-cilk-t4.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t4-vs-clique-cilk-t8.pdf}
+% 
+%     \vspace*{1em}
+% 
+%     \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t8-vs-clique-cilk-t16.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t16-vs-clique-cilk-t32.pdf}
+% 
+%     \caption{Scalability and reproducibility, clique algorithm, Cilk.}
+% \end{figure}
+% 
+% \begin{figure}[tb]
+%     \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t32r-vs-mcsplit-par-t32.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-plain-mcsplit-vs-mcsplit-par-t2.pdf}
+% 
+%     \vspace*{1em}
+% 
+%     \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t2-vs-mcsplit-par-t4.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t4-vs-mcsplit-par-t8.pdf}
+% 
+%     \vspace*{1em}
+% 
+%     \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t8-vs-mcsplit-par-t16.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t16-vs-mcsplit-par-t32.pdf}
+% 
+%     \caption{Scalability and reproducibility, McSplit algorithm.}
+% \end{figure}
 

### 2017-08-07 paper/asify.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+join -1 2 -2 2 <(grep '^[^#]' $1 | sort -k2 ) <(grep '[^#]' $2 | sort -k2 ) | cut -d' ' -f1,2,4 | sort -n -k3,3

### 2017-08-07 paper/paper.tex
Muck around with plotting some more, starting to like this

@@
 
-    \caption{Cumulative performance. The parallel versions, dotted, are always better in aggregate
-    than the sequential versions, solid.}
+    \caption{The cumulative number of instances solved over time, for different families and
+    algorithms. The 32 threaded parallel versions (shown using dotted lines) are always better in
+    aggregate than the sequential versions (shown using solid lines).}
 \end{figure*}
 
+% \begin{figure*}[tb]
+%     \includegraphics*{gen-graph-as-plain.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-as-33v.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-as-33ved.pdf}
+% 
+%     \vspace*{1em}
+% 
+%     \includegraphics*{gen-graph-as-plain-connected.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-as-33ve-connected.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-as-sip.pdf}
+% 
+%     \caption{Aggregate speedups.}
+% \end{figure*}
+
 \begin{figure*}[tb]
-    \includegraphics*{gen-graph-as-plain.pdf}
+    \includegraphics*{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}
     \hfill
-    \includegraphics*{gen-graph-as-33v.pdf}
+    \includegraphics*{gen-graph-scatter-33v-clique-vs-clique-par-t32.pdf}
     \hfill
-    \includegraphics*{gen-graph-as-33ved.pdf}
+    \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-connected-clique-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ve-connected-clique-vs-clique-par-t32.pdf}
 
@@
 
-    \includegraphics*{gen-graph-as-plain-connected.pdf}
+    \includegraphics*{gen-graph-histogram-plain-clique-vs-clique-par-t32.pdf}
     \hfill
-    \includegraphics*{gen-graph-as-33ve-connected.pdf}
+    \includegraphics*{gen-graph-histogram-33v-clique-vs-clique-par-t32.pdf}
     \hfill
-    \includegraphics*{gen-graph-as-sip.pdf}
+    \includegraphics*{gen-graph-histogram-33ved-clique-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-histogram-plain-connected-clique-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-histogram-33ve-connected-clique-vs-clique-par-t32.pdf}
+
+    \caption{On the top row, per-instance speedups, using the clique algorithm. The $x$-axis is
+    sequential performance and the $y$-axis is 32 threaded performance, so points below the diagonal
+    line represent a speedup. Darker points represent instances where the solution is relatively
+    large compared to the order of the input graphs. Below, histograms plotting the distribution of
+    speedups for instances whose sequential runtime was at least 500 milliseconds, and below the
+    timeout.}
 \end{figure*}
+
+\begin{figure}[tb]
+    \hfill
+    \includegraphics{gen-graph-as-clique.pdf}
+    \hfill
+
+    \caption{Aggregate speedups from the clique algorithm using 32 threads, for each supported
+    family of instances.}
+\end{figure}
 
@@
 % 
-% \begin{figure}[tb]
-%     \includegraphics*{gen-graph-scatter-33ved-clique-par-t32r-vs-clique-par-t32.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t2.pdf}
-% 
-%     \vspace*{1em}
-% 
-%     \includegraphics*{gen-graph-scatter-33ved-clique-par-t2-vs-clique-par-t4.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-33ved-clique-par-t4-vs-clique-par-t8.pdf}
-% 
-%     \vspace*{1em}
-% 
-%     \includegraphics*{gen-graph-scatter-33ved-clique-par-t8-vs-clique-par-t16.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-33ved-clique-par-t16-vs-clique-par-t32.pdf}
-% 
-%     \caption{Scalability and reproducibility, clique algorithm.}
-% \end{figure}
+\begin{figure}[tb]
+    \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t2.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t2-vs-clique-par-t4.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t4-vs-clique-par-t8.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t8-vs-clique-par-t16.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t16-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t32r-vs-clique-par-t32.pdf}
+
+    \caption{Per-instance speedups from the clique algorithm on vertex- and edge-labelled, directed
+    instances, when going from sequential to two threads in the first plot, then increasing the
+    number of threads in subsequent plots. The final plot shows 32 threads versus a repeated run
+    also with 32 threads.}
+\end{figure}
 % 

### 2017-08-07 paper/paper.tex
Fuller figured

@@
 
-% \begin{figure*}[tb]
-%     \includegraphics*{gen-graph-as-plain.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-as-33v.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-as-33ved.pdf}
-% 
-%     \vspace*{1em}
-% 
-%     \includegraphics*{gen-graph-as-plain-connected.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-as-33ve-connected.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-as-sip.pdf}
-% 
-%     \caption{Aggregate speedups.}
-% \end{figure*}
+\subsection{A Closer Look at Clique}
 
@@
 \begin{figure}[tb]
-    \hfill
     \includegraphics{gen-graph-as-clique.pdf}
-    \hfill
 
@@
 % \end{figure}
+
+\subsection{A Closer Look at $k{\downarrow}$}
+
+\begin{figure}[tb]
+    \includegraphics*{gen-graph-scatter-plain-kdown-vs-kdown-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-sip-kdown-vs-kdown-par-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics{gen-graph-as-kdown.pdf}
+
+    \caption{On the top row, per-instance speedups, using the $k{\downarrow}$ algorithm. The
+    $x$-axis is sequential performance and the $y$-axis is 32 threaded performance. Below,
+    aggregate speedups for both families.}
+\end{figure}
+
+\subsection{A Closer Look at McSplit}
+
+\begin{figure}[tb]
+    \includegraphics*{gen-graph-scatter-plain-mcsplit-vs-mcsplit-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33v-mcsplit-vs-mcsplit-par-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-33ved-mcsplit-vs-mcsplit-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-connected-mcsplit-vs-mcsplit-par-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics*{gen-graph-scatter-33ve-connected-mcsplit-vs-mcsplit-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-sip-mcsplit-vs-mcsplit-par-t32.pdf}
+
+    \vspace*{1em}
+
+    \includegraphics{gen-graph-as-mcsplit.pdf}
+
+    \caption{On the top three rows, per-instance speedups, using the splitting algorithm. The
+    $x$-axis is sequential performance and the $y$-axis is 32 threaded performance. On the final
+    row, aggregate speedups for both families.}
+\end{figure}
+
+\subsection{Old Text}
 

### 2017-08-08 paper/asify.sh
Slightly betterer

@@ -1,3 +1,3 @@
 #!/bin/bash
 
-join -1 2 -2 2 <(grep '^[^#]' $1 | sort -k2 ) <(grep '[^#]' $2 | sort -k2 ) | cut -d' ' -f1,2,4 | sort -n -k3,3
+join -1 2 -2 2 <(grep 'i$' $1 | sort -k2 ) <(grep 'i$' $2 | sort -k2 ) | cut -d' ' -f1,2,4 | sort -n -k3,3

### 2017-08-08 paper/paper.tex
Maybe we want reproduction plots
Write more, pretend it's progress

@@
 
-\section{Experiments with Parallel Search}
+\section{Parallel Search}
 
-?? Hardware.
+?? Approach
 
-?? Datasets.
+\section{Empirical Evaluation}
+
+We perform our experiments on systems with dual Intel Xeon E5-2640 v2 processors and 64GBytes RAM,
+running Ubuntu 14.04, with GCC 5.3.0 as the compiler. Each machine has a total of sixteen cores, and
+hyper-threading is enabled, so we run experiments with 32 threads except where otherwise noted.
+Most of our instances come from a standard benchmark set
+\citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07} for maximum common subgraph
+problems.  This benchmark set can be used in a number of ways, for different variants of the
+problem. We use it to create five families of instance, as follows:
+
+\begin{description}
+    \item[Unlabelled] instances, by selecting the first ten members of each parameter class where the
+        graphs have up to 50 vertices each---this gives us a total of 4,110 instances.
+
+    \item[Vertex labelled] instances, by selecting the first ten members of each parameter class
+        (and so graphs have up to 100 vertices each), applying the 33\% labelling scheme for
+        vertices only, and treating edges as undirected. This gives 8,140 instances.
+
+    \item[Both labelled, directed] instances, by selecting the first ten members of each parameter
+        class, and applying the 33\% labelling scheme to both vertices and edges. Again, this gives
+        8,140 instances.
+
+    \item[Unlabelled, connected] instances, using the same instances as the \emph{unlabelled} case.
+
+    \item[Both labelled, connected] instances, starting in the same way as the \emph{both labelled,
+        directed} case. These are then converted to undirected graphs by treating edges as
+        undirected, picking the label of the lower-numbered edge.
+\end{description}
+
+\noindent
+Following \citet{DBLP:conf/aaai/HoffmannMR17}, we also work with the 5,725 \textbf{Large} instances
+originally used for studying portfolios of subgraph isomorphism algorithms
+\citep{DBLP:conf/lion/KotthoffMS16}. These graphs are unlabelled and undirected, and can include up
+to 6,671 vertices. We do not use the clique encoding on these instances due to its impractical
+memory requirements for large graphs.
 
@@
 
+\begin{figure}[tb]
+    \includegraphics*{gen-graph-scatter-plain-kdown-par-t4-vs-kdown-par-t16.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-kdown-par-t32r-vs-kdown-par-t32.pdf}
+
+    \caption{Scalability and reproducibility of the $k{\downarrow}$ algorithm.}
+\end{figure}
+
 \subsection{A Closer Look at McSplit}
@@
     row, aggregate speedups for both families.}
+\end{figure}
+
+\begin{figure}[tb]
+    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t4-vs-mcsplit-par-t16.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t32r-vs-mcsplit-par-t32.pdf}
+
+    \caption{Scalability and reproducibility of the splitting algorithm.}
 \end{figure}

### 2017-08-09 paper/paper.tex
More things

@@
 \documentclass[sigconf]{acmart}
-
-\usepackage{showframe}
 
@@
 \newcommand{\citepos}[1]{\citet{#1}'s}
+
+% pretty colours
+\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
+\definecolor{magmapurple}{rgb}{0.505882, 0.145098, 0.505882}
+\definecolor{magmaorange}{rgb}{0.984314, 0.529412, 0.380392}
 
@@
 \section{Introduction}
+
+\begin{figure}[b]
+    \centering
+    \begin{tikzpicture}[scale=0.33]%{{{
+        \begin{scope}
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
+            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
+
+            \draw [very thick] (M1) -- (M2);
+            \draw [very thick] (M2) -- (M4);
+            \draw [very thick] (M3) -- (M5);
+            \draw [color=uofgsandstone!60] (M4) -- (M6);
+            \draw [color=uofgsandstone!60] (M5) -- (M6);
+            \draw [very thick] (M3) -- (M1);
+            \draw [color=uofgsandstone!60] (M6) -- (M7);
+            \draw [color=uofgsandstone!60] (M6) -- (M8);
+            \draw [color=uofgsandstone!60] (M6) -- (M9);
+            \draw [very thick] (M7) -- (M9);
+            \draw [very thick] (M8) -- (M9);
+        \end{scope}
+
+        \begin{scope}[xshift=5cm]
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
+            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};
+
+            \draw [very thick] (M1) -- (M2);
+            \draw [very thick] (M2) -- (M4);
+            \draw [very thick] (M3) -- (M5);
+            \draw [color=uofgsandstone!60] (M4) -- (M6);
+            \draw [color=uofgsandstone!60] (M5) -- (M6);
+            \draw [very thick] (M3) -- (M1);
+            \draw [color=uofgsandstone!60] (M6) -- (M7);
+            \draw [very thick] (M7) -- (M8);
+            \draw [very thick] (M8) -- (M9);
+        \end{scope}
+
+        \begin{scope}[xshift=12cm]
+            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
+            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
+
+            \draw [very thick] (M1) -- (M2);
+            \draw [very thick] (M2) -- (M4);
+            \draw [very thick] (M3) -- (M5);
+            \draw [very thick] (M4) -- (M6);
+            \draw [very thick] (M5) -- (M6);
+            \draw [very thick] (M3) -- (M1);
+            \draw [color=uofgsandstone!60] (M6) -- (M7);
+            \draw [very thick] (M6) -- (M8);
+            \draw [color=uofgsandstone!60] (M6) -- (M9);
+            \draw [color=uofgsandstone!60] (M7) -- (M9);
+            \draw [color=uofgsandstone!60] (M8) -- (M9);
+        \end{scope}
+
+        \begin{scope}[xshift=17cm]
+            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};
+
+            \draw [very thick] (M1) -- (M2);
+            \draw [very thick] (M2) -- (M4);
+            \draw [very thick] (M3) -- (M5);
+            \draw [very thick] (M4) -- (M6);
+            \draw [very thick] (M5) -- (M6);
+            \draw [very thick] (M3) -- (M1);
+            \draw [very thick] (M6) -- (M7);
+            \draw [draw=uofgsandstone!60] (M7) -- (M8);
+            \draw [draw=uofgsandstone!60] (M8) -- (M9);
+        \end{scope}
+\end{tikzpicture}
+\caption{A maximum common subgraph of the first two graphs has eight vertices, shaded. A maximum
+common \emph{connected} subgraph, shown on the right, has only seven vertices.}
+\label{figure:mcsexample}
+\end{figure}
 
@@
     algorithms. The 32 threaded parallel versions (shown using dotted lines) are always better in
-    aggregate than the sequential versions (shown using solid lines).}
+aggregate than the sequential versions (shown using solid lines).}\label{figure:cumulative}
 \end{figure*}
+
+In \cref{figure:cumulative} we plot empirical cumulative distribution functions showing the number
+of instances solved over time, for both sequential (solid lines) and parallel (dotted lines)
+versions of each algorithm.  Cumulative plots are commonly used in constraint programming and
+boolean satisfiability for comparing aggregate behaviour of solvers (sometimes with the axes
+reversed, in which case they are called \emph{cactus} plots).  To read these plots, make a choice of
+timeout along the $x$-axis, which uses a log scale. The $y$ value at that point shows the number of
+instances whose runtime (individually) is at most $x$, for a particular algorithm. In other words,
+at any given $x$ value, the highest line shows which algorithm is able to solve the largest number
+of instances using a per-instance timeout of that $x$ value, bearing in mind that the actual sets of
+instances solved by each algorithm may be completely different.
+
+Each plot gives the same conclusion: if we are working with a timeout of at least $10^2$
+milliseconds, then for any problem family and any sequential algorithm, if given the option of
+switching to the corresponding parallel algorithm, then we should do so.
 

### 2017-08-09 paper/paper.tex
Move
Thing

@@
 
-?? Explain induced, give examples.
+A \emph{subgraph isomorphism} is an injective mapping from a \emph{pattern} graph to a \emph{target}
+graph which maps adjacent vertices to adjacent vertices---that is, it \emph{preserves adjacency}. It
+is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent vertices, preserving
+non-adjacency. When working with labelled graphs, a subgraph isomorphism must preserve labels, and
+on directed graphs, it must preserve orientation. A \emph{common induced subgraph} of two graphs $G$
+and $H$ is a pair of induced subgraph isomorphisms from a pattern graph $P$, one to $G$ and one to
+$H$. A \emph{maximum common induced subgraph} is one with as many vertices as possible.
 
@@
 \citep{DBLP:journals/jcss/VasilyevaTBL16}, and in character recognition problems
-\citep{DBLP:journals/pr/LuRS91}.
+\citep{DBLP:journals/pr/LuRS91}.  A common variant of the problem requires a largest
+\emph{connected} subgraph
+\citep{DBLP:journals/tcs/Koch01,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
+As illustrated in \cref{figure:mcsexample}, the maximum common connected subgraph cannot be deduced
+from the maximum common subgraph.
 
-?? Connected, which occurs in many applications
-\citep{DBLP:journals/tcs/Koch01,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
-Adding the connectedness requirement makes certain special cases solvable in polynomial time,
-including outerplanar graphs of bounded degree \citep{DBLP:journals/algorithms/AkutsuT13} and trees
-\citep{DBLP:conf/mfcs/DroschinskyKM16}, but the general case remains NP-hard. As illustrated in
-\cref{figure:mcsexample}, the maximum common connected subgraph cannot be deduced from the maximum
-common subgraph.
-
-This paper looks at three recent branch and bound algorithms, each of which is the state of the art
-for certain classes of instance.
-
-The ultimate aim of this chapter is to parallelise a state-of-the-art algorithm for maximum common
-subgraph problems, and to see whether we agree with \citepos{DBLP:conf/ictai/MinotNS15} conclusion
-that doing so is unusually challenging.
+Although both variants are NP-hard, some progress has been made towards solving the problem in
+practice.  This paper looks at three recent branch and bound algorithms for maximum common
+(connected) induced subgraph problems, each of which is the state of the art for certain classes of
+instance. We discuss our experiences in adding parallel tree-search to these three algorithms. In
+each case, our results show that the parallel version of the algorithm is clearly better than the
+sequential version, although a closer look at the results shows many nuances.
 
@@
 
+\subsection{Encoding as Maximum Clique}
+
+\subsection{Constraint Programming}
+
 The maximum common subgraph problem may be reformulated as a constraint optimisation problem, as
-follows. We pick whichever input graph has fewer vertices, and call it the \emph{pattern}; the other
-graph is called the \emph{target}. The model is then to construct an injective, partial mapping from
-the pattern to the target, preserving both adjacency and non-adjacency. Thus, for each
-vertex in the pattern, we create a variable, whose domain ranges over each vertex in the target
-graph, plus an additional value $\bot$. We then have three sets of constraints. The first set says
-that for each pair of adjacent vertices in the pattern (that is, for each edge in the pattern),
-if neither of these vertices are mapped to $\bot$ then these vertices must be mapped to an adjacent
-pair of target vertices. The second set is similar, but looks at non-adjacent pairs (or non-edges).
-Finally, the third set ensures injectivity, by enforcing that the variables must be all different
-except when using $\bot$.
+follows.  Observe that an equivalent definition of a common subgraph of graphs $G$ and $H$ is an
+injective partial mapping from $G$ to $H$ which preserves both adjacency and non-adjacency. Hence we
+pick whichever input graph has fewer vertices, and call it the \emph{pattern}; the other graph is
+called the \emph{target}. The model is then to construct an injective, partial mapping from the
+pattern to the target, preserving both adjacency and non-adjacency. Thus, for each vertex in the
+pattern, we create a variable, whose domain ranges over each vertex in the target graph, plus an
+additional value $\bot$. We then have three sets of constraints. The first set says that for each
+pair of adjacent vertices in the pattern (that is, for each edge in the pattern), if neither of
+these vertices are mapped to $\bot$ then these vertices must be mapped to an adjacent pair of target
+vertices. The second set is similar, but looks at non-adjacent pairs (or non-edges).  Finally, the
+third set ensures injectivity, by enforcing that the variables must be all different except when
+using $\bot$.
 
@@
 The iteration of the decision problem (iterating over $k$) leads to an algorithm with rivalling runtimes.
-
-\subsection{Encoding as Maximum Clique}
 

### 2017-08-09 paper/paper.tex
This is definitely progress

@@
 \usepackage{tikz}
+
+\usepackage{placeins}
 
@@
 %Conference
-\acmConference[$\textnormal{IA}^3$ 2017]{Seventh Workshop on Irregular Applications: Architectures and Algorithms}{November 2017}{Denver, Colorado USA}
+\acmConference[IA$^3$ 2017]{Seventh Workshop on Irregular Applications: Architectures and Algorithms}{November 2017}{Denver, Colorado USA}
 \acmYear{2017}
@@
 \citep{o:LuoWSN17}, crisis management \citep{o:DelavalladeFLL16}, deanonymising datasets
-\citep{o:SharadD13}, the analysis of source code \citep{DBLP:journals/tkde/DjokoCH97}, binary
-programs \citep{DBLP:conf/icics/GaoRS08}, and circuit designs \citep{DBLP:journals/jair/CookH94}, in
+\citep{o:SharadD13}, the analysis of source code \citep{DBLP:journals/tkde/DjokoCH97} and binary
+programs \citep{DBLP:conf/icics/GaoRS08}, in
 malware detection \citep{DBLP:journals/compsec/ParkRS13}, social network analysis
@@
 \citep{DBLP:journals/tcs/Koch01,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
-As illustrated in \cref{figure:mcsexample}, the maximum common connected subgraph cannot be deduced
+As illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be determined
 from the maximum common subgraph.
@@
 
-\subsection{Encoding as Maximum Clique}
+There are three competitive approaches for the maximum common subgraph problem, each being the
+strongest on certain classes of instance. The first involves a reduction to the maximum clique
+problem, whilst the other two approaches have a constraint programming feel to them.
+
+\subsection{Reduction to Maximum Clique}
+
+A clique in a graph is a subgraph where every vertex is adjacent to every other. There is a
+well-known reduction from the maximum common subgraph problem to the problem of finding a maximum
+clique in an \emph{association graph}
+\citep{o:Levi73,DBLP:journals/siamcomp/BalasY86,o:DurandPBT99,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/cp/McCreeshNPS16}, which, when combined with
+a modern maximum clique solver \citep{DBLP:journals/ol/SegundoMRH13}, is the current best approach for
+solving the problem on labelled graphs \citep{DBLP:conf/cp/McCreeshNPS16}. A modified clique-like
+algorithm can also be used to solve the maximum common connected subgraph problem, by ensuring
+connectedness during search \citep{DBLP:conf/cp/McCreeshNPS16}; again, this is the best known way of
+solving the problem on labelled graphs.
+
+A major downside of the maximum clique approach is that the association graph encoding is extremely
+memory-intensive, limiting its practical use to pairs of graphs with no more than a few hundred
+vertices.
 
@@
 follows.  Observe that an equivalent definition of a common subgraph of graphs $G$ and $H$ is an
-injective partial mapping from $G$ to $H$ which preserves both adjacency and non-adjacency. Hence we
-pick whichever input graph has fewer vertices, and call it the \emph{pattern}; the other graph is
-called the \emph{target}. The model is then to construct an injective, partial mapping from the
-pattern to the target, preserving both adjacency and non-adjacency. Thus, for each vertex in the
-pattern, we create a variable, whose domain ranges over each vertex in the target graph, plus an
-additional value $\bot$. We then have three sets of constraints. The first set says that for each
-pair of adjacent vertices in the pattern (that is, for each edge in the pattern), if neither of
-these vertices are mapped to $\bot$ then these vertices must be mapped to an adjacent pair of target
-vertices. The second set is similar, but looks at non-adjacent pairs (or non-edges).  Finally, the
-third set ensures injectivity, by enforcing that the variables must be all different except when
-using $\bot$.
-
-?? Either by binary constraints between all pairs, or a specialised propagator.
-
-?? Objective.
-
-?? One approach is branch and bound search over this model. Pick a variable, ... . cite.
+injective \emph{partial} mapping from $G$ to $H$ which preserves both adjacency and non-adjacency.
+Hence we pick whichever input graph has fewer vertices, and call it the \emph{pattern}; the other
+graph is called the \emph{target}. The model then follows from this new definition: for each vertex
+in the pattern, we create a variable, whose domain ranges over each vertex in the target graph, plus
+an additional value $\bot$ representing an unmapped vertex. We then have three sets of constraints.
+The first set says that for each pair of adjacent vertices in the pattern (that is, for each edge in
+the pattern), if neither of these vertices are mapped to $\bot$ then these vertices must be mapped
+to an adjacent pair of target vertices. The second set is similar, but looks at non-adjacent pairs
+(or non-edges).  Finally, the third set ensures injectivity, by enforcing that the variables must be
+all different except when using $\bot$. This final set of constraints may either be implemented
+using binary constraints between all pairs of variables, or a special global ``all different except
+$\bot$'' propagator \citep{DBLP:conf/cp/PetitRB01}. The objective is simply to find an assignment of
+values to variables, maximising the number of variables not set to $\bot$. The state of the art for
+this technique is a dedicated implementation of a forward-checking branch and bound search over this
+model \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}.
 
@@
 
-Due to the special structure of the maximum common subgraph problem, the following property holds
-throughout the search process using the CP model: any two variables either have domains with no
-values in common (with the possible exception of $\bot$), or have identical domains. The algorithm
-McSplit exploits this property, while exploring essentially the same search tree as the CP model.
-Rather than storing a domain for each vertex in $\operatorname{V}(G)$, which results in redundant
-storage in cases where domains are identical, equivalence classes of vertices in
-$\operatorname{V}(G)$ are stored, and a set of candidate vertices in $H$ is maintained for each
-equivalence class. This enables fast propagation of the constraints and smaller memory requirements.
-In addition, the data structure enables good branching heuristics to be calculated cheaply.
+\citet{o:McCreeshPT17} observe that due to the special structure of the maximum common subgraph
+problem, the following property holds throughout the search process using the constraint programming
+model: any two variables either have domains with no values in common (with the possible exception
+of $\bot$), or have identical domains. The McSplit algorithm exploits this property. It explores
+essentially the same search tree as the basic forward-checking constraint programming model, but
+using different supporting algorithms and data structures.  Rather than storing a domain for each
+vertex in the pattern graph, equivalence classes of vertices in both graphs are stored in a special
+data structure which is modified in-place and restored upon backtracking. This enables fast
+propagation of the constraints and smaller memory requirements. In addition, this data structure
+enables stronger branching heuristics to be calculated cheaply.
 
@@
 
-The algorithm as described by Hoffmann, McCreesh, and Reilly \cite{DBLP:conf/aaai/HoffmannMR17} approaches the subgraph isomorphism problem by asking the question ``if a pattern graph cannot be found in the target, how much of the pattern graph can be found?''.
-The $k{\downarrow}$ algorithm tries to solve the subgraph isomorphism problem first for $k=0$ (the whole pattern graph can be found in the target), should that be not satisfiable, it tries to solve the problem for $k=1$ (one vertex cannot be matched) should that also not be satisfiable, it iterates $k$ further by a small number to keep the shrinking pattern graph within a reasonable size.
-By unmatching some vertices  invariants are weakened but kept effective.
-Specifically, invariants based around paths and degrees of vertices hold.
+A different take on the constraint programming model is presented by
+\citet{DBLP:conf/aaai/HoffmannMR17}. They approach the maximum common subgraph via the
+subgraph isomorphism problem, asking the question ``if a pattern graph cannot be found in the
+target, how much of the pattern graph can be found?''. The $k{\downarrow}$ algorithm tries to solve
+the subgraph isomorphism problem first for $k=0$ (asking whether the whole pattern graph can be found in the
+target). Should that be not satisfiable, it tries to solve the problem for $k=1$ (one vertex cannot
+be matched), and should that also not be satisfiable, it iteratively increases $k$ until the result
+is satisfiable. The advantage of this approach is that strong invariants using paths and the degrees
+of vertices may be used to prune large portions of the search space.
 
-This technique bridges the subgraph isomorphism problem with the maximal common subgraph problem.
-The iteration of the decision problem (iterating over $k$) leads to an algorithm with rivalling runtimes.
+This algorithm is aimed primarily at large instances, where the two graphs are of different orders,
+and where it is expected that the solution will involve most of the smaller graph (that is, $k$ is
+expected to be low). The implementation we work with only handles unlabelled instances, and does
+not support the connected variant.
 
@@
 
-?? Approach
+\subsection{Anomalies}
+
+\subsection{Parallel Maximum Clique}
+
+?? For the clique algorithm.
+
+?? Discuss the TOPC paper.
+
+?? Reproducible
+
+\subsection{Parallel Constraint-Based Search}
+
+?? Approach for kdown and mcsplit
 
@@
 
+?? There is more to be done.
+
+\FloatBarrier
+
 \bibliographystyle{ACM-Reference-Format}

### 2017-08-12 cilk-clique-code/clique.cc
More options

@@ -366,7 +366,7 @@ namespace
 
                     auto v = p_order[n];
 
-                    cilk_spawn [this, &c, shrinking_p, v] {
+                    if (params.deep && c.size() < 2) {
                         auto new_c = c;
 
                         // consider taking v
@@ -380,7 +380,24 @@ namespace
                             expand(new_c, new_p);
                         else
                             incumbent.update(new_c);
-                    }();
+                    }
+                    else {
+                        cilk_spawn [this, &c, shrinking_p, v] {
+                            auto new_c = c;
+
+                            // consider taking v
+                            new_c.push_back(v);
+
+                            // filter p to contain vertices adjacent to v
+                            FixedBitSet<n_words_> new_p = shrinking_p;
+                            graph.intersect_with_row(v, new_p);
+
+                            if (! new_p.empty())
+                                expand(new_c, new_p);
+                            else
+                                incumbent.update(new_c);
+                        }();
+                    }
 
                     // now consider not taking v
                     shrinking_p.unset(v);

### 2017-08-12 cilk-clique-code/solve_max_common_subgraph.cc

@@ -205,6 +205,8 @@ auto main(int argc, char * argv[]) -> int
             ("unlabelled",                            "Make the graph unlabelled)")
             ("no-edge-labels",                        "Get rid of edge labels, but keep vertex labels")
             ("undirected",                            "Make the graph undirected")
+            ("parallel-for", "Use parallel for")
+            ("deep", "Steal deeper")
             ;
 
         po::options_description all_options{ "All options" };
@@ -244,6 +246,9 @@ auto main(int argc, char * argv[]) -> int
 
         /* Figure out what our options should be. */
         Params params;
+
+        params.parallel_for = options_vars.count("parallel-for");
+        params.deep = options_vars.count("deep");
 
         /* Create graphs */
         auto graphs = std::make_pair(

### 2017-08-12 paper/paper.tex
Words

@@
 
-\section{Parallel Search}
+\section{Benchmark Instances}
 
-\subsection{Anomalies}
-
-\subsection{Parallel Maximum Clique}
-
-?? For the clique algorithm.
-
-?? Discuss the TOPC paper.
-
-?? Reproducible
-
-\subsection{Parallel Constraint-Based Search}
-
-?? Approach for kdown and mcsplit
-
-\section{Empirical Evaluation}
-
-We perform our experiments on systems with dual Intel Xeon E5-2640 v2 processors and 64GBytes RAM,
-running Ubuntu 14.04, with GCC 5.3.0 as the compiler. Each machine has a total of sixteen cores, and
-hyper-threading is enabled, so we run experiments with 32 threads except where otherwise noted.
-Most of our instances come from a standard benchmark set
-\citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07} for maximum common subgraph
-problems.  This benchmark set can be used in a number of ways, for different variants of the
-problem. We use it to create five families of instance, as follows:
+Most of the benchmark instances we will use come from a standard database for maximum common
+subgraph problems \citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}. This benchmark
+set can be used in a number of ways, for different variants of the problem. We use it to create five
+families of instance, as follows:
 
@@
 \citep{DBLP:conf/lion/KotthoffMS16}. These graphs are unlabelled and undirected, and can include up
-to 6,671 vertices. We do not use the clique encoding on these instances due to its impractical
-memory requirements for large graphs.
+to 6,671 vertices. We do not use the clique encoding on these instances due to its
+memory requirements.
+
+\section{Parallel Search}
+
+?? Tree search. History.
+
+\subsection{Anomalies}
+
+\subsection{Parallel Maximum Clique}
+
+?? For the clique algorithm.
+
+?? Discuss the TOPC paper.
+
+?? Reproducible
+
+\subsection{Parallel Constraint-Based Search}
+
+?? Approach for kdown and mcsplit
+
+?? kdown isn't quite conventional branch and bound.
+
+?? mcsplit uses in-place data structures
+
+\section{Empirical Evaluation}
+
+We perform our experiments on systems with dual Intel Xeon E5-2640 v2 processors and 64GBytes RAM,
+running Ubuntu 14.04, with GCC 5.3.0 as the compiler. Each machine has a total of sixteen cores, and
+hyper-threading is enabled, so we run experiments with 32 threads except where otherwise noted.
 

### 2017-08-13 paper/paper.tex
More things

@@
 \documentclass[sigconf]{acmart}
+
+\usepackage{showframe}
 
@@
 \usepackage{tikz}
+\usepackage{nicefrac}
 
@@
 
-?? Tree search. History.
+The clique and $k{\downarrow}$ algorithms already make use of fine granularity bit-parallelism. To
+introduce more coarse-grained thread parallelism, we will parallelise the search process: viewing
+backtracking search as forming a tree, we can explore different portions of the tree on different
+processors.
 
-\subsection{Anomalies}
+?? quick history
+
+?? Anomalies
 
@@
 
-?? For the clique algorithm.
+The maximum clique algorithm underlying . Differs only by a preprocessing step.
 
-?? Discuss the TOPC paper.
+The clique-inspired connected algorithm described by \citet{} may be parallelised in exactly the
+same way.
 
-?? Reproducible
+?? Note the very high number of recursive calls per second. Fitting everything in cache. Not
+compatible with framework abstractions.
+
+?? Splitting depth. Balance.
+
+?? Discuss the TOPC paper.  ?? Reproducible, anomalies, etc
+
+\subsection{Anomalies}
 
@@
 
-\begin{figure*}[tb]
+\begin{figure*}[p]
     \includegraphics*{gen-graph-cumulative-plain.pdf}
@@
 versions of each algorithm.  Cumulative plots are commonly used in constraint programming and
-boolean satisfiability for comparing aggregate behaviour of solvers (sometimes with the axes
-reversed, in which case they are called \emph{cactus} plots).  To read these plots, make a choice of
-timeout along the $x$-axis, which uses a log scale. The $y$ value at that point shows the number of
-instances whose runtime (individually) is at most $x$, for a particular algorithm. In other words,
-at any given $x$ value, the highest line shows which algorithm is able to solve the largest number
-of instances using a per-instance timeout of that $x$ value, bearing in mind that the actual sets of
-instances solved by each algorithm may be completely different.
+boolean satisfiability for comparing aggregate behaviour of solvers in the presence of timeouts.
+(Sometimes the axes are reversed, in which case they are called \emph{cactus} plots.)  To read these
+plots, make a choice of timeout along the $x$-axis, which uses a log scale. The $y$ value at that
+point shows the number of instances whose runtime (individually) is at most $x$, for a particular
+algorithm. In other words, at any given $x$ value, the highest line shows which algorithm is able to
+solve the largest number of instances using a per-instance timeout of that $x$ value, bearing in
+mind that the actual sets of instances solved by each algorithm may be completely different.
 
@@
 
+However, although they are good at showing general trends, cumulative plots can hide interesting
+details. We therefore now take a closer look at each of the three algorithms.
+
 \subsection{A Closer Look at Clique}
 
-\begin{figure*}[tb]
+\begin{figure*}[p]
     \includegraphics*{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}
@@
     speedups for instances whose sequential runtime was at least 500 milliseconds, and below the
-    timeout.}
+    timeout.}\label{figure:cliquespeedups}
 \end{figure*}
 
-\begin{figure}[tb]
+In the top row of \cref{figure:cliquespeedups}, we see scatter plots comparing the sequential and
+parallel runtimes of the clique algorithm on an instance by instance basis, using a log-log plot.
+Instances which timed out using one algorithm but not the other are shown as points along the outer
+axes. Points below the $x{-}y$ diagonal line represent speedups. The colour of the points indicates
+the relative size of the solution---darker points represent instances where the solution uses most
+of the vertices of the input graphs.
+
+?? Quick summary
+
+?? Try to turn this into a number, giving a speedup.
+
+?? Best fit line. Pretend they are outliners. Can get away with it if they are rare enough, but they
+are not for many algorithms.
+
+?? Histograms. Ignore low runtimes. Also have to ignore instances solved in parallel but not
+sequential. Underestimate superlinear speedups. We do not consider this to be a particularly
+satisfactory way of measuring speedups.
+
+?? We propose a different way of . Refer back to the cumulative plots. The usual way of reading
+these plots is to pick whichever line is furthest up, and measuring the vertical difference between
+lines says how many more the parallel algorithm can solve than the sequential algorithm can with a
+particular choice of timeout. However, measuring the \emph{horizontal} distance between lines also
+conveys information. Suppose the sequential algorithm can solve $y$ instances with a selected timeout
+of $s$. By moving to the left on a cumulative plot, we can find the timeout $p$ required for the
+parallel algorithm to solve the same number of instances, bearing in mind that \emph{the two sets of
+instances could in principle have completely different members}. We define the \emph{aggregate
+speedup} to be $\nicefrac{s}{p}$; this can be expressed as a function of time (i.e.\ $s$) or of the
+number of instances solved ($y$).
+
+\begin{figure}[p]
     \includegraphics{gen-graph-as-clique.pdf}
 
-    \caption{Aggregate speedups from the clique algorithm using 32 threads, for each supported
-    family of instances.}
+    \caption{Aggregate speedups from 32 threads on 16 hyper-threaded cores, shown as a function of
+    sequential runtime, for each family supported by the clique algorithm.}
 \end{figure}
@@
 % 
-\begin{figure}[tb]
+\begin{figure}[p]
     \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t2.pdf}
@@
 
-\subsection{Old Text}
-
-Cumulative performance plots comparing each parallel algorithm to its sequential version are given
-in ??, using 32 threads on 16 core hyper-threaded Intel Xeon
-E5-2640 v2 systems. The results are consistent: in each case the parallel algorithm is clearly and
-substantially better than its sequential version, except when the sequential runtime is below 100
-milliseconds.
-
-?? gives instance-by-instance comparisons for a representative subset of
-these algorithm and problem combinations. For the clique algorithm on unlabelled graphs (top left),
-the results show fairly consistent speedups for harder instances: there is only a single instance
-whose sequential runtime is over two seconds where we obtain a speedup of less than ten. We also see
-a significant number of superlinear speedups. This is because there are relatively long proofs of
-optimality for most of the harder instances, but strong solutions are usually found quickly (and
-where they are not, the additional diversity helps). The results are similarly positive on
-labelled graphs (centre left), with speedups of greater than ten being the norm for harder
-instances, although there is more variety in the results.
-
-For $k{\downarrow}$ on the unlabelled maximum common subgraph instances (top centre), the typical speedup
-is between two and ten, and occasionally there are speedups of between ten and one hundred. A few
-instances show small slowdowns: remember that when running 32 threads on a 16 core, hyper-threaded
-system, each individual thread runs somewhat slower than it would if only a single thread were
-running---in other words, these slight slowdowns are due to hardware effects, not changes to the
-search tree.
-
-On the larger subgraph isomorphism instances (bottom left), we see a concentration of speedups
-of around ten. We also see strongly superlinear speedups in some cases, with eight
-instances finishing in under one second which timed out sequentially at one thousand seconds.
-Examining these instances closely shows that in each case, they are satisfiable with either $k = 0$
-or very low values of $k$, and that the superlinear speedups are due to the parallel search
-introducing early diversity, offsetting an incorrect initial value-ordering heuristic choice. We
-also see more instances with slight absolute slowdowns: these are instances where the additional
-threads contribute no helpful work to the solution, and where the slowdown due to not having 32
-times as many resources when using 32 threads is particularly pronounced (memory contention is much
-more of a problem for these larger graphs). Although not ideal, in aggregate these results are rare
-and are more than offset by the superlinear speedups for other instances.
-
-What about the splitting algorithm? On unlabelled instances (top right), our speedups are typically
-between those obtained from the clique algorithm and those from $k{\downarrow}$, being mainly a
-little below ten times. However, in some cases we see absolute slowdowns of up to four times: in
-these cases, parallelism does little, and is not enough to offset the costs we paid when switching
-from an in-place data structure to a data structure which requires copying. Interestingly, we also
-see many more superlinear speedups for this algorithm, which suggests that its value ordering
-heuristics are particularly poor early-on in search. (This last point is not particularly
-surprising: compared to $k{\downarrow}$, the splitting algorithm does not try to reduce domains at
-all at the top of search. The clique encoding, meanwhile, is able to capture even richer information
-about variable-value relationships. The splitting algorithm is fast, but comparatively stupid.)
-
-Similar trends occur with the subgraph isomorphism instances (bottom right), and the absolute
-slowdowns go as high as a factor of ten---this is on instances with particularly large
-target graphs, where copying is most expensive, and where there is a high branching factor at the
-root of the search tree. Again, superlinear speedups are common.
-
-On labelled instances (centre right), the parallel performance is less erratic. Although we no
-longer see absolute slowdowns (excluding on very easy instances), we also see fewer superlinear
-speedups, with most speedups being between one and ten. In many of these instances, parallelism
-does very little: restricting splitting to a depth of five in these cases gives work balance
-problems, but does help us avoid the substantial overheads that come with deeper splitting. This is
-because the search trees for these instances tend to have a very low branching factor---in
-constraint programming terms, the labels make each domain contain only a few values.
-
-?? show that for all three
-algorithms, parallel search is beneficial, and also (reasonably) risk-free: although in a few cases the
-constant factor slowdowns due to hardware limitations and having to make changes to data structures
-are unpleasant, we do not encounter any exponential slowdowns due to search tree changes. But what
-about reproducibility and scalability? Further experiments in ??
-establish that both of these properties also hold. In the left-hand column, we plot each algorithm
-run against itself, using 32 threads in both cases. For the clique and $k{\downarrow}$ approaches,
-every instance is on the central diagonal line, showing an extremely high level of reproducibility.
-For the splitting algorithm, there is a little more variability, particularly for easier instances:
-inspecting the results suggests that this is not down to different amounts of search work being
-performed, but rather whether or not helper threads end up triggering additional copying.
-
-In the right-hand column of ??, we plot the effects of moving from 8
-threads (on the same system) to 32 threads. Points below the diagonal line indicate an improvement.
-For the clique approach, increasing the number of threads is unequivocally beneficial. For the
-$k{\downarrow}$ algorithm, the benefits are slim, and sometimes performance worsens slightly due to
-memory contention (not work done) but we do not see any exponential slowdowns. Finally, for the
-splitting algorithm, the results are usually strictly better, although in a few cases the overheads
-increase further giving a slowdown; we also see an increase in the occurrence of strongly
-superlinear speedups.
-
-It is likely that the parallel results for the splitting algorithm can be improved further. In the
-labelled case, we are often not obtaining a good work balance. However, deeper splitting is
-expensive, particularly with larger graphs. One option to consider would be recomputing parts of the
-search tree to avoid the costs of potentially allowing a copy in the future. This would be extremely
-unpleasant from an implementation perspective, but may be necessary to obtain more even results.
-However, even without this change, the parallel algorithm is clearly preferable to the sequential
-one.
+% \subsection{Cilk}
+%
+% \begin{figure}[tb]
+%     \includegraphics*{gen-graph-scatter-33ved-clique-par-t32r-vs-clique-par-t32.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t32r-vs-clique-cilk-t32.pdf}
+%
+%     \vspace*{1em}
+%
+%     \includegraphics*{gen-graph-scatter-sip-kdown-par-t32r-vs-kdown-par-t32.pdf}
+%     \hfill
+%     \includegraphics*{gen-graph-scatter-sip-kdown-cilk-t32r-vs-kdown-cilk-t32.pdf}
+%
+%     \caption{Reproducibility using Cilk.}
+% \end{figure}
 

### 2017-08-14 paper/paper.tex
Definitely progress

@@
 \usepackage{nicefrac}
-
-\usepackage{placeins}
 
@@
 % The default list of authors is too long for headers}
-\renewcommand{\shortauthors}{R. Hoffmann et al.}
+\renewcommand{\shortauthors}{R. Hoffmann, C. McCreesh et al.}
 
 \begin{abstract}
-    We discuss our experiences adapting three recent algorithms for maximum common (connected) subgraph
-    problems to exploit multi-core parallelism. We consider the effects of different work
-    distribution strategies and their interaction with value-ordering heuristics, and propose a
-    search method which is risk-free, reproducible, and scalable, as well as beneficial. Our results
-    show that each algorithm can be parallelised successfully using this strategy, with the
+    We discuss our experiences adapting three recent algorithms for maximum common (connected)
+    subgraph problems to exploit multi-core parallelism. These algorithms are heavily irregular and
+    very sensitive to search order changes, and so do not easily lend themselves to parallel search.
+    Nonetheless, our results show that each algorithm can be parallelised successfully, with the
     multi-core algorithms we create being clearly better than the sequential ones. We then look in
-    more detail at the results, on an instance by instance basis, and discuss how we should be
-    measuring speedups for these kinds of algorithm.
+    more detail at the results, and discuss how we should be
+    measuring speedups for these kinds of algorithm. Because of the difficulty in measuring an
+    average speedup when so-called anomalous (superlinear and sublinear) speedups are common, we
+    propose a new measure called \emph{aggregate speedup}.
 \end{abstract}
@@
 
-\keywords{Maximum common subgraph, Parallel branch and bound, Constraint programming}
+\keywords{Maximum common subgraph, Parallel branch and bound}
 
@@
 clique in an \emph{association graph}
-\citep{o:Levi73,DBLP:journals/siamcomp/BalasY86,o:DurandPBT99,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/cp/McCreeshNPS16}, which, when combined with
+\citep{o:Levi73,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/cp/McCreeshNPS16}, which, when combined with
 a modern maximum clique solver \citep{DBLP:journals/ol/SegundoMRH13}, is the current best approach for
@@
 
-?? Ensuring connectedness
+Two approaches exist for ensuring connectedness: either a conventional global constraint and
+propagator can be used \citep{DBLP:conf/cp/McCreeshNPS16}, or a special branching rule can be used
+to enforce connectedness during search \citep{DBLP:conf/mco/VismaraV08}. The two techniques are
+broadly comparable performance-wise \citep{DBLP:conf/cp/McCreeshNPS16}, but the branching rule is
+much simpler to implement.
 
@@
 propagation of the constraints and smaller memory requirements. In addition, this data structure
-enables stronger branching heuristics to be calculated cheaply.
+enables stronger branching heuristics to be calculated cheaply. The McSplit algorithm effectively
+dominates conventional constraint programming approaches, being consistently over an order of
+magnitude faster.
 
@@
 introduce more coarse-grained thread parallelism, we will parallelise the search process: viewing
-backtracking search as forming a tree, we can explore different portions of the tree on different
-processors.
-
-?? quick history
-
-?? Anomalies
+backtracking search as forming a tree, we can explore different portions of the tree using different
+threads. We use a shared incumbent, so better solutions found by one thread can be used others
+immediately. This technique has a long history \citep{o:BaderHC05}. Of particular
+interest to us are so-called \emph{anomalies}
+\citep{DBLP:journals/cacm/LaiS84,DBLP:journals/tc/LiW86,DBLP:conf/irregular/BruinKT95}:
+because we are not performing a fixed amount of work, we should have no expectation of a linear
+speedup, and instead we could see a sublinear speedup (much less than $n$ from $n$ processors) or a
+superlinear speedup (much more than $n$ from $n$ processors). An absolute slowdown (speedup much
+less than 1) is also possible under some models.
 
@@
 
-The maximum clique algorithm underlying . Differs only by a preprocessing step.
+Thread-parallel versions of state-of-the-art maximum clique algorithms have already been produced
+\citep{DBLP:journals/algorithms/McCreeshP13,DBLP:journals/jcisd/DepolliKRTJ13,DBLP:conf/icde/XiangGA13}.
+\citet{DBLP:journals/topc/McCreeshP15} compare several of these approaches, and make an important
+observation: although work balance is a problem due to the irregularity of the search tree, often
+the interaction between search order and parallel work decomposition is the dominating factor in
+determining speedups. They explain why anomalies are in fact common in practice: many clique problem
+instances benefit immensely from having found a strong incumbent, but have solutions which are
+either unique or rare, and are hard to find. They propose a work splitting mechanism which offsets
+anomalies, guaranteeing reproducibility (two runs with the same instance on the same hardware will
+give similar runtimes), scalability (increasing the number of cores cannot make things worse), and
+no absolute slowdowns.  Additionally, this mechanism explicitly offsets the commitment to early
+branching choices, where search ordering heuristics are most likely to be inaccurate
+\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/cp/ChuSS09}.
 
-The clique-inspired connected algorithm described by \citet{} may be parallelised in exactly the
-same way.
-
-?? Note the very high number of recursive calls per second. Fitting everything in cache. Not
-compatible with framework abstractions.
-
-?? Splitting depth. Balance.
-
-?? Discuss the TOPC paper.  ?? Reproducible, anomalies, etc
-
-\subsection{Anomalies}
+We will use this mechanism for our experiments.  The clique-based maximum common subgraph algorithm
+effectively differs only in the preprocessing stage, and the clique-inspired connected algorithm
+described by \citet{DBLP:conf/cp/McCreeshNPS16} is sufficiently similar that may be parallelised in
+exactly the same way. Based upon preliminary experiments, we use a splitting depth limit of five
+rather than the original three, since maximum common subgraph instances appear to be even more
+irregular than normal clique problem instances.
 
@@
 
-?? Approach for kdown and mcsplit
+A similar approach may be used for the $k{\downarrow}$ algorithm. Although it is not quite a
+conventional branch and bound algorithm, each individual $k$ pass is a tree search, and may be
+parallelised. For each pass, we use the same work splitting mechanism as in the clique algorithm,
+starting by splitting only at the top level of search to explicitly introduce diversity, and then
+iteratively increasing the splitting depth as additional work is needed (up to a limit of five
+levels deep).
 
-?? kdown isn't quite conventional branch and bound.
+Because the $k{\downarrow}$ algorithm uses a conventional constraint programming domain store, there
+is no need to use recomputation; the state is naturally copied at each branching point.
 
-?? mcsplit uses in-place data structures
+In principle the McSplit algorithm may be parallelised in exactly the same way. However, this
+algorithm makes heavy use of an in-place, backtrackable data structure, which is not copied for
+recursive calls. In order to introduce the \emph{potential} for parallelism, we must make
+copies of the state data structure. Implemented na{\"\i}vely, this can give an order of magnitude
+slowdown to the sequential algorithm, which can be hard to recover using parallelism. To lessen the
+effects, rather than copying state for each recursive call, we copy once before the main branching
+loop, and then copy that copy in each ``helper'' thread, replaying the branching loop without making
+duplicate recursive calls.  (We are not convinced that a better approach is not possible\ldots)
 
@@
 running Ubuntu 14.04, with GCC 5.3.0 as the compiler. Each machine has a total of sixteen cores, and
-hyper-threading is enabled, so we run experiments with 32 threads except where otherwise noted.
+hyper-threading is enabled, so we run experiments with 32 threads except where otherwise noted, but
+we do not expect speedups of 32 even in the best case.
+
+?? Hundred second timeout.
 
@@
 \end{figure*}
-
-In \cref{figure:cumulative} we plot empirical cumulative distribution functions showing the number
-of instances solved over time, for both sequential (solid lines) and parallel (dotted lines)
-versions of each algorithm.  Cumulative plots are commonly used in constraint programming and
-boolean satisfiability for comparing aggregate behaviour of solvers in the presence of timeouts.
-(Sometimes the axes are reversed, in which case they are called \emph{cactus} plots.)  To read these
-plots, make a choice of timeout along the $x$-axis, which uses a log scale. The $y$ value at that
-point shows the number of instances whose runtime (individually) is at most $x$, for a particular
-algorithm. In other words, at any given $x$ value, the highest line shows which algorithm is able to
-solve the largest number of instances using a per-instance timeout of that $x$ value, bearing in
-mind that the actual sets of instances solved by each algorithm may be completely different.
-
-Each plot gives the same conclusion: if we are working with a timeout of at least $10^2$
-milliseconds, then for any problem family and any sequential algorithm, if given the option of
-switching to the corresponding parallel algorithm, then we should do so.
-
-However, although they are good at showing general trends, cumulative plots can hide interesting
-details. We therefore now take a closer look at each of the three algorithms.
-
-\subsection{A Closer Look at Clique}
 
@@
 
-In the top row of \cref{figure:cliquespeedups}, we see scatter plots comparing the sequential and
-parallel runtimes of the clique algorithm on an instance by instance basis, using a log-log plot.
-Instances which timed out using one algorithm but not the other are shown as points along the outer
-axes. Points below the $x{-}y$ diagonal line represent speedups. The colour of the points indicates
-the relative size of the solution---darker points represent instances where the solution uses most
-of the vertices of the input graphs.
+In \cref{figure:cumulative} we plot empirical cumulative distribution functions showing the number
+of instances solved over time, for both sequential (solid lines) and parallel (dotted lines)
+versions of each algorithm.  Cumulative plots are commonly used in constraint programming and
+boolean satisfiability for comparing aggregate behaviour of solvers in the presence of timeouts.
+(Sometimes the axes are reversed, in which case they are called \emph{cactus} plots.)  To read these
+plots, make a choice of timeout along the $x$-axis, which uses a log scale. The $y$ value at that
+point shows the number of instances whose runtime (individually) is at most $x$, for a particular
+algorithm. In other words, at any given $x$ value, the highest line shows which algorithm is able to
+solve the largest number of instances using a per-instance timeout of that $x$ value, bearing in
+mind that the actual sets of instances solved by each algorithm may be completely different.
 
-?? Quick summary
+Each plot gives the same conclusion: if we are working with a timeout of at least $10^2$
+milliseconds, then for any problem family and any sequential algorithm, if given the option of
+switching to the corresponding parallel algorithm, then we should do so.
 
-?? Try to turn this into a number, giving a speedup.
+However, although they are good at showing general trends, cumulative plots can hide interesting
+details. We therefore now take a closer look at each of the three algorithms.
 
-?? Best fit line. Pretend they are outliners. Can get away with it if they are rare enough, but they
-are not for many algorithms.
-
-?? Histograms. Ignore low runtimes. Also have to ignore instances solved in parallel but not
-sequential. Underestimate superlinear speedups. We do not consider this to be a particularly
-satisfactory way of measuring speedups.
-
-?? We propose a different way of . Refer back to the cumulative plots. The usual way of reading
-these plots is to pick whichever line is furthest up, and measuring the vertical difference between
-lines says how many more the parallel algorithm can solve than the sequential algorithm can with a
-particular choice of timeout. However, measuring the \emph{horizontal} distance between lines also
-conveys information. Suppose the sequential algorithm can solve $y$ instances with a selected timeout
-of $s$. By moving to the left on a cumulative plot, we can find the timeout $p$ required for the
-parallel algorithm to solve the same number of instances, bearing in mind that \emph{the two sets of
-instances could in principle have completely different members}. We define the \emph{aggregate
-speedup} to be $\nicefrac{s}{p}$; this can be expressed as a function of time (i.e.\ $s$) or of the
-number of instances solved ($y$).
+\subsection{Clique Results In Depth}
 
@@
     \caption{Aggregate speedups from 32 threads on 16 hyper-threaded cores, shown as a function of
-    sequential runtime, for each family supported by the clique algorithm.}
+    sequential runtime, for each family supported by the clique algorithm.}\label{figure:cliqueas}
 \end{figure}
-
-% \begin{figure*}[tb]
-%     \includegraphics*{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-plain-kdown-vs-kdown-par-t32.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-plain-mcsplit-vs-mcsplit-par-t32.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-33v-clique-vs-clique-par-t32.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-33v-mcsplit-vs-mcsplit-par-t32.pdf}
-% 
-%     \vspace*{1em}
-% 
-%     \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t32.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-33ved-mcsplit-vs-mcsplit-par-t32.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-sip-kdown-vs-kdown-par-t32.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-sip-mcsplit-vs-mcsplit-par-t32.pdf}
-% 
-%     \caption{Per-instance speedups. The $x$-axis is sequential performance and the $y$-axis is 32
-%     threaded performance, so points below the diagonal line represent a speedup.}
-% \end{figure*}
-% 
 \begin{figure}[p]
@@
     number of threads in subsequent plots. The final plot shows 32 threads versus a repeated run
-    also with 32 threads.}
+    also with 32 threads.}\label{figure:cliquescale}
 \end{figure}
-% 
-% \begin{figure}[tb]
-%     \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t32r-vs-clique-cilk-t32.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-cilk-t2.pdf}
-% 
-%     \vspace*{1em}
-% 
-%     \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t2-vs-clique-cilk-t4.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t4-vs-clique-cilk-t8.pdf}
-% 
-%     \vspace*{1em}
-% 
-%     \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t8-vs-clique-cilk-t16.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t16-vs-clique-cilk-t32.pdf}
-% 
-%     \caption{Scalability and reproducibility, clique algorithm, Cilk.}
-% \end{figure}
-% 
-% \begin{figure}[tb]
-%     \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t32r-vs-mcsplit-par-t32.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-plain-mcsplit-vs-mcsplit-par-t2.pdf}
-% 
-%     \vspace*{1em}
-% 
-%     \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t2-vs-mcsplit-par-t4.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t4-vs-mcsplit-par-t8.pdf}
-% 
-%     \vspace*{1em}
-% 
-%     \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t8-vs-mcsplit-par-t16.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t16-vs-mcsplit-par-t32.pdf}
-% 
-%     \caption{Scalability and reproducibility, McSplit algorithm.}
-% \end{figure}
 
-\subsection{A Closer Look at $k{\downarrow}$}
+In the top row of \cref{figure:cliquespeedups}, we see scatter plots comparing the sequential and
+parallel runtimes of the clique algorithm on an instance by instance basis, using a log-log plot.
+Instances which timed out using one algorithm but not the other are shown as points along the outer
+axes. Points below the $x{-}y$ diagonal line represent speedups. The colour of the points indicates
+the relative size of the solution---darker points represent instances where the solution uses most
+of the vertices of the input graphs. (We use these conventions for scatter plots throughout this
+paper.)
+
+Broadly speaking, the results are similar on each of the five families. For runtimes below 100
+milliseconds, overheads and the preprocessing step dominate, and we are usually only able to achieve
+a small speedup. At higher runtimes, most speedups appear to be between ten and twenty, except on
+the final family, where they are mostly between five and ten. For a few instances, the speedups are
+lower (but they are still clearly speedups), whilst in the first four families, we also see evidence
+of super-linear speedups being relatively common.
+
+However, attempting to determine a speedup by staring at a scatter plot is not particularly
+quantitative. We could attempt to find a best fit line through these points, pretending that the
+superlinear speedups are outliners. We could perhaps get away with this if outliers were rare
+enough, but in practice we are not expecting linear speedups, and for the other two algorithms, we
+will see that superlinear speedups are even more common. Alternatively, we could rig our experiments
+to remove anomalies, by priming search with a known-optimal
+solution---\citet{DBLP:journals/topc/McCreeshP15} criticise this practice in more detail.
+
+A more principled approach is given in the bottom row of \cref{figure:cliquespeedups}. For instances
+where the sequential run both succeeded and took at least 500 milliseconds, we plot the distribution
+of speedups obtained. These histograms confirm our informal observations. However, these plots are
+not particularly satisfactory: in order to calculate a speedup, we can only consider instances where
+the sequential algorithm succeeded. These plots therefore underestimate superlinear speedups.
+
+To avoid this weakness, we propose a new way of characterising speedups. Refer back to the
+cumulative plots in \cref{figure:cumulative}. The usual way of comparing two algorithms on these
+plots is to measuring the vertical difference between lines, which says how many more instances the
+parallel algorithm can solve than the sequential algorithm can with a particular choice of timeout.
+However, measuring the \emph{horizontal} distance between lines also conveys information. Suppose
+the sequential algorithm can solve $y$ instances with a selected timeout of $s$. By moving to the
+left on a cumulative plot, we can find the timeout $p$ required for the parallel algorithm to solve
+the same number of instances, bearing in mind that \emph{the two sets of instances could in
+principle have completely different members}. We therefore define the \emph{aggregate speedup} to be
+$\nicefrac{s}{p}$; this can be expressed as a function of time (i.e.\ $s$) or of the number of
+instances solved ($y$).
+
+We plot aggregate speedups as a function of time in \cref{figure:cliqueas}. For a sequential timeout
+of one hundred seconds, we get speedups of twenty to twenty eight in the unlabelled, vertex
+labelled, and both labelled, directed cases. In the unlabelled connected case, our aggregate speedup
+is thirty seven, which (when taking hyper-threading into account) is strongly super-linear.  With
+some detailed knowledge of the underlying sequential algorithm, this should perhaps not surprise us:
+for instances with a large solution, once we have found that solution, a proof of optimality is
+relatively easy. However, finding that solution can be unusually hard, particularly since the
+branching strategy for the connected constraint necessarily interferes with the tailored search
+order used by modern clique algorithms.  In contrast, for the both labelled connected case, our
+aggregate speedup is only six. A closer inspection of the results shows that the search tree is
+unusually narrow and deep for these instances, making work balance harder and contention higher.
+
+What about scalability and reproducibility? The first plot in \cref{figure:cliquescale} show the
+effects of going from sequential to threaded with two cores, and the next four plots show the
+effects of doubling the number of threads each time. These plots show that most of the superlinear
+effects occur with fairly small numbers of threads, with nearly all of the benefits of increased
+diversity in search being obtained once eight threads are used. As expected, in no case does
+increasing the number of threads make things substantially worse, although going from sixteen to
+thirty two threads does show some small slowdowns---this is due to hyper-threading.
+
+The final plot in \cref{figure:cliquescale} shows that runtimes are reproducible: running the same
+instance on the same hardware twice takes almost exactly the same amount of time. (We could not
+reproduce the three outliers in a third run---we suspect other system programs may have woken up
+and caused slight interference during the second run.)
+
+These results are hopefully comforting: they show that anomalies can be controlled, and that
+switching to a parallel algorithm is not only better, but also safe from a scientific
+reproducibility perspective.
+
+\subsection{$k{\downarrow}$ Results In Depth}
 
@@
 
-    \caption{On the top row, per-instance speedups, using the $k{\downarrow}$ algorithm. The
-    $x$-axis is sequential performance and the $y$-axis is 32 threaded performance. Below,
-    aggregate speedups for both families.}
-\end{figure}
+    \vspace*{1em}
 
-\begin{figure}[tb]
     \includegraphics*{gen-graph-scatter-plain-kdown-par-t4-vs-kdown-par-t16.pdf}
@@
 
-    \caption{Scalability and reproducibility of the $k{\downarrow}$ algorithm.}
+    \caption{On the top row, per-instance speedups, using the $k{\downarrow}$ algorithm. The
+    $x$-axis is sequential performance and the $y$-axis is 32 threaded performance. Below,
+    aggregate speedups for both families. Finally, scalability and reproducibility.}\label{figure:kdownscatters}
 \end{figure}
 
-\subsection{A Closer Look at McSplit}
+In \cref{figure:kdownscatters} we show per-instance and aggregate speedups for the $k{\downarrow}$
+algorithm. On unlabelled instances, we see a range of speedups between slightly less than one and
+ten, with an aggregate speedup of four. These results are not as good as with the clique algorithm.
+A closer look at the results suggests that memory bandwidth is to blame: although we have somewhat
+more than sixteen times as much processing power available to us, our available memory bandwidth at
+best doubles from introducing parallelism. Unlike the clique algorithm, which has very small,
+cache-friendly data structures which are modified in-place, the state for the $k{\downarrow}$
+algorithm is large and much of the runtime is spent copying data structures.
+
+For the large instances, our aggregate speedup is higher, at over ten. This has two causes: for
+larger graphs, the computational effort per recursive call increases by more than the amount the
+memory copying does, reducing the memory bandwidth problem slightly, and additionally a much larger
+number of super-linear speedups occurred with this family of instances. We could perhaps anticipate
+this latter effect: in many of these instances the maximum common subgraph covers all or nearly all
+of the smaller of the two graphs, and so once it is found, the proof of optimality is trivial.
+However, finding a witness can be difficult. We should also expect value ordering heuristics in
+these algorithms to be weak at the top of search (they are based upon degree, and many graphs do not
+have a large degree spread), and so the benefits of high-up diversity can be extremely large
+\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/cp/ChuSS09,DBLP:journals/topc/McCreeshP15}. Indeed,
+similar results were seen with a parallel version of the subgraph isomorphism algorithm upon which
+$k{\downarrow}$ is based \citep{DBLP:conf/cp/McCreeshP15}.
+
+The final row of \cref{figure:kdownscatters} shows that as with the clique algorithm, this
+parallelism is reproducible, and that runtimes do not get worse when the number of threads is
+increased. (Although not shown, we also tried to parallelise $k{\downarrow}$ using randomised
+work-stealing from Intel Cilk Plus---similar to the clique algorithm
+\citep{DBLP:journals/topc/McCreeshP15}, doing so gives generally reasonable results on average, but
+now repeat runtimes can differ by more than an order of magnitude.)
+
+\subsection{McSplit Results In Depth}
 
@@
 
+    \caption{Per-instance speedups, using the splitting algorithm. The $x$-axis is sequential
+    performance and the $y$-axis is 32 threaded performance.}\label{figure:mcsplitscatter}
+\end{figure}
+
+Finally, we look at our attempts to parallelise the McSplit algorithm. Recall that doing so required
+heavy modifications to the algorithm, introducing significant amounts of speculative copying of a
+data structure that is usually backtrackable and modified in-place. \Cref{figure:mcsplitscatter}
+shows ?? what?
+
+?? Aggregate speedups in \cref{figure:mcsplitas}. Also show scalability and reproducibility.
+
+\begin{figure}[tb]
+    \includegraphics{gen-graph-as-mcsplit.pdf}
+
     \vspace*{1em}
 
-    \includegraphics{gen-graph-as-mcsplit.pdf}
-
-    \caption{On the top three rows, per-instance speedups, using the splitting algorithm. The
-    $x$-axis is sequential performance and the $y$-axis is 32 threaded performance. On the final
-    row, aggregate speedups for both families.}
-\end{figure}
-
-\begin{figure}[tb]
     \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t4-vs-mcsplit-par-t16.pdf}
@@
 
-    \caption{Scalability and reproducibility of the splitting algorithm.}
+    \caption{On the top row row, aggregate speedups for the McSplit algorithm. Below, scalability
+    and reproducibility.}\label{figure:mcsplitas}
 \end{figure}
@@
 
-\FloatBarrier
+?? Note the very high number of recursive calls per second. Fitting everything in cache. Not
+compatible with framework abstractions. Recomputation versus state copying.
+
+?? Aggregate speedups. Use of multiple measures.
+
+?? Same algorithm, different instance families, very sensitive
 

### 2017-08-15 paper/paper.tex
Things

@@
 \title[Observations from Parallelising Three Subgraph Algorithms]{Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms}
-\titlenote{This work was supported by the Engineering and Physical Sciences
-    Research Council [grant numbers EP/K503058/1, EP/M508056, and EP/P026842/1] and the ANR project
-SoLStiCe (ANR-13-BS02-0002-01)}
 
@@
     more detail at the results, and discuss how we should be
-    measuring speedups for these kinds of algorithm. Because of the difficulty in measuring an
+    measuring speedups for this kind of algorithm. Because of the difficulty in measuring an
     average speedup when so-called anomalous (superlinear and sublinear) speedups are common, we
@@
 
-\keywords{Maximum common subgraph, Parallel branch and bound}
+\keywords{Maximum common subgraph, Maximum clique, Parallel branch and bound, Constraint
+programming, Irregular parallelism, Parallel tree search}
 
@@
 
-\begin{figure}[b]
-    \centering
-    \begin{tikzpicture}[scale=0.33]%{{{
-        \begin{scope}
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
-            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
-
-            \draw [very thick] (M1) -- (M2);
-            \draw [very thick] (M2) -- (M4);
-            \draw [very thick] (M3) -- (M5);
-            \draw [color=uofgsandstone!60] (M4) -- (M6);
-            \draw [color=uofgsandstone!60] (M5) -- (M6);
-            \draw [very thick] (M3) -- (M1);
-            \draw [color=uofgsandstone!60] (M6) -- (M7);
-            \draw [color=uofgsandstone!60] (M6) -- (M8);
-            \draw [color=uofgsandstone!60] (M6) -- (M9);
-            \draw [very thick] (M7) -- (M9);
-            \draw [very thick] (M8) -- (M9);
-        \end{scope}
-
-        \begin{scope}[xshift=5cm]
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
-            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};
-
-            \draw [very thick] (M1) -- (M2);
-            \draw [very thick] (M2) -- (M4);
-            \draw [very thick] (M3) -- (M5);
-            \draw [color=uofgsandstone!60] (M4) -- (M6);
-            \draw [color=uofgsandstone!60] (M5) -- (M6);
-            \draw [very thick] (M3) -- (M1);
-            \draw [color=uofgsandstone!60] (M6) -- (M7);
-            \draw [very thick] (M7) -- (M8);
-            \draw [very thick] (M8) -- (M9);
-        \end{scope}
-
-        \begin{scope}[xshift=12cm]
-            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
-            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
-
-            \draw [very thick] (M1) -- (M2);
-            \draw [very thick] (M2) -- (M4);
-            \draw [very thick] (M3) -- (M5);
-            \draw [very thick] (M4) -- (M6);
-            \draw [very thick] (M5) -- (M6);
-            \draw [very thick] (M3) -- (M1);
-            \draw [color=uofgsandstone!60] (M6) -- (M7);
-            \draw [very thick] (M6) -- (M8);
-            \draw [color=uofgsandstone!60] (M6) -- (M9);
-            \draw [color=uofgsandstone!60] (M7) -- (M9);
-            \draw [color=uofgsandstone!60] (M8) -- (M9);
-        \end{scope}
-
-        \begin{scope}[xshift=17cm]
-            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};
-
-            \draw [very thick] (M1) -- (M2);
-            \draw [very thick] (M2) -- (M4);
-            \draw [very thick] (M3) -- (M5);
-            \draw [very thick] (M4) -- (M6);
-            \draw [very thick] (M5) -- (M6);
-            \draw [very thick] (M3) -- (M1);
-            \draw [very thick] (M6) -- (M7);
-            \draw [draw=uofgsandstone!60] (M7) -- (M8);
-            \draw [draw=uofgsandstone!60] (M8) -- (M9);
-        \end{scope}
-\end{tikzpicture}
-\caption{A maximum common subgraph of the first two graphs has eight vertices, shaded. A maximum
-common \emph{connected} subgraph, shown on the right, has only seven vertices.}
-\label{figure:mcsexample}
-\end{figure}
+% \begin{figure}[b]
+%     \centering
+%     \begin{tikzpicture}[scale=0.33]%{{{
+%         \begin{scope}
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
+%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
+% 
+%             \draw [very thick] (M1) -- (M2);
+%             \draw [very thick] (M2) -- (M4);
+%             \draw [very thick] (M3) -- (M5);
+%             \draw [color=uofgsandstone!60] (M4) -- (M6);
+%             \draw [color=uofgsandstone!60] (M5) -- (M6);
+%             \draw [very thick] (M3) -- (M1);
+%             \draw [color=uofgsandstone!60] (M6) -- (M7);
+%             \draw [color=uofgsandstone!60] (M6) -- (M8);
+%             \draw [color=uofgsandstone!60] (M6) -- (M9);
+%             \draw [very thick] (M7) -- (M9);
+%             \draw [very thick] (M8) -- (M9);
+%         \end{scope}
+% 
+%         \begin{scope}[xshift=5cm]
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
+%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
+%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};
+% 
+%             \draw [very thick] (M1) -- (M2);
+%             \draw [very thick] (M2) -- (M4);
+%             \draw [very thick] (M3) -- (M5);
+%             \draw [color=uofgsandstone!60] (M4) -- (M6);
+%             \draw [color=uofgsandstone!60] (M5) -- (M6);
+%             \draw [very thick] (M3) -- (M1);
+%             \draw [color=uofgsandstone!60] (M6) -- (M7);
+%             \draw [very thick] (M7) -- (M8);
+%             \draw [very thick] (M8) -- (M9);
+%         \end{scope}
+% 
+%         \begin{scope}[xshift=12cm]
+%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
+%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
+%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
+%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
+% 
+%             \draw [very thick] (M1) -- (M2);
+%             \draw [very thick] (M2) -- (M4);
+%             \draw [very thick] (M3) -- (M5);
+%             \draw [very thick] (M4) -- (M6);
+%             \draw [very thick] (M5) -- (M6);
+%             \draw [very thick] (M3) -- (M1);
+%             \draw [color=uofgsandstone!60] (M6) -- (M7);
+%             \draw [very thick] (M6) -- (M8);
+%             \draw [color=uofgsandstone!60] (M6) -- (M9);
+%             \draw [color=uofgsandstone!60] (M7) -- (M9);
+%             \draw [color=uofgsandstone!60] (M8) -- (M9);
+%         \end{scope}
+% 
+%         \begin{scope}[xshift=17cm]
+%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
+%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
+%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
+%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};
+% 
+%             \draw [very thick] (M1) -- (M2);
+%             \draw [very thick] (M2) -- (M4);
+%             \draw [very thick] (M3) -- (M5);
+%             \draw [very thick] (M4) -- (M6);
+%             \draw [very thick] (M5) -- (M6);
+%             \draw [very thick] (M3) -- (M1);
+%             \draw [very thick] (M6) -- (M7);
+%             \draw [draw=uofgsandstone!60] (M7) -- (M8);
+%             \draw [draw=uofgsandstone!60] (M8) -- (M9);
+%         \end{scope}
+% \end{tikzpicture}
+% \caption{A maximum common subgraph of the first two graphs has eight vertices, shaded. A maximum
+% common \emph{connected} subgraph, shown on the right, has only seven vertices.}
+% \label{figure:mcsexample}
+% \end{figure}
 
@@
 is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent vertices, preserving
-non-adjacency. When working with labelled graphs, a subgraph isomorphism must preserve labels, and
+non-adjacency. When working with labelled graphs, a subgraph isomorphism must also preserve labels, and
 on directed graphs, it must preserve orientation. A \emph{common induced subgraph} of two graphs $G$
 and $H$ is a pair of induced subgraph isomorphisms from a pattern graph $P$, one to $G$ and one to
-$H$. A \emph{maximum common induced subgraph} is one with as many vertices as possible.
+$H$. A \emph{maximum common induced subgraph} is one with as many vertices as possible. (The
+\emph{maximum common partial subgraph} problem is non-induced, with as many edges as
+possible; this paper discusses only induced problems.)
 
@@
 \citep{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
-represent molecules, but also in applications including computer vision
+represent molecules or reactions, and also in computer vision
 \citep{DBLP:journals/jair/CookH94,DBLP:conf/gbrpr/CombierDS13}, computer-aided manufacturing
-\citep{o:LuoWSN17}, crisis management \citep{o:DelavalladeFLL16}, deanonymising datasets
-\citep{o:SharadD13}, the analysis of source code \citep{DBLP:journals/tkde/DjokoCH97} and binary
-programs \citep{DBLP:conf/icics/GaoRS08}, in
-malware detection \citep{DBLP:journals/compsec/ParkRS13}, social network analysis
+\citep{o:LuoWSN17}, crisis management \citep{o:DelavalladeFLL16}, the analysis of programs and
+malware \citep{DBLP:conf/icics/GaoRS08,DBLP:journals/compsec/ParkRS13}, social network analysis
 \citep{DBLP:journals/tkde/FangYZZ15}, graph database query explanations
@@
 \citep{DBLP:journals/tcs/Koch01,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
-As illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be determined
-from the maximum common subgraph.
 
-Although both variants are NP-hard, some progress has been made towards solving the problem in
+% As illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be determined
+% from the maximum common subgraph.
+
+Although both variants are NP-hard, recently progress has been made towards solving the problem in
 practice.  This paper looks at three recent branch and bound algorithms for maximum common
@@
 values to variables, maximising the number of variables not set to $\bot$. The state of the art for
-this technique is a dedicated implementation of a forward-checking branch and bound search over this
-model \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}.
+this technique is a dedicated (non-toolkit) implementation of a forward-checking branch and bound
+search over this
+model \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}; attempts to parallelise this
+algorithm by static decomposition have had limited success \citep{DBLP:conf/ictai/MinotNS15}.
 
@@
 A different take on the constraint programming model is presented by
-\citet{DBLP:conf/aaai/HoffmannMR17}. They approach the maximum common subgraph via the
+\citet{DBLP:conf/aaai/HoffmannMR17}. They approach maximum common subgraph via the
 subgraph isomorphism problem, asking the question ``if a pattern graph cannot be found in the
@@
 subgraph problems \citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}. This benchmark
-set can be used in a number of ways, for different variants of the problem. We use it to create five
-families of instance, as follows:
+set can be used in a number of ways, for different variants of the problem. Following other recent
+work \citep{DBLP:conf/cp/McCreeshNPS16,DBLP:conf/aaai/HoffmannMR17,o:McCreeshPT17}, we use it to
+create five families of instance, as follows:
 
@@
 
-The clique and $k{\downarrow}$ algorithms already make use of fine granularity bit-parallelism. To
+The clique and $k{\downarrow}$ algorithms already make use of fine-granularity bit-parallelism. To
 introduce more coarse-grained thread parallelism, we will parallelise the search process: viewing
@@
 because we are not performing a fixed amount of work, we should have no expectation of a linear
-speedup, and instead we could see a sublinear speedup (much less than $n$ from $n$ processors) or a
-superlinear speedup (much more than $n$ from $n$ processors). An absolute slowdown (speedup much
-less than 1) is also possible under some models.
+speedup, and instead we could see a sublinear speedup (much less than $n$ from $n$ processors, if
+speculative work turns out to be wasted) or a superlinear speedup (much more than $n$ from $n$
+processors, if a strong incumbent is found more quickly). An absolute slowdown (a speedup much less
+than 1) is also possible under some models.
 
@@
 
-Thread-parallel versions of state-of-the-art maximum clique algorithms have already been produced
-\citep{DBLP:journals/algorithms/McCreeshP13,DBLP:journals/jcisd/DepolliKRTJ13,DBLP:conf/icde/XiangGA13}.
+Thread-parallel versions of state-of-the-art maximum clique algorithms have already been produced.
 \citet{DBLP:journals/topc/McCreeshP15} compare several of these approaches, and make an important
@@
 branching choices, where search ordering heuristics are most likely to be inaccurate
-\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/cp/ChuSS09}.
+\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/cp/ChuSS09}. Under this approach, superlinear speedups
+are common.
 
@@
 exactly the same way. Based upon preliminary experiments, we use a splitting depth limit of five
-rather than the original three, since maximum common subgraph instances appear to be even more
-irregular than normal clique problem instances.
+rather than the original three, since maximum common subgraph instances appear to give even more
+irregular search trees than normal clique problem instances.
 
@@
 iteratively increasing the splitting depth as additional work is needed (up to a limit of five
-levels deep).
-
-Because the $k{\downarrow}$ algorithm uses a conventional constraint programming domain store, there
-is no need to use recomputation; the state is naturally copied at each branching point.
+levels deep).  Because the $k{\downarrow}$ algorithm uses a conventional constraint programming
+domain store, there is no need to use recomputation; the state is naturally copied at each branching
+point.
 
@@
 hyper-threading is enabled, so we run experiments with 32 threads except where otherwise noted, but
-we do not expect speedups of 32 even in the best case.
-
-?? Hundred second timeout.
+we do not expect speedups of 32 even in the best case.  We run all our experiments with a one
+hundred second timeout for each instance. All of our sequential runtimes are from optimised
+implementations which were not designed with parallelism in mind---that is, they are genuine
+improvements over the state of the art.
 
@@
 versions of each algorithm.  Cumulative plots are commonly used in constraint programming and
-boolean satisfiability for comparing aggregate behaviour of solvers in the presence of timeouts.
-(Sometimes the axes are reversed, in which case they are called \emph{cactus} plots.)  To read these
-plots, make a choice of timeout along the $x$-axis, which uses a log scale. The $y$ value at that
-point shows the number of instances whose runtime (individually) is at most $x$, for a particular
-algorithm. In other words, at any given $x$ value, the highest line shows which algorithm is able to
-solve the largest number of instances using a per-instance timeout of that $x$ value, bearing in
-mind that the actual sets of instances solved by each algorithm may be completely different.
+boolean satisfiability experiments for comparing the aggregate behaviour of solvers in the presence
+of timeouts.  (Sometimes the axes are reversed, in which case they are called \emph{cactus} plots.)
+To read these plots, make a choice of timeout along the $x$-axis, which uses a log scale. The $y$
+value at that point shows the number of instances whose runtime (individually) is at most $x$, for a
+particular algorithm. In other words, at any given $x$ value, the highest line shows which algorithm
+is able to solve the largest number of instances using a per-instance timeout of that $x$ value,
+bearing in mind that the actual sets of instances solved by each algorithm may be completely
+different.
 
@@
 Instances which timed out using one algorithm but not the other are shown as points along the outer
-axes. Points below the $x{-}y$ diagonal line represent speedups. The colour of the points indicates
+borders. Points below the $x{-}y$ diagonal line represent speedups. The colour of the points indicates
 the relative size of the solution---darker points represent instances where the solution uses most
@@
 a small speedup. At higher runtimes, most speedups appear to be between ten and twenty, except on
-the final family, where they are mostly between five and ten. For a few instances, the speedups are
-lower (but they are still clearly speedups), whilst in the first four families, we also see evidence
-of super-linear speedups being relatively common.
+the final family of both labelled connected instances, where they are mostly between five and ten.
+For a few instances, the speedups are lower (but they are still clearly speedups), whilst in the
+first four families, we also see evidence of superlinear speedups being relatively common.
 
 However, attempting to determine a speedup by staring at a scatter plot is not particularly
-quantitative. We could attempt to find a best fit line through these points, pretending that the
-superlinear speedups are outliners. We could perhaps get away with this if outliers were rare
-enough, but in practice we are not expecting linear speedups, and for the other two algorithms, we
-will see that superlinear speedups are even more common. Alternatively, we could rig our experiments
-to remove anomalies, by priming search with a known-optimal
-solution---\citet{DBLP:journals/topc/McCreeshP15} criticise this practice in more detail.
+quantitative. We \emph{could} attempt to find a best fit line through these points, pretending that
+the superlinear speedups are outliners. We might perhaps get away with this if outliers were rare
+enough, but in practice we are not expecting linear speedups (and for the other two algorithms, we
+will see that superlinear speedups are even more common). Alternatively, we could rig our
+experiments to remove anomalies, by priming search with a known-optimal
+solution---\citet{DBLP:journals/topc/McCreeshP15} criticise this disappointingly common practice in
+more detail.
 
@@
 of speedups obtained. These histograms confirm our informal observations. However, these plots are
-not particularly satisfactory: in order to calculate a speedup, we can only consider instances where
-the sequential algorithm succeeded. These plots therefore underestimate superlinear speedups.
+still not especially satisfactory: in order to calculate a speedup, we can only consider instances
+where the sequential algorithm succeeded. These plots therefore underestimate superlinear speedups.
 
@@
 cumulative plots in \cref{figure:cumulative}. The usual way of comparing two algorithms on these
-plots is to measuring the vertical difference between lines, which says how many more instances the
-parallel algorithm can solve than the sequential algorithm can with a particular choice of timeout.
-However, measuring the \emph{horizontal} distance between lines also conveys information. Suppose
-the sequential algorithm can solve $y$ instances with a selected timeout of $s$. By moving to the
-left on a cumulative plot, we can find the timeout $p$ required for the parallel algorithm to solve
-the same number of instances, bearing in mind that \emph{the two sets of instances could in
-principle have completely different members}. We therefore define the \emph{aggregate speedup} to be
-$\nicefrac{s}{p}$; this can be expressed as a function of time (i.e.\ $s$) or of the number of
-instances solved ($y$).
+plots is by measuring the vertical difference between lines, which would tell us how many more
+instances the parallel algorithm can solve than the sequential algorithm can with a particular
+choice of timeout.  However, measuring the \emph{horizontal} distance between lines also conveys
+information. Suppose the sequential algorithm can solve $y$ instances with a selected timeout of
+$s$. By moving to the left on a cumulative plot, we can find the timeout $p$ required for the
+parallel algorithm to solve the same number of instances, bearing in mind that \emph{the two sets of
+instances could in principle have completely different members}. We therefore define the
+\emph{aggregate speedup} to be $\nicefrac{s}{p}$; this can be expressed as a function of time (i.e.\
+$s$) or of the number of instances solved ($y$).
 
@@
 labelled, and both labelled, directed cases. In the unlabelled connected case, our aggregate speedup
-is thirty seven, which (when taking hyper-threading into account) is strongly super-linear.  With
+is thirty seven, which (when taking hyper-threading into account) is strongly superlinear.  With
 some detailed knowledge of the underlying sequential algorithm, this should perhaps not surprise us:
@@
 
-These results are hopefully comforting: they show that anomalies can be controlled, and that
+These results are comforting: they show that anomalies can be controlled, and that
 switching to a parallel algorithm is not only better, but also safe from a scientific
@@
 
-    \vspace*{1em}
+    \vspace*{0.2em}
 
@@
 ten, with an aggregate speedup of four. These results are not as good as with the clique algorithm.
-A closer look at the results suggests that memory bandwidth is to blame: although we have somewhat
-more than sixteen times as much processing power available to us, our available memory bandwidth at
-best doubles from introducing parallelism. Unlike the clique algorithm, which has very small,
+A closer look suggests that memory bandwidth is to blame: although we have somewhat more than
+sixteen times as much processing power available to us, our available memory bandwidth at best
+doubles from introducing parallelism. Unlike the clique algorithm, which has very small,
 cache-friendly data structures which are modified in-place, the state for the $k{\downarrow}$
@@
 memory copying does, reducing the memory bandwidth problem slightly, and additionally a much larger
-number of super-linear speedups occurred with this family of instances. We could perhaps anticipate
+number of superlinear speedups occurred with this family of instances. We could perhaps anticipate
 this latter effect: in many of these instances the maximum common subgraph covers all or nearly all
 of the smaller of the two graphs, and so once it is found, the proof of optimality is trivial.
-However, finding a witness can be difficult. We should also expect value ordering heuristics in
+However, finding a witness can be difficult. We should also expect value-ordering heuristics in
 these algorithms to be weak at the top of search (they are based upon degree, and many graphs do not
@@
 
-\begin{figure}[tb]
+\begin{figure}[p]
     \includegraphics*{gen-graph-scatter-plain-mcsplit-vs-mcsplit-par-t32.pdf}
@@
 
-    \caption{Per-instance speedups, using the splitting algorithm. The $x$-axis is sequential
-    performance and the $y$-axis is 32 threaded performance.}\label{figure:mcsplitscatter}
-\end{figure}
+    \vspace*{0.6em}
 
-Finally, we look at our attempts to parallelise the McSplit algorithm. Recall that doing so required
-heavy modifications to the algorithm, introducing significant amounts of speculative copying of a
-data structure that is usually backtrackable and modified in-place. \Cref{figure:mcsplitscatter}
-shows ?? what?
-
-?? Aggregate speedups in \cref{figure:mcsplitas}. Also show scalability and reproducibility.
-
-\begin{figure}[tb]
     \includegraphics{gen-graph-as-mcsplit.pdf}
 
-    \vspace*{1em}
+    \vspace*{0.2em}
 
@@
 
-    \caption{On the top row row, aggregate speedups for the McSplit algorithm. Below, scalability
-    and reproducibility.}\label{figure:mcsplitas}
+    \caption{On the first three rows, per-instance speedups, using McSplit. On the
+    fourth row, aggregate speedups, and on the final row, scalability and
+    reproducibility.}\label{figure:mcsplitscatter}
 \end{figure}
 
-% \subsection{Cilk}
-%
-% \begin{figure}[tb]
-%     \includegraphics*{gen-graph-scatter-33ved-clique-par-t32r-vs-clique-par-t32.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-33ved-clique-cilk-t32r-vs-clique-cilk-t32.pdf}
-%
-%     \vspace*{1em}
-%
-%     \includegraphics*{gen-graph-scatter-sip-kdown-par-t32r-vs-kdown-par-t32.pdf}
-%     \hfill
-%     \includegraphics*{gen-graph-scatter-sip-kdown-cilk-t32r-vs-kdown-cilk-t32.pdf}
-%
-%     \caption{Reproducibility using Cilk.}
-% \end{figure}
+Finally, we look at our attempts to parallelise the McSplit algorithm. Recall that doing so required
+heavy modifications to the algorithm, introducing significant amounts of speculative copying of a
+data structure that is usually backtrackable and modified in-place. For unlabelled, unlabelled
+connected, and large instances, \cref{figure:mcsplitscatter} shows a particularly high proportion of
+strongly superlinear speedups. This is because the McSplit algorithm is focussed upon
+exploring the search space very quickly, and its branching heuristics do not have the advantage of
+the domain filtering performed by $k{\downarrow}$, or the rich inter-domain knowledge coming from
+the combination of the association graph encoding and the colour ordering used by clique algorithms.
+Thus making a correct value-ordering choice at the top of search is harder for McSplit than for
+other algorithms, and so increased diversity can be particularly beneficial.
+
+For the large instances, we see evidence of work balance problems. McSplit's use of a ``smallest
+domain first'' variable-ordering heuristic, combined with the presence of $\bot$ in domains, tends
+to produce narrow (nearly binary) and deep search trees. These balance problems are even more
+evident in the labelled cases (where following a guessed assignment, many domains are left with only
+two values), and often lead to little to no speedup being obtained.
+
+The scatter plots also show occasional absolute slowdowns, sometimes by up to an order of magnitude.
+These are due to the changes which had to be made to the sequential algorithm (and because we are
+benchmarking against the sequential algorithm, not a parallel algorithm with one thread), rather
+than search order effects. In cases where parallelism cannot be exploited, the cost of speculatively
+copying domains at each level of search can dominate the runtimes. Because of this, fixing work
+balance problems by increasing the splitting depth typically makes matters much worse, not better.
+
+What about scalability and reproducibility? The final row of \cref{figure:mcsplitscatter} presents a
+less ideal picture than for the previous two algorithms---again, this is due to speculative
+overheads that fail to pay off, rather than being anomalies in the classical sense.
 
@@
 
-?? There is more to be done.
+We have parallelised three state-of-the-art maximum common (connected) subgraph algorithms with a
+reasonable degree of success by using dynamic work-splitting. The parallel algorithms are not just
+better in aggregate, but also preserve the desirable reproducibility properties of sequential
+algorithms. A large part of our success was down to using parallelism to explicitly introduce
+diversity into the search process, offsetting weak early value-ordering branching choices.
 
-?? Note the very high number of recursive calls per second. Fitting everything in cache. Not
-compatible with framework abstractions. Recomputation versus state copying.
+Despite having a branch and bound flavour, all three of the sequential algorithms had their own
+difficulties and performance characteristics which prevented them from cleanly fitting into common
+abstraction frameworks.
 
-?? Aggregate speedups. Use of multiple measures.
+There is room for improvement, particularly with respect to work balance. However, improvements to
+work balance must not come at the expense of the search order properties, nor at the cost of
+increased overheads. We would also like to see distributed versions of these algorithms.
 
-?? Same algorithm, different instance families, very sensitive
+We introduced the idea of \emph{aggregate speedups}, to deal with measuring a speedup in the
+presence of anomalies. These figures informed part of our analysis, but our results highlight the
+importance of viewing results in multiple ways, and in using large families of instances with
+different characteristics when evaluating parallel search algorithms.
+
+\appendix
+\begin{acks}
+    This work was supported by the Engineering and Physical Sciences Research Council [grant
+numbers EP/K503058/1, EP/M508056, and EP/P026842/1] and the ANR project SoLStiCe
+(ANR-13-BS02-0002-01).
+\end{acks}
 

### 2017-08-15 paper/paper.tex
Last bits

@@
 
-% \begin{figure}[b]
-%     \centering
-%     \begin{tikzpicture}[scale=0.33]%{{{
-%         \begin{scope}
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
-%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
-% 
-%             \draw [very thick] (M1) -- (M2);
-%             \draw [very thick] (M2) -- (M4);
-%             \draw [very thick] (M3) -- (M5);
-%             \draw [color=uofgsandstone!60] (M4) -- (M6);
-%             \draw [color=uofgsandstone!60] (M5) -- (M6);
-%             \draw [very thick] (M3) -- (M1);
-%             \draw [color=uofgsandstone!60] (M6) -- (M7);
-%             \draw [color=uofgsandstone!60] (M6) -- (M8);
-%             \draw [color=uofgsandstone!60] (M6) -- (M9);
-%             \draw [very thick] (M7) -- (M9);
-%             \draw [very thick] (M8) -- (M9);
-%         \end{scope}
-% 
-%         \begin{scope}[xshift=5cm]
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
-%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
-%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};
-% 
-%             \draw [very thick] (M1) -- (M2);
-%             \draw [very thick] (M2) -- (M4);
-%             \draw [very thick] (M3) -- (M5);
-%             \draw [color=uofgsandstone!60] (M4) -- (M6);
-%             \draw [color=uofgsandstone!60] (M5) -- (M6);
-%             \draw [very thick] (M3) -- (M1);
-%             \draw [color=uofgsandstone!60] (M6) -- (M7);
-%             \draw [very thick] (M7) -- (M8);
-%             \draw [very thick] (M8) -- (M9);
-%         \end{scope}
-% 
-%         \begin{scope}[xshift=12cm]
-%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
-%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
-%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
-%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
-% 
-%             \draw [very thick] (M1) -- (M2);
-%             \draw [very thick] (M2) -- (M4);
-%             \draw [very thick] (M3) -- (M5);
-%             \draw [very thick] (M4) -- (M6);
-%             \draw [very thick] (M5) -- (M6);
-%             \draw [very thick] (M3) -- (M1);
-%             \draw [color=uofgsandstone!60] (M6) -- (M7);
-%             \draw [very thick] (M6) -- (M8);
-%             \draw [color=uofgsandstone!60] (M6) -- (M9);
-%             \draw [color=uofgsandstone!60] (M7) -- (M9);
-%             \draw [color=uofgsandstone!60] (M8) -- (M9);
-%         \end{scope}
-% 
-%         \begin{scope}[xshift=17cm]
-%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
-%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
-%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
-%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};
-% 
-%             \draw [very thick] (M1) -- (M2);
-%             \draw [very thick] (M2) -- (M4);
-%             \draw [very thick] (M3) -- (M5);
-%             \draw [very thick] (M4) -- (M6);
-%             \draw [very thick] (M5) -- (M6);
-%             \draw [very thick] (M3) -- (M1);
-%             \draw [very thick] (M6) -- (M7);
-%             \draw [draw=uofgsandstone!60] (M7) -- (M8);
-%             \draw [draw=uofgsandstone!60] (M8) -- (M9);
-%         \end{scope}
-% \end{tikzpicture}
-% \caption{A maximum common subgraph of the first two graphs has eight vertices, shaded. A maximum
-% common \emph{connected} subgraph, shown on the right, has only seven vertices.}
-% \label{figure:mcsexample}
-% \end{figure}
+\begin{figure}[b]
+    \centering
+    \begin{tikzpicture}[scale=0.25]%{{{
+        \begin{scope}
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M1) at (90:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M2) at (150:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M3) at (30:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M4) at (210:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M5) at (330:1.5) {\phantom{0}};
+            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M6) at (270:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
+
+            \draw [very thick] (M1) -- (M2);
+            \draw [very thick] (M2) -- (M4);
+            \draw [very thick] (M3) -- (M5);
+            \draw [color=uofgsandstone!60] (M4) -- (M6);
+            \draw [color=uofgsandstone!60] (M5) -- (M6);
+            \draw [very thick] (M3) -- (M1);
+            \draw [color=uofgsandstone!60] (M6) -- (M7);
+            \draw [color=uofgsandstone!60] (M6) -- (M8);
+            \draw [color=uofgsandstone!60] (M6) -- (M9);
+            \draw [very thick] (M7) -- (M9);
+            \draw [very thick] (M8) -- (M9);
+        \end{scope}
+
+        \begin{scope}[xshift=5cm]
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M1) at (90:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M2) at (150:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M3) at (30:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M4) at (210:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M5) at (330:1.5) {\phantom{0}};
+            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M6) at (270:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};
+
+            \draw [very thick] (M1) -- (M2);
+            \draw [very thick] (M2) -- (M4);
+            \draw [very thick] (M3) -- (M5);
+            \draw [color=uofgsandstone!60] (M4) -- (M6);
+            \draw [color=uofgsandstone!60] (M5) -- (M6);
+            \draw [very thick] (M3) -- (M1);
+            \draw [color=uofgsandstone!60] (M6) -- (M7);
+            \draw [very thick] (M7) -- (M8);
+            \draw [very thick] (M8) -- (M9);
+        \end{scope}
+
+        \begin{scope}[xshift=12cm]
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M1) at (90:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M2) at (150:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M3) at (30:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M4) at (210:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M5) at (330:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M6) at (270:1.5) {\phantom{0}};
+            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
+
+            \draw [very thick] (M1) -- (M2);
+            \draw [very thick] (M2) -- (M4);
+            \draw [very thick] (M3) -- (M5);
+            \draw [very thick] (M4) -- (M6);
+            \draw [very thick] (M5) -- (M6);
+            \draw [very thick] (M3) -- (M1);
+            \draw [color=uofgsandstone!60] (M6) -- (M7);
+            \draw [very thick] (M6) -- (M8);
+            \draw [color=uofgsandstone!60] (M6) -- (M9);
+            \draw [color=uofgsandstone!60] (M7) -- (M9);
+            \draw [color=uofgsandstone!60] (M8) -- (M9);
+        \end{scope}
+
+        \begin{scope}[xshift=17cm]
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M1) at (90:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M2) at (150:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M3) at (30:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M4) at (210:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M5) at (330:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M6) at (270:1.5) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};
+
+            \draw [very thick] (M1) -- (M2);
+            \draw [very thick] (M2) -- (M4);
+            \draw [very thick] (M3) -- (M5);
+            \draw [very thick] (M4) -- (M6);
+            \draw [very thick] (M5) -- (M6);
+            \draw [very thick] (M3) -- (M1);
+            \draw [very thick] (M6) -- (M7);
+            \draw [draw=uofgsandstone!60] (M7) -- (M8);
+            \draw [draw=uofgsandstone!60] (M8) -- (M9);
+        \end{scope}
+\end{tikzpicture}
+\caption{A maximum common subgraph of the first two graphs has eight vertices, shaded. A maximum
+common \emph{connected} subgraph, shown on the right, has only seven vertices.}
+\label{figure:mcsexample}
+\end{figure}
 
@@
 \citep{o:LuoWSN17}, crisis management \citep{o:DelavalladeFLL16}, the analysis of programs and
-malware \citep{DBLP:conf/icics/GaoRS08,DBLP:journals/compsec/ParkRS13}, social network analysis
-\citep{DBLP:journals/tkde/FangYZZ15}, graph database query explanations
-\citep{DBLP:journals/jcss/VasilyevaTBL16}, and in character recognition problems
-\citep{DBLP:journals/pr/LuRS91}.  A common variant of the problem requires a largest
+malware \citep{DBLP:conf/icics/GaoRS08,DBLP:journals/compsec/ParkRS13}, and social network analysis
+\citep{DBLP:journals/tkde/FangYZZ15}. A common variant of the problem requires a largest
 \emph{connected} subgraph
-\citep{DBLP:journals/tcs/Koch01,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
-
-% As illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be determined
-% from the maximum common subgraph.
+\citep{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
+As illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be determined
+from the maximum common subgraph.
 

### 2017-08-15 paper/paper.tex
Patrick's comments

@@
 backtracking search as forming a tree, we can explore different portions of the tree using different
-threads. We use a shared incumbent, so better solutions found by one thread can be used others
+threads. We use a shared incumbent, so better solutions found by one thread can be used by others
 immediately. This technique has a long history \citep{o:BaderHC05}. Of particular
@@
 effectively differs only in the preprocessing stage, and the clique-inspired connected algorithm
-described by \citet{DBLP:conf/cp/McCreeshNPS16} is sufficiently similar that may be parallelised in
+described by \citet{DBLP:conf/cp/McCreeshNPS16} is sufficiently similar that it may be parallelised in
 exactly the same way. Based upon preliminary experiments, we use a splitting depth limit of five
@@
 loop, and then copy that copy in each ``helper'' thread, replaying the branching loop without making
-duplicate recursive calls.  (We are not convinced that a better approach is not possible\ldots)
+duplicate recursive calls.  (We believe a better approach using partial recomputation may be
+possible, and intend to investigate this further in the future.)
 

### 2017-08-15 paper/paper.tex
Ruth's moanings

@@
 \citep{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
-As illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be determined
+As illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be inferred
 from the maximum common subgraph.
@@
 
-A clique in a graph is a subgraph where every vertex is adjacent to every other. There is a
+A \emph{clique} in a graph is a subgraph where every vertex is adjacent to every other. There is a
 well-known reduction from the maximum common subgraph problem to the problem of finding a maximum
@@
 
-The maximum common subgraph problem may be reformulated as a constraint optimisation problem, as
-follows.  Observe that an equivalent definition of a common subgraph of graphs $G$ and $H$ is an
-injective \emph{partial} mapping from $G$ to $H$ which preserves both adjacency and non-adjacency.
-Hence we pick whichever input graph has fewer vertices, and call it the \emph{pattern}; the other
-graph is called the \emph{target}. The model then follows from this new definition: for each vertex
-in the pattern, we create a variable, whose domain ranges over each vertex in the target graph, plus
-an additional value $\bot$ representing an unmapped vertex. We then have three sets of constraints.
-The first set says that for each pair of adjacent vertices in the pattern (that is, for each edge in
-the pattern), if neither of these vertices are mapped to $\bot$ then these vertices must be mapped
-to an adjacent pair of target vertices. The second set is similar, but looks at non-adjacent pairs
-(or non-edges).  Finally, the third set ensures injectivity, by enforcing that the variables must be
-all different except when using $\bot$. This final set of constraints may either be implemented
-using binary constraints between all pairs of variables, or a special global ``all different except
-$\bot$'' propagator \citep{DBLP:conf/cp/PetitRB01}. The objective is simply to find an assignment of
-values to variables, maximising the number of variables not set to $\bot$. The state of the art for
-this technique is a dedicated (non-toolkit) implementation of a forward-checking branch and bound
-search over this
-model \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}; attempts to parallelise this
-algorithm by static decomposition have had limited success \citep{DBLP:conf/ictai/MinotNS15}.
+The maximum common induced subgraph problem may be reformulated as a constraint optimisation
+problem, as follows.  Observe that an equivalent definition of a common subgraph of graphs $G$ and
+$H$ is an injective \emph{partial} mapping from $G$ to $H$ which preserves both adjacency and
+non-adjacency.  Hence we pick whichever input graph has fewer vertices, and call it the
+\emph{pattern}; the other graph is called the \emph{target}. The model then follows from this new
+definition: for each vertex in the pattern, we create a variable, whose domain ranges over each
+vertex in the target graph, plus an additional value $\bot$ representing an unmapped vertex. We then
+have three sets of constraints.  The first set says that for each pair of adjacent vertices in the
+pattern (that is, for each edge in the pattern), if neither of these vertices are mapped to $\bot$
+then these vertices must be mapped to an adjacent pair of target vertices. The second set is
+similar, but looks at non-adjacent pairs (or non-edges).  Finally, the third set ensures
+injectivity, by enforcing that the variables must be all different except when using $\bot$. This
+final set of constraints may either be implemented using binary constraints between all pairs of
+variables, or a special global ``all different except $\bot$'' propagator
+\citep{DBLP:conf/cp/PetitRB01}. The objective is simply to find an assignment of values to
+variables, maximising the number of variables not set to $\bot$. The state of the art for this
+technique is a dedicated (non-toolkit) implementation of a forward-checking branch and bound search
+over this model \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}; attempts to parallelise
+this algorithm by static decomposition have had limited success \citep{DBLP:conf/ictai/MinotNS15}.
 
@@
 
-\subsection{Domain Splitting}
+\subsection{Domain Splitting (McSplit)}
 
@@
 of timeouts.  (Sometimes the axes are reversed, in which case they are called \emph{cactus} plots.)
-To read these plots, make a choice of timeout along the $x$-axis, which uses a log scale. The $y$
+To read these plots, make a choice of timeout along the $x$-axis (which uses a log scale). The $y$
 value at that point shows the number of instances whose runtime (individually) is at most $x$, for a

### 2017-08-15 paper/paper.tex
More tweaking

@@
     Nonetheless, our results show that each algorithm can be parallelised successfully, with the
-    multi-core algorithms we create being clearly better than the sequential ones. We then look in
+    threaded algorithms we create being clearly better than the sequential ones. We then look in
     more detail at the results, and discuss how we should be
-    measuring speedups for this kind of algorithm. Because of the difficulty in measuring an
+    measuring speedups for this kind of algorithm. Because of the difficulty in quantifying an
     average speedup when so-called anomalous (superlinear and sublinear) speedups are common, we
@@
 
+Finding a maximum common subgraph is the key step in measuring the similarity or difference between
+two graphs \citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}: to determine
+the difference between two graphs, we find what they have in common, and then take everything left
+over. Because of this, maximum common subgraph problems frequently arise in biology and chemistry
+\citep{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
+represent molecules or reactions, and also in computer vision
+\citep{DBLP:journals/jair/CookH94,DBLP:conf/gbrpr/CombierDS13}, computer-aided manufacturing
+\citep{o:LuoWSN17}, the analysis of programs and malware
+\citep{DBLP:conf/icics/GaoRS08,DBLP:journals/compsec/ParkRS13}, crisis management
+\citep{o:DelavalladeFLL16}, and social network analysis \citep{DBLP:journals/tkde/FangYZZ15}.
+
 A \emph{subgraph isomorphism} is an injective mapping from a \emph{pattern} graph to a \emph{target}
@@
 
-Finding a maximum common subgraph is the key step in measuring the similarity or difference between
-two graphs \citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}: to determine
-the difference between two graphs, we find what they have in common, and then take everything left
-over. Because of this, maximum common subgraph problems frequently arise in biology and chemistry
-\citep{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
-represent molecules or reactions, and also in computer vision
-\citep{DBLP:journals/jair/CookH94,DBLP:conf/gbrpr/CombierDS13}, computer-aided manufacturing
-\citep{o:LuoWSN17}, crisis management \citep{o:DelavalladeFLL16}, the analysis of programs and
-malware \citep{DBLP:conf/icics/GaoRS08,DBLP:journals/compsec/ParkRS13}, and social network analysis
-\citep{DBLP:journals/tkde/FangYZZ15}. A common variant of the problem requires a largest
-\emph{connected} subgraph
-\citep{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
-As illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be inferred
+ A common variant of the problem requires a largest \emph{connected} subgraph
+ \citep{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.  As
+ illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be inferred
  from the maximum common subgraph.
@@
 clique in an \emph{association graph}
-\citep{o:Levi73,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/cp/McCreeshNPS16}, which, when combined with
-a modern maximum clique solver \citep{DBLP:journals/ol/SegundoMRH13}, is the current best approach for
-solving the problem on labelled graphs \citep{DBLP:conf/cp/McCreeshNPS16}. A modified clique-like
-algorithm can also be used to solve the maximum common connected subgraph problem, by ensuring
-connectedness during search \citep{DBLP:conf/cp/McCreeshNPS16}; again, this is the best known way of
-solving the problem on labelled graphs.
-
-A major downside of the maximum clique approach is that the association graph encoding is extremely
-memory-intensive, limiting its practical use to pairs of graphs with no more than a few hundred
-vertices.
+\citep{o:Levi73,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/cp/McCreeshNPS16}, which, when combined
+with a modern maximum clique solver \citep{DBLP:journals/ol/SegundoMRH13}, is the current best
+approach for solving the problem on labelled graphs \citep{DBLP:conf/cp/McCreeshNPS16}. A modified
+clique-like algorithm can also be used to solve the maximum common connected subgraph problem, by
+ensuring connectedness during search \citep{DBLP:conf/cp/McCreeshNPS16}; again, this is the best
+known way of solving the problem on labelled graphs. However, the association graph encoding is
+extremely memory-intensive, limiting its practical use to pairs of graphs with no more than a few
+hundred vertices.
 
@@
 However, although they are good at showing general trends, cumulative plots can hide interesting
-details. We therefore now take a closer look at each of the three algorithms.
+details. We therefore now take a closer look at each of the three algorithms in turn.
 
@@
 We have parallelised three state-of-the-art maximum common (connected) subgraph algorithms with a
-reasonable degree of success by using dynamic work-splitting. The parallel algorithms are not just
-better in aggregate, but also preserve the desirable reproducibility properties of sequential
-algorithms. A large part of our success was down to using parallelism to explicitly introduce
-diversity into the search process, offsetting weak early value-ordering branching choices.
-
-Despite having a branch and bound flavour, all three of the sequential algorithms had their own
-difficulties and performance characteristics which prevented them from cleanly fitting into common
-abstraction frameworks.
+reasonable degree of success by using dynamic work-splitting. Despite having a branch and bound
+flavour, all three sequential algorithms had their own difficulties and performance characteristics
+which prevented them from cleanly fitting into common abstraction frameworks. Nonetheless, our
+results show that the parallel algorithms are not just better in aggregate, but also preserve the
+desirable reproducibility properties of sequential algorithms. A large part of our success was down
+to using parallelism to explicitly introduce diversity into the search process, offsetting weak
+early value-ordering branching choices.
 
@@
 work balance must not come at the expense of the search order properties, nor at the cost of
-increased overheads. We would also like to see distributed versions of these algorithms.
+increased overheads. We would also like to see distributed versions of these algorithms, to make use
+of clusters of multi-core machines.
 
-We introduced the idea of \emph{aggregate speedups}, to deal with measuring a speedup in the
-presence of anomalies. These figures informed part of our analysis, but our results highlight the
-importance of viewing results in multiple ways, and in using large families of instances with
-different characteristics when evaluating parallel search algorithms.
+Aside from the parallel algorithms, we introduced the idea of \emph{aggregate speedups}, to deal
+with measuring a speedup in the presence of anomalies. These figures informed part of our analysis,
+but our results highlight the importance of viewing results in multiple ways, and in using large
+families of instances with different characteristics when evaluating parallel search algorithms.
 

### 2017-08-15 paper/paper.tex
To not be or to be not

@@
 the subgraph isomorphism problem first for $k=0$ (asking whether the whole pattern graph can be found in the
-target). Should that be not satisfiable, it tries to solve the problem for $k=1$ (one vertex cannot
+target). Should that not be satisfiable, it tries to solve the problem for $k=1$ (one vertex cannot
 be matched), and should that also not be satisfiable, it iteratively increases $k$ until the result

### 2017-08-15 paper/paper.tex
Clean up a bit

@@
     threaded algorithms we create being clearly better than the sequential ones. We then look in
-    more detail at the results, and discuss how we should be
-    measuring speedups for this kind of algorithm. Because of the difficulty in quantifying an
-    average speedup when so-called anomalous (superlinear and sublinear) speedups are common, we
-    propose a new measure called \emph{aggregate speedup}.
+    more detail at the results, and discuss speedups should be measured for this kind of algorithm.
+    Because of the difficulty in quantifying an average speedup when so-called anomalous
+    speedups (superlinear and sublinear) are common, we propose a new measure called \emph{aggregate
+    speedup}.
 \end{abstract}
@@
 \keywords{Maximum common subgraph, Maximum clique, Parallel branch and bound, Constraint
-programming, Irregular parallelism, Parallel tree search}
+programming, Irregular parallelism, Parallel tree-search}
 
@@
 A \emph{subgraph isomorphism} is an injective mapping from a \emph{pattern} graph to a \emph{target}
-graph which maps adjacent vertices to adjacent vertices---that is, it \emph{preserves adjacency}. It
-is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent vertices, preserving
+graph which maps adjacent vertices to adjacent vertices---that is, it \emph{preserves adjacency}.
+The isomorphism is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent vertices, preserving
 non-adjacency. When working with labelled graphs, a subgraph isomorphism must also preserve labels, and
@@
 
-Although both variants are NP-hard, recently progress has been made towards solving the problem in
-practice.  This paper looks at three recent branch and bound algorithms for maximum common
-(connected) induced subgraph problems, each of which is the state of the art for certain classes of
-instance. We discuss our experiences in adding parallel tree-search to these three algorithms. In
-each case, our results show that the parallel version of the algorithm is clearly better than the
-sequential version, although a closer look at the results shows many nuances.
+Although both the connected and non-connected variants are NP-hard, recently progress has been made
+towards solving the problem in practice.  This paper looks at three recent branch and bound
+algorithms for maximum common (connected) induced subgraph problems, each of which is the state of
+the art for certain classes of instance. We discuss our experiences in adding parallel tree-search
+to these three algorithms. In each case, our results show that the parallel version of the algorithm
+is clearly better than the sequential version, although a closer look at the results shows many
+nuances.
 
@@
 Following \citet{DBLP:conf/aaai/HoffmannMR17}, we also work with the 5,725 \textbf{Large} instances
-originally used for studying portfolios of subgraph isomorphism algorithms
+originally introduced for studying portfolios of subgraph isomorphism algorithms
 \citep{DBLP:conf/lion/KotthoffMS16}. These graphs are unlabelled and undirected, and can include up
@@
 branching choices, where search ordering heuristics are most likely to be inaccurate
-\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/cp/ChuSS09}. Under this approach, superlinear speedups
-are common.
+\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/cp/ChuSS09}, making superlinear speedups
+common.
 
@@
 A similar approach may be used for the $k{\downarrow}$ algorithm. Although it is not quite a
-conventional branch and bound algorithm, each individual $k$ pass is a tree search, and may be
+conventional branch and bound algorithm, each individual $k$ pass is a tree-search, and may be
 parallelised. For each pass, we use the same work splitting mechanism as in the clique algorithm,
@@
 hundred second timeout for each instance. All of our sequential runtimes are from optimised
-implementations which were not designed with parallelism in mind---that is, they are genuine
-improvements over the state of the art.
+implementations which were not designed with parallelism in mind---that is, speedups from
+parallelism are genuine improvements over the state of the art.
 
@@
 
-However, although they are good at showing general trends, cumulative plots can hide interesting
-details. We therefore now take a closer look at each of the three algorithms in turn.
+Although they are good at showing general trends, cumulative plots can hide interesting details. We
+therefore now take a closer look at each of the three algorithms in turn.
 
@@
 parallel algorithm to solve the same number of instances, bearing in mind that \emph{the two sets of
-instances could in principle have completely different members}. We therefore define the
-\emph{aggregate speedup} to be $\nicefrac{s}{p}$; this can be expressed as a function of time (i.e.\
-$s$) or of the number of instances solved ($y$).
+instances could have completely different members}. We define the \emph{aggregate speedup} to be
+$\nicefrac{s}{p}$; this can be expressed as a function of time (i.e.\ $s$) or of the number of
+instances solved ($y$).
 
@@
 A closer look suggests that memory bandwidth is to blame: although we have somewhat more than
-sixteen times as much processing power available to us, our available memory bandwidth at best
-doubles from introducing parallelism. Unlike the clique algorithm, which has very small,
-cache-friendly data structures which are modified in-place, the state for the $k{\downarrow}$
+sixteen times as much processing power available to us when using threads, our available memory
+bandwidth at best doubles. Unlike the clique algorithm, which has very
+small, cache-friendly data structures which are modified in-place, the state for the $k{\downarrow}$
 algorithm is large and much of the runtime is spent copying data structures.
@@
 Finally, we look at our attempts to parallelise the McSplit algorithm. Recall that doing so required
-heavy modifications to the algorithm, introducing significant amounts of speculative copying of a
+heavy modifications to the implementation, introducing significant amounts of speculative copying of a
 data structure that is usually backtrackable and modified in-place. For unlabelled, unlabelled

### 2017-08-15 paper/paper.tex
More things

@@
     threaded algorithms we create being clearly better than the sequential ones. We then look in
-    more detail at the results, and discuss speedups should be measured for this kind of algorithm.
+    more detail at the results, and discuss how speedups should be measured for this kind of algorithm.
     Because of the difficulty in quantifying an average speedup when so-called anomalous
@@
             \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M6) at (270:1.5) {\phantom{0}};
-            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M7) at ($(330:1.5) + (M6)$) {\phantom{0}};
+            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M8) at ($(210:1.5) + (M6)$) {\phantom{0}};
             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
@@
 hyper-threading is enabled, so we run experiments with 32 threads except where otherwise noted, but
-we do not expect speedups of 32 even in the best case.  We run all our experiments with a one
+we do not expect speedups of 32 even in the ideal case.  We run all our experiments with a one
 hundred second timeout for each instance. All of our sequential runtimes are from optimised

### 2017-11-18 paper/paper.tex
CPAIOR?

@@
 
-\documentclass[sigconf]{acmart}
+\documentclass{llncs}
 
@@
 
-\usetikzlibrary{decorations, decorations.pathreplacing, calc, backgrounds}
+% \usepackage{showframe}
 
-\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
-\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
-\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
-\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
-\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
+\usetikzlibrary{decorations, decorations.pathreplacing, calc, backgrounds, positioning}
 
 % cref style
-\crefname{figure}{Figure}{Figures}
-\Crefname{figure}{Figure}{Figures}
-
-\newcommand{\citepos}[1]{\citet{#1}'s}
+\crefname{figure}{Fig.}{Figs.}
+\Crefname{figure}{Fig.}{Figs.}
 
@@
 
-% http://tex.stackexchange.com/questions/22100/the-bar-and-overline-commands
-\newcommand{\shortoverline}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
-
 \newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
 
-% Copyright
-%\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
+\title{Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms}
 
-% DOI
-\acmDOI{10.475/123_4}
+\author{
+    Ruth Hoffmann\inst{1}
+    \and Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
+        Research Council [grant numbers EP/K503058/1, EP/M508056, and EP/P026842/1]}\inst{2}
+    \and Samba Ndojh Ndiaye\thanks{This work
+    was supported by the ANR project SoLStiCe (ANR-13-BS02-0002-01)}\inst{3}
+    \and Patrick Prosser\samethanks[1]\inst{2}
+    \and Craig Reilly\samethanks[1]\inst{2}
+    \and Christine Solnon\samethanks[2]\inst{4}
+    \and James Trimble\samethanks[1]\inst{2}}
 
-% ISBN
-\acmISBN{123-4567-24-567/08/06}
-
-%Conference
-\acmConference[IA$^3$ 2017]{Seventh Workshop on Irregular Applications: Architectures and Algorithms}{November 2017}{Denver, Colorado USA}
-\acmYear{2017}
-\copyrightyear{2017}
-
-\acmPrice{666.00}
+\institute{
+    University of St Andrews, St Andrews, United Kingdom \and
+    University of Glasgow, Glasgow, Scotland \and
+    Universit\'e Lyon 1, LIRIS, UMR5205, F-69621, France
+    \and INSA-Lyon, LIRIS, UMR5205, F-69621, France}
 
 \begin{document}
-\title[Observations from Parallelising Three Subgraph Algorithms]{Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms}
 
-\author{Ruth Hoffmann}
-\affiliation{%
-    \institution{University of St Andrews}
-    \city{St Andrews, Scotland}
-}
-
-\author{Ciaran McCreesh}
-\affiliation{%
-    \institution{University of Glasgow}
-    \city{Glasgow, Scotland}
-}
-\email{ciaran.mccreesh@glasgow.ac.uk}
-
-\author{Samba Ndojh Ndiaye}
-\affiliation{
-    \institution{Universit\'e Lyon 1, LIRIS, UMR5205}
-    \city{F-69621, France}
-}
-
-\author{Patrick Prosser}
-\affiliation{%
-    \institution{University of Glasgow}
-    \city{Glasgow, Scotland}
-}
-
-\author{Craig Reilly}
-\affiliation{%
-    \institution{University of Glasgow}
-    \city{Glasgow, Scotland}
-}
-
-\author{Christine Solnon}
-\affiliation{
-    \institution{INSA-Lyon, LIRIS, UMR5205}
-    \city{F-69621, France}
-}
-
-\author{James Trimble}
-\affiliation{%
-    \institution{University of Glasgow}
-    \city{Glasgow, Scotland}
-}
-
-% The default list of authors is too long for headers}
-\renewcommand{\shortauthors}{R. Hoffmann, C. McCreesh et al.}
+\maketitle
 
@@
 
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10003752.10003809.10003716.10011136</concept_id>
-<concept_desc>Theory of computation~Discrete optimization</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003752.10003809.10010170</concept_id>
-<concept_desc>Theory of computation~Parallel algorithms</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003752.10003809.10011254.10011256</concept_id>
-<concept_desc>Theory of computation~Branch-and-bound</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003752.10003790.10003795</concept_id>
-<concept_desc>Theory of computation~Constraint and logic programming</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Discrete optimization}
-\ccsdesc[500]{Theory of computation~Parallel algorithms}
-\ccsdesc[500]{Theory of computation~Branch-and-bound}
-\ccsdesc[300]{Theory of computation~Constraint and logic programming}
-
-\keywords{Maximum common subgraph, Maximum clique, Parallel branch and bound, Constraint
-programming, Irregular parallelism, Parallel tree-search}
-
-\maketitle
-
 \section{Introduction}
 
-\begin{figure}[b]
+\begin{figure}[t]
     \centering
@@
 Finding a maximum common subgraph is the key step in measuring the similarity or difference between
-two graphs \citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}: to determine
-the difference between two graphs, we find what they have in common, and then take everything left
-over. Because of this, maximum common subgraph problems frequently arise in biology and chemistry
-\citep{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
+two graphs \cite{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}: to determine
+the difference between two graphs, we instead find what they have in common.
+Because of this, maximum common subgraph problems frequently arise in biology and chemistry
+\cite{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
 represent molecules or reactions, and also in computer vision
-\citep{DBLP:journals/jair/CookH94,DBLP:conf/gbrpr/CombierDS13}, computer-aided manufacturing
-\citep{o:LuoWSN17}, the analysis of programs and malware
-\citep{DBLP:conf/icics/GaoRS08,DBLP:journals/compsec/ParkRS13}, crisis management
-\citep{o:DelavalladeFLL16}, and social network analysis \citep{DBLP:journals/tkde/FangYZZ15}.
+\cite{DBLP:journals/jair/CookH94,DBLP:conf/gbrpr/CombierDS13}, computer-aided manufacturing
+\cite{o:LuoWSN17}, the analysis of programs and malware
+\cite{DBLP:conf/icics/GaoRS08,DBLP:journals/compsec/ParkRS13}, crisis management
+\cite{o:DelavalladeFLL16}, and social network analysis \cite{DBLP:journals/tkde/FangYZZ15}.
 
 A \emph{subgraph isomorphism} is an injective mapping from a \emph{pattern} graph to a \emph{target}
-graph which maps adjacent vertices to adjacent vertices---that is, it \emph{preserves adjacency}.
+graph which \emph{preserves adjacency}---that is, if it maps adjacent vertices to adjacent vertices.
 The isomorphism is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent vertices, preserving
@@
 possible; this paper discusses only induced problems.)
-
 A common variant of the problem requires a largest \emph{connected} subgraph
-\citep{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.  As
+\cite{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.  As
 illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be inferred
@@
 Although both the connected and non-connected variants are NP-hard, recently progress has been made
-towards solving the problem in practice.  This paper looks at three recent branch and bound
-algorithms for maximum common (connected) induced subgraph problems, each of which is the state of
-the art for certain classes of instance. We discuss our experiences in adding parallel tree-search
-to these three algorithms. In each case, our results show that the parallel version of the algorithm
-is clearly better than the sequential version, although a closer look at the results shows many
+towards solving the problem in practice.  This paper looks at three branch and bound algorithms for
+maximum common (connected) induced subgraph problems, each of which is the state of the art for
+certain classes of instance. We discuss our experiences in adding parallel tree-search to these
+three algorithms. In each case, our results show that the parallel version of the algorithm is
+clearly better than the sequential version, although a closer look at the results shows many
 nuances.
@@
 clique in an \emph{association graph}
-\citep{o:Levi73,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/cp/McCreeshNPS16}, which, when combined
-with a modern maximum clique solver \citep{DBLP:journals/ol/SegundoMRH13}, is the current best
-approach for solving the problem on labelled graphs \citep{DBLP:conf/cp/McCreeshNPS16}. A modified
-clique-like algorithm can also be used to solve the maximum common connected subgraph problem, by
-ensuring connectedness during search \citep{DBLP:conf/cp/McCreeshNPS16}; again, this is the best
-known way of solving the problem on labelled graphs. However, the association graph encoding is
-extremely memory-intensive, limiting its practical use to pairs of graphs with no more than a few
-hundred vertices.
+\cite{o:Levi73,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/cp/McCreeshNPS16}; this reduction
+resembles the microstructure encoding \cite{DBLP:conf/aaai/Jegou93a} of the constraint programming
+approach described below. When combined with a modern maximum clique solver
+\cite{DBLP:journals/ol/SegundoMRH13}, this is the current best approach for solving the problem on
+labelled graphs \cite{DBLP:conf/cp/McCreeshNPS16}. A modified clique-like algorithm can also be
+used to solve the maximum common connected subgraph problem, by ensuring connectedness during search
+\cite{DBLP:conf/cp/McCreeshNPS16}; again, this is the best known way of solving the problem on
+labelled graphs. However, the association graph encoding is extremely memory-intensive, limiting its
+practical use to pairs of graphs with no more than a few hundred vertices.
 
@@
 variables, or a special global ``all different except $\bot$'' propagator
-\citep{DBLP:conf/cp/PetitRB01}. The objective is simply to find an assignment of values to
+\cite{DBLP:conf/cp/PetitRB01}. The objective is simply to find an assignment of values to
 variables, maximising the number of variables not set to $\bot$. The state of the art for this
 technique is a dedicated (non-toolkit) implementation of a forward-checking branch and bound search
-over this model \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}; attempts to parallelise
-this algorithm by static decomposition have had limited success \citep{DBLP:conf/ictai/MinotNS15}.
+over this model \cite{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}.
 
 Two approaches exist for ensuring connectedness: either a conventional global constraint and
-propagator can be used \citep{DBLP:conf/cp/McCreeshNPS16}, or a special branching rule can be used
-to enforce connectedness during search \citep{DBLP:conf/mco/VismaraV08}. The two techniques are
-broadly comparable performance-wise \citep{DBLP:conf/cp/McCreeshNPS16}, but the branching rule is
+propagator can be used \cite{DBLP:conf/cp/McCreeshNPS16}, or a special branching rule can be used
+to enforce connectedness during search \cite{DBLP:conf/mco/VismaraV08}. The two techniques are
+broadly comparable performance-wise \cite{DBLP:conf/cp/McCreeshNPS16}, but the branching rule is
 much simpler to implement.
@@
 
-\citet{o:McCreeshPT17} observe that due to the special structure of the maximum common subgraph
-problem, the following property holds throughout the search process using the constraint programming
-model: any two variables either have domains with no values in common (with the possible exception
-of $\bot$), or have identical domains. The McSplit algorithm exploits this property. It explores
-essentially the same search tree as the basic forward-checking constraint programming model, but
-using different supporting algorithms and data structures.  Rather than storing a domain for each
-vertex in the pattern graph, equivalence classes of vertices in both graphs are stored in a special
-data structure which is modified in-place and restored upon backtracking. This enables fast
-propagation of the constraints and smaller memory requirements. In addition, this data structure
-enables stronger branching heuristics to be calculated cheaply. The McSplit algorithm effectively
-dominates conventional constraint programming approaches, being consistently over an order of
-magnitude faster.
+McCreesh et al.\ \cite{o:McCreeshPT17} observe that due to the special structure of the maximum
+common subgraph problem, the following property holds throughout the search process using the
+constraint programming model: any two variables either have domains with no values in common (with
+the possible exception of $\bot$), or have identical domains. The McSplit algorithm exploits this
+property. It explores essentially the same search tree as the basic forward-checking constraint
+programming model, but using different supporting algorithms and data structures.  Rather than
+storing a domain for each vertex in the pattern graph, equivalence classes of vertices in both
+graphs are stored in a special data structure which is modified in-place and restored upon
+backtracking. This enables fast propagation of the constraints and smaller memory requirements. In
+addition, this data structure enables stronger branching heuristics to be calculated cheaply. The
+McSplit algorithm effectively dominates conventional constraint programming approaches, being
+consistently over an order of magnitude faster.
 
@@
 A different take on the constraint programming model is presented by
-\citet{DBLP:conf/aaai/HoffmannMR17}. They approach maximum common subgraph via the
+Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17}. They approach maximum common subgraph via the
 subgraph isomorphism problem, asking the question ``if a pattern graph cannot be found in the
@@
 be matched), and should that also not be satisfiable, it iteratively increases $k$ until the result
-is satisfiable. The advantage of this approach is that strong invariants using paths and the degrees
-of vertices may be used to prune large portions of the search space.
+is satisfiable. This approach exploits strong invariants using paths and the degrees of vertices to
+prune large portions of the search space.
 
@@
 and where it is expected that the solution will involve most of the smaller graph (that is, $k$ is
-expected to be low). The implementation we work with only handles unlabelled instances, and does
-not support the connected variant.
+expected to be low). The sequential implementation we work with does not support labels or the
+connected variant.
 
@@
 Most of the benchmark instances we will use come from a standard database for maximum common
-subgraph problems \citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}. This benchmark
+subgraph problems \cite{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}. This benchmark
 set can be used in a number of ways, for different variants of the problem. Following other recent
-work \citep{DBLP:conf/cp/McCreeshNPS16,DBLP:conf/aaai/HoffmannMR17,o:McCreeshPT17}, we use it to
+work \cite{DBLP:conf/cp/McCreeshNPS16,DBLP:conf/aaai/HoffmannMR17,o:McCreeshPT17}, we use it to
 create five families of instance, as follows:
@@
 \noindent
-Following \citet{DBLP:conf/aaai/HoffmannMR17}, we also work with the 5,725 \textbf{Large} instances
+Following Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17}, we also work with the 5,725 \textbf{Large} instances
 originally introduced for studying portfolios of subgraph isomorphism algorithms
-\citep{DBLP:conf/lion/KotthoffMS16}. These graphs are unlabelled and undirected, and can include up
+\cite{DBLP:conf/lion/KotthoffMS16}. These graphs are unlabelled and undirected, and can include up
 to 6,671 vertices. We do not use the clique encoding on these instances due to its
@@
 The clique and $k{\downarrow}$ algorithms already make use of fine-granularity bit-parallelism. To
-introduce more coarse-grained thread parallelism, we will parallelise the search process: viewing
-backtracking search as forming a tree, we can explore different portions of the tree using different
-threads. We use a shared incumbent, so better solutions found by one thread can be used by others
-immediately. This technique has a long history \citep{o:BaderHC05}. Of particular
-interest to us are so-called \emph{anomalies}
-\citep{DBLP:journals/cacm/LaiS84,DBLP:journals/tc/LiW86,DBLP:conf/irregular/BruinKT95}:
-because we are not performing a fixed amount of work, we should have no expectation of a linear
-speedup, and instead we could see a sublinear speedup (much less than $n$ from $n$ processors, if
-speculative work turns out to be wasted) or a superlinear speedup (much more than $n$ from $n$
-processors, if a strong incumbent is found more quickly). An absolute slowdown (a speedup much less
-than 1) is also possible under some models.
+introduce coarse-grained thread parallelism, we will parallelise search: viewing backtracking search
+as forming a tree, we can explore different portions of the tree using different threads. We use a
+shared incumbent, so better solutions found by one thread can be used by others immediately. In this
+paper we use C++11 native threads, and so only support shared memory systems.
+
+Parallel tree-search has a long history \cite{o:BaderHC05}. Of particular interest to us are
+so-called \emph{anomalies}
+\cite{DBLP:journals/cacm/LaiS84,DBLP:journals/tc/LiW86,DBLP:conf/irregular/BruinKT95}: because we
+are not performing a fixed amount of work, we should have no expectation of a linear speedup, and
+instead we could see a sublinear speedup (much less than $n$ from $n$ processors, if speculative
+work turns out to be wasted) or a superlinear speedup (much more than $n$ from $n$ processors, if a
+strong incumbent is found more quickly). An absolute slowdown (a speedup much less than 1) is also
+possible under some models.
+
+?? Common. Hence not using EPS
+
+?? attempts to parallelise this model by static decomposition have had limited success
+\cite{DBLP:conf/ictai/MinotNS15}
 
@@
 
-Thread-parallel versions of state-of-the-art maximum clique algorithms have already been produced.
-\citet{DBLP:journals/topc/McCreeshP15} compare several of these approaches, and make an important
-observation: although work balance is a problem due to the irregularity of the search tree, often
-the interaction between search order and parallel work decomposition is the dominating factor in
-determining speedups. They explain why anomalies are in fact common in practice: many clique problem
-instances benefit immensely from having found a strong incumbent, but have solutions which are
-either unique or rare, and are hard to find. They propose a work splitting mechanism which offsets
-anomalies, guaranteeing reproducibility (two runs with the same instance on the same hardware will
-give similar runtimes), scalability (increasing the number of cores cannot make things worse), and
-no absolute slowdowns.  Additionally, this mechanism explicitly offsets the commitment to early
-branching choices, where search ordering heuristics are most likely to be inaccurate
-\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/cp/ChuSS09}, making superlinear speedups
-common.
+Thread-parallel versions of state-of-the-art maximum clique algorithms already exist.
+McCreesh et al.\ \cite{DBLP:journals/topc/McCreeshP15} compare several of these approaches, and make
+an important observation: although work balance is a problem due to the irregularity of the search
+tree, often the interaction between search order and parallel work decomposition is the dominating
+factor in determining speedups. They explain why anomalies are in fact common in practice: many
+clique problem instances benefit immensely from having found a strong incumbent, but have solutions
+which are either unique or rare, and are hard to find. They propose a work splitting mechanism which
+offsets anomalies, guaranteeing reproducibility (two runs with the same instance on the same
+hardware will give similar runtimes), scalability (increasing the number of cores cannot make things
+worse), and no absolute slowdowns.  Additionally, this mechanism explicitly offsets the commitment
+to early branching choices, where search ordering heuristics are most likely to be inaccurate
+\cite{DBLP:conf/ijcai/HarveyG95,DBLP:conf/cp/ChuSS09}, making superlinear speedups common.
 
@@
 effectively differs only in the preprocessing stage, and the clique-inspired connected algorithm
-described by \citet{DBLP:conf/cp/McCreeshNPS16} is sufficiently similar that it may be parallelised in
+described by McCreesh et al.\ \cite{DBLP:conf/cp/McCreeshNPS16} is sufficiently similar that it may be parallelised in
 exactly the same way. Based upon preliminary experiments, we use a splitting depth limit of five
@@
 
-\begin{figure*}[p]
+\begin{figure}[p]
     \includegraphics*{gen-graph-cumulative-plain.pdf}
@@
     \includegraphics*{gen-graph-cumulative-33v.pdf}
-    \hfill
-    \includegraphics*{gen-graph-cumulative-33ved.pdf}
 
@@
 
+    \includegraphics*{gen-graph-cumulative-33ved.pdf}
+    \hfill
     \includegraphics*{gen-graph-cumulative-plain-connected.pdf}
-    \hfill
+
+    \vspace*{1em}
+
     \includegraphics*{gen-graph-cumulative-33ve-connected.pdf}
@@
 aggregate than the sequential versions (shown using solid lines).}\label{figure:cumulative}
-\end{figure*}
-
-\begin{figure*}[p]
-    \includegraphics*{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33v-clique-vs-clique-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-plain-connected-clique-vs-clique-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33ve-connected-clique-vs-clique-par-t32.pdf}
-
-    \vspace*{1em}
-
-    \includegraphics*{gen-graph-histogram-plain-clique-vs-clique-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-histogram-33v-clique-vs-clique-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-histogram-33ved-clique-vs-clique-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-histogram-plain-connected-clique-vs-clique-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-histogram-33ve-connected-clique-vs-clique-par-t32.pdf}
-
-    \caption{On the top row, per-instance speedups, using the clique algorithm. The $x$-axis is
-    sequential performance and the $y$-axis is 32 threaded performance, so points below the diagonal
-    line represent a speedup. Darker points represent instances where the solution is relatively
-    large compared to the order of the input graphs. Below, histograms plotting the distribution of
-    speedups for instances whose sequential runtime was at least 500 milliseconds, and below the
-    timeout.}\label{figure:cliquespeedups}
-\end{figure*}
+\end{figure}
 
@@
 of instances solved over time, for both sequential (solid lines) and parallel (dotted lines)
-versions of each algorithm.  Cumulative plots are commonly used in constraint programming and
-boolean satisfiability experiments for comparing the aggregate behaviour of solvers in the presence
-of timeouts.  (Sometimes the axes are reversed, in which case they are called \emph{cactus} plots.)
-To read these plots, make a choice of timeout along the $x$-axis (which uses a log scale). The $y$
-value at that point shows the number of instances whose runtime (individually) is at most $x$, for a
-particular algorithm. In other words, at any given $x$ value, the highest line shows which algorithm
-is able to solve the largest number of instances using a per-instance timeout of that $x$ value,
-bearing in mind that the actual sets of instances solved by each algorithm may be completely
-different.
+versions of each algorithm. To read these plots, make a choice of timeout along the $x$-axis (which
+uses a log scale). The $y$ value at that point shows the number of instances whose runtime
+(individually) is at most $x$, for a particular algorithm. In other words, at any given $x$ value,
+the highest line shows which algorithm is able to solve the largest number of instances using a
+per-instance timeout of that $x$ value, bearing in mind that the actual sets of instances solved by
+each algorithm may be completely different.
 
@@
 \begin{figure}[p]
+    \centering
+    \begin{tikzpicture}%{{{
+        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-plain-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-plain-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Unlabelled}};
+    \end{tikzpicture}
+
+    \centering
+    \begin{tikzpicture}%{{{
+        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-33v-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33v-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-33v-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Vertex labelled}};
+    \end{tikzpicture}
+
+    \centering
+    \begin{tikzpicture}%{{{
+        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33ved-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-33ved-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Both labelled,}};
+        \node[anchor=west, right = -0.2 of C0.east] (C0) {\rotatebox{90}{\scriptsize directed}};
+    \end{tikzpicture}
+
+    \centering
+    \begin{tikzpicture}%{{{
+        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-plain-connected-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-plain-connected-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-plain-connected-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Unlabelled,}};
+        \node[anchor=west, right = -0.2 of C0.east] (C0) {\rotatebox{90}{\scriptsize connected}};
+    \end{tikzpicture}
+
+    \centering
+    \begin{tikzpicture}%{{{
+        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-33ve-connected-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33ve-connected-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-33ve-connected-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Vertex labelled,}};
+        \node[anchor=west, right = -0.2 of C0.east] (C0) {\rotatebox{90}{\scriptsize connected}};
+    \end{tikzpicture}
+
+    \caption{In the left column, per-instance speedups, using the clique algorithm. The $x$-axis is
+    sequential performance and the $y$-axis is 32 threaded performance. In the centre, performance
+    profiles. On the right, histograms plotting the distribution of speedups for instances whose
+    sequential runtime was at least 500 milliseconds, and below the
+    timeout.}\label{figure:cliquespeedups}
+\end{figure}
+
+\begin{figure}[tb]
+    \centering
     \includegraphics{gen-graph-as-clique.pdf}
@@
 \end{figure}
-\begin{figure}[p]
+\begin{figure}[tb]
     \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t2.pdf}
@@
     \includegraphics*{gen-graph-scatter-33ved-clique-par-t2-vs-clique-par-t4.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t4-vs-clique-par-t8.pdf}
 
@@
 
-    \includegraphics*{gen-graph-scatter-33ved-clique-par-t4-vs-clique-par-t8.pdf}
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t8-vs-clique-par-t16.pdf}
     \hfill
-    \includegraphics*{gen-graph-scatter-33ved-clique-par-t8-vs-clique-par-t16.pdf}
-
-    \vspace*{1em}
-
     \includegraphics*{gen-graph-scatter-33ved-clique-par-t16-vs-clique-par-t32.pdf}
@@
 
-In the top row of \cref{figure:cliquespeedups}, we see scatter plots comparing the sequential and
+In the first column of \cref{figure:cliquespeedups}, we see scatter plots comparing the sequential and
 parallel runtimes of the clique algorithm on an instance by instance basis, using a log-log plot.
-Instances which timed out using one algorithm but not the other are shown as points along the outer
-borders. Points below the $x{-}y$ diagonal line represent speedups. The colour of the points indicates
-the relative size of the solution---darker points represent instances where the solution uses most
-of the vertices of the input graphs. (We use these conventions for scatter plots throughout this
-paper.)
+Each point represents one instance, with the $x$-axis being the sequential runtime and the $y$-axis
+the parallel runtime.  Instances which timed out using one algorithm but not the other are shown as
+points along the outer borders. Points below the $x{-}y$ diagonal line represent speedups. The
+colour of the points indicates the relative size of the solution---darker points represent instances
+where the solution uses most of the vertices of the input graphs. (We use these conventions for
+scatter plots throughout this paper.)
 
@@
 experiments to remove anomalies, by priming search with a known-optimal
-solution---\citet{DBLP:journals/topc/McCreeshP15} criticise this disappointingly common practice in
-more detail.
+solution, ?? \cite{DBLP:journals/topc/McCreeshP15}.
 
-A more principled approach is given in the bottom row of \cref{figure:cliquespeedups}. For instances
-where the sequential run both succeeded and took at least 500 milliseconds, we plot the distribution
-of speedups obtained. These histograms confirm our informal observations. However, these plots are
-still not especially satisfactory: in order to calculate a speedup, we can only consider instances
-where the sequential algorithm succeeded. These plots therefore underestimate superlinear speedups.
+A more principled approach is given in the second column of \cref{figure:cliquespeedups}. For
+instances where the sequential run both succeeded and took at least 500 milliseconds, we plot the
+distribution of speedups obtained. These histograms confirm our informal observations. However,
+these plots are still not especially satisfactory: in order to calculate a speedup, we can only
+consider instances where the sequential algorithm succeeded. These plots therefore underestimate
+superlinear speedups. We must also make an arbitrary choice of the lowest sequential runtime to
+consider.
 
-To avoid this weakness, we propose a new way of characterising speedups. Refer back to the
+?? Third column, performance profiles. Three options. Include easy instances, or exclude them. If we
+do, how do we deal with instances solved by one algorithm and not the other? Treat it as timeout, or
+par10. See different results. For example, for unlabelled connected, if we do not use par10, it
+looks like there are around 400 instances where the speedup is between 1 and 10, whereas in reality
+there are very few.
+
+To avoid these weaknesses, we propose a new way of characterising speedups. Refer back to the
 cumulative plots in \cref{figure:cumulative}. The usual way of comparing two algorithms on these
@@
     \includegraphics*{gen-graph-scatter-sip-kdown-vs-kdown-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-kdown-par-t4-vs-kdown-par-t16.pdf}
 
@@
 
+    \includegraphics*{gen-graph-scatter-plain-kdown-par-t32r-vs-kdown-par-t32.pdf}
+    \hfill
     \includegraphics{gen-graph-as-kdown.pdf}
 
-    \vspace*{0.2em}
-
-    \includegraphics*{gen-graph-scatter-plain-kdown-par-t4-vs-kdown-par-t16.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-plain-kdown-par-t32r-vs-kdown-par-t32.pdf}
-
-    \caption{On the top row, per-instance speedups, using the $k{\downarrow}$ algorithm. The
-    $x$-axis is sequential performance and the $y$-axis is 32 threaded performance. Below,
-    aggregate speedups for both families. Finally, scalability and reproducibility.}\label{figure:kdownscatters}
+    \caption{In the first two plots, per-instance speedups, using the $k{\downarrow}$ algorithm. The
+    $x$-axis is sequential performance and the $y$-axis is 32 threaded performance. Next,
+    scalability and reproducibility, and finally, aggregate speedups for both
+    families.}\label{figure:kdownscatters}
 \end{figure}
@@
 ten, with an aggregate speedup of four. These results are not as good as with the clique algorithm.
-A closer look suggests that memory bandwidth is to blame: although we have somewhat more than
+Profiling suggests memory allocation problems: ?? although we have somewhat more than
 sixteen times as much processing power available to us when using threads, our available memory
@@
 larger graphs, the computational effort per recursive call increases by more than the amount the
-memory copying does, reducing the memory bandwidth problem slightly, and additionally a much larger
+memory copying does, reducing the memory problem slightly, and additionally a much larger
 number of superlinear speedups occurred with this family of instances. We could perhaps anticipate
@@
 have a large degree spread), and so the benefits of high-up diversity can be extremely large
-\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/cp/ChuSS09,DBLP:journals/topc/McCreeshP15}. Indeed,
+\cite{DBLP:conf/ijcai/HarveyG95,DBLP:conf/cp/ChuSS09,DBLP:journals/topc/McCreeshP15}. Indeed,
 similar results were seen with a parallel version of the subgraph isomorphism algorithm upon which
-$k{\downarrow}$ is based \citep{DBLP:conf/cp/McCreeshP15}.
+$k{\downarrow}$ is based \cite{DBLP:conf/cp/McCreeshP15}.
 
@@
 work-stealing from Intel Cilk Plus---similar to the clique algorithm
-\citep{DBLP:journals/topc/McCreeshP15}, doing so gives generally reasonable results on average, but
+\cite{DBLP:journals/topc/McCreeshP15}, doing so gives generally reasonable results on average, but
 now repeat runtimes can differ by more than an order of magnitude.)
@@
     \includegraphics*{gen-graph-scatter-33v-mcsplit-vs-mcsplit-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-mcsplit-vs-mcsplit-par-t32.pdf}
 
@@
 
-    \includegraphics*{gen-graph-scatter-33ved-mcsplit-vs-mcsplit-par-t32.pdf}
+    \includegraphics*{gen-graph-scatter-plain-connected-mcsplit-vs-mcsplit-par-t32.pdf}
     \hfill
-    \includegraphics*{gen-graph-scatter-plain-connected-mcsplit-vs-mcsplit-par-t32.pdf}
-
-    \vspace*{0.2em}
-
     \includegraphics*{gen-graph-scatter-33ve-connected-mcsplit-vs-mcsplit-par-t32.pdf}
@@
 
-    \vspace*{0.6em}
+    \vspace*{2em}
 
+    \begin{minipage}[c]{0.62\textwidth}
     \includegraphics{gen-graph-as-mcsplit.pdf}
+    \end{minipage}
+    \hfill
+    \begin{minipage}[t]{0.35\textwidth}
+    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t4-vs-mcsplit-par-t16.pdf}
+    \vspace*{0.2em}
+    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t32r-vs-mcsplit-par-t32.pdf}
+    \end{minipage}
 
-    \vspace*{0.2em}
-
-    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t4-vs-mcsplit-par-t16.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-plain-mcsplit-par-t32r-vs-mcsplit-par-t32.pdf}
-
-    \caption{On the first three rows, per-instance speedups, using McSplit. On the
-    fourth row, aggregate speedups, and on the final row, scalability and
-    reproducibility.}\label{figure:mcsplitscatter}
+    \caption{On the first two rows, per-instance speedups, using McSplit. Below, aggregate speedups
+    on the left, and on the right, scalability and reproducibility.}\label{figure:mcsplitscatter}
 \end{figure}
@@
 
-Aside from the parallel algorithms, we introduced the idea of \emph{aggregate speedups}, to deal
-with measuring a speedup in the presence of anomalies. These figures informed part of our analysis,
-but our results highlight the importance of viewing results in multiple ways, and in using large
-families of instances with different characteristics when evaluating parallel search algorithms.
+Finally, we introduced the idea of \emph{aggregate speedups}, to deal with measuring a speedup in
+the presence of anomalies. These figures informed part of our analysis, but our results highlight
+the importance of viewing results in multiple ways, and in using large families of instances with
+different characteristics when evaluating parallel search algorithms.
 
-\appendix
-\begin{acks}
-    This work was supported by the Engineering and Physical Sciences Research Council [grant
-numbers EP/K503058/1, EP/M508056, and EP/P026842/1] and the ANR project SoLStiCe
-(ANR-13-BS02-0002-01).
-\end{acks}
+?? Decision problems and optimisation problems where solution density is low and bounding is strong.
 
-\bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{splncs}
 \bibliography{combined}

### 2017-11-18 paper/paper.tex
Fatter nodes

@@
 
-We perform our experiments on systems with dual Intel Xeon E5-2640 v2 processors and 64GBytes RAM,
-running Ubuntu 14.04, with GCC 5.3.0 as the compiler. Each machine has a total of sixteen cores, and
-hyper-threading is enabled, so we run experiments with 32 threads except where otherwise noted, but
-we do not expect speedups of 32 even in the ideal case.  We run all our experiments with a one
-hundred second timeout for each instance. All of our sequential runtimes are from optimised
-implementations which were not designed with parallelism in mind---that is, speedups from
-parallelism are genuine improvements over the state of the art.
+We perform our experiments on systems with dual Intel Xeon E5-2697A v4 processors and 512GBytes RAM,
+running Ubuntu 17.04, with GCC 6.3.0 as the compiler. Each machine has a total of thirty two cores.
+We run all our experiments with a one thousand second timeout for each instance. All of our
+sequential runtimes are from optimised implementations which were not designed with parallelism in
+mind---that is, speedups from parallelism are genuine improvements over the state of the art.
 
@@
     scalability and reproducibility, and finally, aggregate speedups for both
-    families.}\label{figure:kdownscatters}
+    families (shown using both types of hardware).}\label{figure:kdownscatters}
 \end{figure}

### 2017-11-21 paper/paper.tex
Yes, CPAIOR

@@
 
-\begin{figure}[t]
-    \centering
-    \begin{tikzpicture}[scale=0.25]%{{{
-        \begin{scope}
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M1) at (90:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M2) at (150:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M3) at (30:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M4) at (210:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M5) at (330:1.5) {\phantom{0}};
-            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M6) at (270:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
-
-            \draw [very thick] (M1) -- (M2);
-            \draw [very thick] (M2) -- (M4);
-            \draw [very thick] (M3) -- (M5);
-            \draw [color=uofgsandstone!60] (M4) -- (M6);
-            \draw [color=uofgsandstone!60] (M5) -- (M6);
-            \draw [very thick] (M3) -- (M1);
-            \draw [color=uofgsandstone!60] (M6) -- (M7);
-            \draw [color=uofgsandstone!60] (M6) -- (M8);
-            \draw [color=uofgsandstone!60] (M6) -- (M9);
-            \draw [very thick] (M7) -- (M9);
-            \draw [very thick] (M8) -- (M9);
-        \end{scope}
-
-        \begin{scope}[xshift=5cm]
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M1) at (90:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M2) at (150:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M3) at (30:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M4) at (210:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M5) at (330:1.5) {\phantom{0}};
-            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M6) at (270:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, fill=magmapurple, inner sep=0.2pt, font=\small] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};
-
-            \draw [very thick] (M1) -- (M2);
-            \draw [very thick] (M2) -- (M4);
-            \draw [very thick] (M3) -- (M5);
-            \draw [color=uofgsandstone!60] (M4) -- (M6);
-            \draw [color=uofgsandstone!60] (M5) -- (M6);
-            \draw [very thick] (M3) -- (M1);
-            \draw [color=uofgsandstone!60] (M6) -- (M7);
-            \draw [very thick] (M7) -- (M8);
-            \draw [very thick] (M8) -- (M9);
-        \end{scope}
-
-        \begin{scope}[xshift=12cm]
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M1) at (90:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M2) at (150:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M3) at (30:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M4) at (210:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M5) at (330:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M6) at (270:1.5) {\phantom{0}};
-            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M7) at ($(330:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M8) at ($(210:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
-
-            \draw [very thick] (M1) -- (M2);
-            \draw [very thick] (M2) -- (M4);
-            \draw [very thick] (M3) -- (M5);
-            \draw [very thick] (M4) -- (M6);
-            \draw [very thick] (M5) -- (M6);
-            \draw [very thick] (M3) -- (M1);
-            \draw [color=uofgsandstone!60] (M6) -- (M7);
-            \draw [very thick] (M6) -- (M8);
-            \draw [color=uofgsandstone!60] (M6) -- (M9);
-            \draw [color=uofgsandstone!60] (M7) -- (M9);
-            \draw [color=uofgsandstone!60] (M8) -- (M9);
-        \end{scope}
-
-        \begin{scope}[xshift=17cm]
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M1) at (90:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M2) at (150:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M3) at (30:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M4) at (210:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M5) at (330:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M6) at (270:1.5) {\phantom{0}};
-            \node[draw, circle, fill=magmaorange, inner sep=0.2pt, font=\small] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
-            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.2pt, font=\small] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};
-
-            \draw [very thick] (M1) -- (M2);
-            \draw [very thick] (M2) -- (M4);
-            \draw [very thick] (M3) -- (M5);
-            \draw [very thick] (M4) -- (M6);
-            \draw [very thick] (M5) -- (M6);
-            \draw [very thick] (M3) -- (M1);
-            \draw [very thick] (M6) -- (M7);
-            \draw [draw=uofgsandstone!60] (M7) -- (M8);
-            \draw [draw=uofgsandstone!60] (M8) -- (M9);
-        \end{scope}
-\end{tikzpicture}
-\caption{A maximum common subgraph of the first two graphs has eight vertices, shaded. A maximum
-common \emph{connected} subgraph, shown on the right, has only seven vertices.}
-\label{figure:mcsexample}
-\end{figure}
-
 Finding a maximum common subgraph is the key step in measuring the similarity or difference between
-two graphs \cite{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}: to determine
-the difference between two graphs, we instead find what they have in common.
-Because of this, maximum common subgraph problems frequently arise in biology and chemistry
+two graphs \cite{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}.  Because of
+this, maximum common subgraph problems frequently arise in biology and chemistry
 \cite{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
@@
 A common variant of the problem requires a largest \emph{connected} subgraph
-\cite{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.  As
-illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be inferred
-from the maximum common subgraph.
+\cite{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
 
@@
 clearly better than the sequential version, although a closer look at the results shows many
-nuances.
+nuances. Thus this paper focusses primarily on presenting and interpreting the experimental data,
+rather than heavy implementation details, in the hopes that the lessons we learned are helpful to
+other practitioners---in particular, we introduce a new measure called \emph{aggregate speedup}
+which is suitable for determining speedups for decision problems or optimisation problems where
+anomalous speedups are common.
 
@@
 
-?? Common. Hence not using EPS
-
-?? attempts to parallelise this model by static decomposition have had limited success
-\cite{DBLP:conf/ictai/MinotNS15}
+We stress that these anomalies are due to changes in the amount of work done, and are not due to
+work balance problems (although work balance is \emph{also} unusually difficult for this problem).
+Anomalies can have a very strong effect on these algorithms, and we will therefore try to mitigate
+them as far as possible. In the evaluation of their ``embarrassingly parallel search'' technique,
+Malapert et al.\ \cite{DBLP:journals/jair/MalapertRR16} ``consider unsatisfiable, enumeration and
+optimization [problem] instances'', and ``ignore the problem of finding a first feasible solution
+because the parallel speedup can be completely uncorrelated to the number of workers, making the
+results hard to analyze''. They do ``consider optimization problems for which the same variability
+can be observed, but at a lesser extent because the optimality proof is required''. Unfortunately,
+many of the instances we consider behave more like decision problem instances than optimisation
+instances: due to the combination of a low solution density, good value-ordering heuristics, and a
+strong bound function in cases where the optimal solution is relatively large, it is often the case
+that the runtime is determined almost entirely by how long it takes to find an optimal solution,
+with the proof of optimality being nearly trivial.  Indeed, attempts to parallelise the basic
+constraint programming model by static decomposition have had limited success
+\cite{DBLP:conf/ictai/MinotNS15}.
 
@@
 We run all our experiments with a one thousand second timeout for each instance. All of our
-sequential runtimes are from optimised implementations which were not designed with parallelism in
-mind---that is, speedups from parallelism are genuine improvements over the state of the art.
+sequential runtimes are from optimised implementations by their original authors which were not
+designed with parallelism in mind---that is, speedups from parallelism are genuine improvements over
+the state of the art.
 
@@
 
-    \caption{The cumulative number of instances solved over time, for different families and
-    algorithms. The 32 threaded parallel versions (shown using dotted lines) are always better in
-aggregate than the sequential versions (shown using solid lines).}\label{figure:cumulative}
+    \caption{The cumulative number of instances solved over time. Except in the bottom left plot,
+    the 32 threaded parallel versions (shown using dotted lines) are always better in aggregate than
+    the sequential versions (shown using solid lines).}\label{figure:cumulative}
 \end{figure}
@@
 
-Each plot gives the same conclusion: if we are working with a timeout of at least $10^2$
-milliseconds, then for any problem family and any sequential algorithm, if given the option of
-switching to the corresponding parallel algorithm, then we should do so.
+With one exception, each plot gives the same conclusion: if we are working with a timeout of at
+least $10^2$ milliseconds, then for any problem family and any sequential algorithm, if given the
+option of switching to the corresponding parallel algorithm, then we should do so. For the McSplit
+algorithm on both labelled, connected instances, the parallel algorithm does not quite catch up to
+the sequential algorithm.
 
-Although they are good at showing general trends, cumulative plots can hide interesting details. We
-therefore now take a closer look at each of the three algorithms in turn.
+Although good at showing general trends, cumulative plots can hide interesting details. We therefore
+now take a closer look at each of the three algorithms in turn.
 
@@
     \caption{In the left column, per-instance speedups, using the clique algorithm. The $x$-axis is
-    sequential performance and the $y$-axis is 32 threaded performance. In the centre, performance
-    profiles. On the right, histograms plotting the distribution of speedups for instances whose
-    sequential runtime was at least 500 milliseconds, and below the
-    timeout.}\label{figure:cliquespeedups}
+    sequential performance and the $y$-axis is 32 threaded performance. In the centre, histograms
+    plotting the distribution of speedups for instances whose sequential runtime was at least 500
+    milliseconds, and below the timeout. On the right, performance
+    profiles.}\label{figure:cliquespeedups}
 \end{figure}
@@
     \centering
-    \includegraphics{gen-graph-as-clique.pdf}
+    \includegraphics*{gen-graph-as-clique.pdf}
 
@@
     sequential runtime, for each family supported by the clique algorithm.}\label{figure:cliqueas}
-\end{figure}
-\begin{figure}[tb]
-    \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t2.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33ved-clique-par-t2-vs-clique-par-t4.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33ved-clique-par-t4-vs-clique-par-t8.pdf}
-
-    \vspace*{1em}
-
-    \includegraphics*{gen-graph-scatter-33ved-clique-par-t8-vs-clique-par-t16.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33ved-clique-par-t16-vs-clique-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-33ved-clique-par-t32r-vs-clique-par-t32.pdf}
-
-    \caption{Per-instance speedups from the clique algorithm on vertex- and edge-labelled, directed
-    instances, when going from sequential to two threads in the first plot, then increasing the
-    number of threads in subsequent plots. The final plot shows 32 threads versus a repeated run
-    also with 32 threads.}\label{figure:cliquescale}
 \end{figure}
@@
 milliseconds, overheads and the preprocessing step dominate, and we are usually only able to achieve
-a small speedup. At higher runtimes, most speedups appear to be between ten and twenty, except on
+a small speedup. At higher runtimes, most speedups appear to be between ten and thirty, except on
 the final family of both labelled connected instances, where they are mostly between five and ten.
@@
 will see that superlinear speedups are even more common). Alternatively, we could rig our
-experiments to remove anomalies, by priming search with a known-optimal
-solution, ?? \cite{DBLP:journals/topc/McCreeshP15}.
+experiments to remove anomalies, by priming search with a known-optimal solution; however, since the
+time to find an optimal solution (but not prove its optimality) is so important, we do not consider
+this to be a fair measure of algorithm performance \cite{DBLP:journals/topc/McCreeshP15}.
 
@@
 these plots are still not especially satisfactory: in order to calculate a speedup, we can only
-consider instances where the sequential algorithm succeeded. These plots therefore underestimate
-superlinear speedups. We must also make an arbitrary choice of the lowest sequential runtime to
-consider.
+consider instances where the sequential algorithm succeeded, and so these plots therefore underestimate
+superlinear speedups. The choice of a 500 millisecond minimum sequential runtime is also rather
+arbitrary, and is acceptable only if we expect the parallel algorithms will only be used on
+relatively hard instances.
 
-?? Third column, performance profiles. Three options. Include easy instances, or exclude them. If we
-do, how do we deal with instances solved by one algorithm and not the other? Treat it as timeout, or
-par10. See different results. For example, for unlabelled connected, if we do not use par10, it
-looks like there are around 400 instances where the speedup is between 1 and 10, whereas in reality
-there are very few.
+In the third column we show performance profiles \cite{DBLP:journals/mp/DolanM02}. A performance
+profile is a cumulative plot of the relative performance of the parallel algorithm to the sequential
+algorithm---instances occurring to the left of the vertical $x = 1$ line are speedups. Each plot
+shows three options as different lines. The `all' line include easy instances whose sequential
+runtime is below 500 milliseconds, whilst the other two lines include them. The `hard' line treats
+sequential timeouts as having been solved at the time limit, whilst the `par10' line treats timeouts
+as taking ten times longer than the timeout (this convention is common in portfolios).
+
+Unfortunately, these three lines can paint very different pictures. For example, for unlabelled
+instances on the top row, if we include easy instances, it appears that parallelism is better on
+just over half of the instances, whilst if we exclude them, parallelism is always beneficial. If we
+do not use the par10 scheme, the performance profile also suggests that there are around 500
+instances where the speedup is below 10, whilst using par10 correctly shows that such instances are
+rare. However, par10 is only effective in this regard because the ``typical'' speedup is in the
+region of 10 (and this is a particular inconvenience because we seek a way of characterising
+speedups which does not rely upon us already knowing that 10 is a reasonable choice of penalty).
+
+A further problem is that to deal with the large superlinear speedups sometimes observed, a log
+scale must be used on the $x$-axis; this makes speedups of 10 and 30 look very similar, whilst in
+practice the difference is important.
 
@@
 We plot aggregate speedups as a function of time in \cref{figure:cliqueas}. For a sequential timeout
-of one hundred seconds, we get speedups of twenty to twenty eight in the unlabelled, vertex
-labelled, and both labelled, directed cases. In the unlabelled connected case, our aggregate speedup
-is thirty seven, which (when taking hyper-threading into account) is strongly superlinear.  With
-some detailed knowledge of the underlying sequential algorithm, this should perhaps not surprise us:
-for instances with a large solution, once we have found that solution, a proof of optimality is
-relatively easy. However, finding that solution can be unusually hard, particularly since the
-branching strategy for the connected constraint necessarily interferes with the tailored search
-order used by modern clique algorithms.  In contrast, for the both labelled connected case, our
-aggregate speedup is only six. A closer inspection of the results shows that the search tree is
-unusually narrow and deep for these instances, making work balance harder and contention higher.
+of one thousand seconds, we get speedups of thirty to forty in the unlabelled, vertex labelled, and
+both labelled, directed cases. In the unlabelled cases, our aggregate speedup are over thirty two,
+which is superlinear.  With some detailed knowledge of the underlying sequential algorithm, this
+should perhaps not surprise us: for instances with a large solution, once we have found that
+solution, a proof of optimality is relatively easy. However, finding that solution can be unusually
+hard, particularly since the branching strategy for the connected constraint necessarily interferes
+with the tailored search order used by modern clique algorithms.  In contrast, for the both labelled
+connected case, our aggregate speedup is barely larger than one. A closer inspection of the results
+shows that the search tree is unusually narrow and deep for these instances, making work balance
+harder and contention higher.
 
@@
 diversity in search being obtained once eight threads are used. As expected, in no case does
-increasing the number of threads make things substantially worse, although going from sixteen to
-thirty two threads does show some small slowdowns---this is due to hyper-threading.
+increasing the number of threads make things substantially worse.
+The final plot in \cref{figure:cliquescale} shows that runtimes are reproducible: running the same
+instance on the same hardware twice takes almost exactly the same amount of time.
 
-The final plot in \cref{figure:cliquescale} shows that runtimes are reproducible: running the same
-instance on the same hardware twice takes almost exactly the same amount of time. (We could not
-reproduce the three outliers in a third run---we suspect other system programs may have woken up
-and caused slight interference during the second run.)
+\begin{figure}[p]
+    \includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t2.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t2-vs-clique-par-t4.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t4-vs-clique-par-t8.pdf}
+
+    \vspace*{0.2em}
+
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t8-vs-clique-par-t16.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t16-vs-clique-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-33ved-clique-par-t32r-vs-clique-par-t32.pdf}
+
+    \caption{Per-instance speedups from the clique algorithm on vertex- and edge-labelled, directed
+    instances, when going from sequential to two threads in the first plot, then increasing the
+    number of threads in subsequent plots. The final plot shows 32 threads versus a repeated run
+    also with 32 threads. ?? 8 threaded results still running}\label{figure:cliquescale}
+\end{figure}
+
+\begin{figure}[p]
+    \includegraphics*{gen-graph-scatter-plain-kdown-vs-kdown-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-sip-kdown-vs-kdown-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-scatter-plain-kdown-par-t4-vs-kdown-par-t16.pdf}
+
+    \vspace*{0.2em}
+
+    \includegraphics*{gen-graph-scatter-plain-kdown-par-t32r-vs-kdown-par-t32.pdf}
+    \hfill
+    \includegraphics*{gen-graph-as-kdown.pdf}
+
+    \caption{In the first two plots, per-instance speedups, using the $k{\downarrow}$ algorithm. The
+    $x$-axis is sequential performance and the $y$-axis is 32 threaded performance. Next,
+    scalability and reproducibility, and finally, aggregate speedups for both
+    families.}\label{figure:kdownscatters}
+\end{figure}
 
@@
 
-\begin{figure}[tb]
-    \includegraphics*{gen-graph-scatter-plain-kdown-vs-kdown-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-sip-kdown-vs-kdown-par-t32.pdf}
-    \hfill
-    \includegraphics*{gen-graph-scatter-plain-kdown-par-t4-vs-kdown-par-t16.pdf}
-
-    \vspace*{1em}
-
-    \includegraphics*{gen-graph-scatter-plain-kdown-par-t32r-vs-kdown-par-t32.pdf}
-    \hfill
-    \includegraphics{gen-graph-as-kdown.pdf}
-
-    \caption{In the first two plots, per-instance speedups, using the $k{\downarrow}$ algorithm. The
-    $x$-axis is sequential performance and the $y$-axis is 32 threaded performance. Next,
-    scalability and reproducibility, and finally, aggregate speedups for both
-    families (shown using both types of hardware).}\label{figure:kdownscatters}
-\end{figure}
-
 In \cref{figure:kdownscatters} we show per-instance and aggregate speedups for the $k{\downarrow}$
 algorithm. On unlabelled instances, we see a range of speedups between slightly less than one and
-ten, with an aggregate speedup of four. These results are not as good as with the clique algorithm.
-Profiling suggests memory allocation problems: ?? although we have somewhat more than
-sixteen times as much processing power available to us when using threads, our available memory
-bandwidth at best doubles. Unlike the clique algorithm, which has very
-small, cache-friendly data structures which are modified in-place, the state for the $k{\downarrow}$
-algorithm is large and much of the runtime is spent copying data structures.
+ten, with an aggregate speedup of seven. These results are not as good as with the clique algorithm.
+Profiling suggests memory allocation problems: although the amount of work done would suggest good
+parallelism, the time taken to perform each domain copy operation increases as the number of threads
+increases. Unlike the clique algorithm, which has very small, cache-friendly data structures which
+are modified in-place, the state for the $k{\downarrow}$ algorithm is large and much of the runtime
+is spent copying data structures. (Our hardware is a dual multi-core processor configuration, and
+each core has its own low-level cache, but memory bandwidth is shared. Interestingly, on older Xeon
+E5 v2 systems, this problem is much more pronounced.)
 
-For the large instances, our aggregate speedup is higher, at over ten. This has two causes: for
+For the large instances, our aggregate speedup is higher, at around twenty. This has two causes: for
 larger graphs, the computational effort per recursive call increases by more than the amount the
@@
 
-The final row of \cref{figure:kdownscatters} shows that as with the clique algorithm, this
-parallelism is reproducible, and that runtimes do not get worse when the number of threads is
+The third and fourth plots in \Cref{figure:kdownscatters} show that as with the clique algorithm,
+this parallelism is reproducible, and that runtimes do not get worse when the number of threads is
 increased. (Although not shown, we also tried to parallelise $k{\downarrow}$ using randomised
@@
     \begin{minipage}[c]{0.62\textwidth}
-    \includegraphics{gen-graph-as-mcsplit.pdf}
+    \includegraphics*{gen-graph-as-mcsplit.pdf}
     \end{minipage}
@@
 heavy modifications to the implementation, introducing significant amounts of speculative copying of a
-data structure that is usually backtrackable and modified in-place. For unlabelled, unlabelled
-connected, and large instances, \cref{figure:mcsplitscatter} shows a particularly high proportion of
-strongly superlinear speedups. This is because the McSplit algorithm is focussed upon
-exploring the search space very quickly, and its branching heuristics do not have the advantage of
-the domain filtering performed by $k{\downarrow}$, or the rich inter-domain knowledge coming from
-the combination of the association graph encoding and the colour ordering used by clique algorithms.
-Thus making a correct value-ordering choice at the top of search is harder for McSplit than for
-other algorithms, and so increased diversity can be particularly beneficial.
+data structure that is usually backtrackable and modified in-place.
+
+For unlabelled, unlabelled connected, and large instances, \cref{figure:mcsplitscatter} shows a
+particularly high proportion of strongly superlinear speedups. This is because the McSplit algorithm
+is focussed upon exploring the search space very quickly, and its branching heuristics do not have
+the advantage of the domain filtering performed by $k{\downarrow}$, or the rich inter-domain
+knowledge coming from the combination of the association graph encoding and the colour ordering used
+by clique algorithms.  Thus making a correct value-ordering choice at the top of search is harder
+for McSplit than for other algorithms, and so increased diversity can be particularly beneficial.
 
@@
 evident in the labelled cases (where following a guessed assignment, many domains are left with only
-two values), and often lead to little to no speedup being obtained.
+two values), and often lead to little to no speedup being obtained. Indeed, for the labelled,
+connected case, we see a slight aggregate \emph{slowdown}.
 
-The scatter plots also show occasional absolute slowdowns, sometimes by up to an order of magnitude.
-These are due to the changes which had to be made to the sequential algorithm (and because we are
-benchmarking against the sequential algorithm, not a parallel algorithm with one thread), rather
-than search order effects. In cases where parallelism cannot be exploited, the cost of speculatively
-copying domains at each level of search can dominate the runtimes. Because of this, fixing work
-balance problems by increasing the splitting depth typically makes matters much worse, not better.
+The scatter plots also show occasional large absolute slowdowns, sometimes by over an order of
+magnitude.  These are due to the changes which had to be made to the sequential algorithm (and
+because we are benchmarking against the sequential algorithm, not a parallel algorithm with one
+thread), rather than search order effects. In cases where parallelism cannot be exploited, the cost
+of speculatively copying domains at each level of search can dominate the runtimes. Because of this,
+fixing work balance problems by increasing the splitting depth typically makes matters much worse,
+not better.
 
-What about scalability and reproducibility? The final row of \cref{figure:mcsplitscatter} presents a
+What about scalability and reproducibility? \Cref{figure:mcsplitscatter} presents a
 less ideal picture than for the previous two algorithms---again, this is due to speculative
@@
 work balance must not come at the expense of the search order properties, nor at the cost of
-increased overheads. We would also like to see distributed versions of these algorithms, to make use
-of clusters of multi-core machines.
+increased overheads.
 
-Finally, we introduced the idea of \emph{aggregate speedups}, to deal with measuring a speedup in
-the presence of anomalies. These figures informed part of our analysis, but our results highlight
-the importance of viewing results in multiple ways, and in using large families of instances with
-different characteristics when evaluating parallel search algorithms.
-
-?? Decision problems and optimisation problems where solution density is low and bounding is strong.
+More generally, we introduced the idea of \emph{aggregate speedups}, to deal with measuring a
+speedup in the presence of anomalies. This measure gives sensible answers even when working with
+instances which behave like decision problems. Aggregate speedups informed part of our analysis, but
+our results highlight the importance of viewing results in multiple ways, and in using large
+families of instances with different characteristics when evaluating parallel search
+algorithms---had we looked only at unlabelled instances, or only at labelled connected instances,
+our conclusion would be very different.
 

### 2017-11-26 paper/paper.tex
Eight threaded freaks
Ruth's comments

@@
 graph which \emph{preserves adjacency}---that is, if it maps adjacent vertices to adjacent vertices.
-The isomorphism is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent vertices, preserving
-non-adjacency. When working with labelled graphs, a subgraph isomorphism must also preserve labels, and
-on directed graphs, it must preserve orientation. A \emph{common induced subgraph} of two graphs $G$
-and $H$ is a pair of induced subgraph isomorphisms from a pattern graph $P$, one to $G$ and one to
-$H$. A \emph{maximum common induced subgraph} is one with as many vertices as possible. (The
-\emph{maximum common partial subgraph} problem is non-induced, with as many edges as
-possible; this paper discusses only induced problems.)
-A common variant of the problem requires a largest \emph{connected} subgraph
+The isomorphism is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent
+vertices, preserving non-adjacency as well. When working with labelled graphs, a subgraph
+isomorphism must preserve labels, and on directed graphs, it must preserve orientation. A
+\emph{common induced subgraph} of two graphs $G$ and $H$ is a pair of induced subgraph isomorphisms
+from a pattern graph $P$, one to $G$ and one to $H$. A \emph{maximum common induced subgraph} is one
+with as many vertices as possible. (The \emph{maximum common partial subgraph} problem is
+non-induced, with as many edges as possible; this paper discusses only induced problems.) A common
+variant of the problem requires a largest \emph{connected} subgraph
 \cite{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
@@
 
-\subsection{Domain Splitting (McSplit)}
+\subsection{Domain Splitting (McSplit and McSplit$\downarrow$)}
 
@@
 consistently over an order of magnitude faster.
+
+The McSplit$\downarrow$ algorithm is a variant designed for instances where we expect nearly all of
+the smaller graph to be found. It branches first on result size, from largest possible result
+downwards.
 
@@
 and where it is expected that the solution will involve most of the smaller graph (that is, $k$ is
-expected to be low). The sequential implementation we work with does not support labels or the
+expected to be low). The sequential implementation we start with does not support labels or the
 connected variant.
@@
 work \cite{DBLP:conf/cp/McCreeshNPS16,DBLP:conf/aaai/HoffmannMR17,o:McCreeshPT17}, we use it to
-create five families of instance, as follows:
+create five families of instances, as follows:
 
@@
 
-With one exception, each plot gives the same conclusion: if we are working with a timeout of at
+With one exception, each plot gives the same conclusion: if we are working with a solving time of at
 least $10^2$ milliseconds, then for any problem family and any sequential algorithm, if given the
@@
 these plots are still not especially satisfactory: in order to calculate a speedup, we can only
-consider instances where the sequential algorithm succeeded, and so these plots therefore underestimate
+consider instances where the sequential algorithm succeeded, and so these plots underestimate
 superlinear speedups. The choice of a 500 millisecond minimum sequential runtime is also rather
@@
 runtime is below 500 milliseconds, whilst the other two lines include them. The `hard' line treats
-sequential timeouts as having been solved at the time limit, whilst the `par10' line treats timeouts
-as taking ten times longer than the timeout (this convention is common in portfolios).
+sequential timeouts as having been solved at the time limit, whilst the `PAR10' line treats timeouts
+as taking ten times longer than the timeout (this convention is common in portfolios
+\cite{DBLP:conf/aaai/XuHL10}).
 
@@
 just over half of the instances, whilst if we exclude them, parallelism is always beneficial. If we
-do not use the par10 scheme, the performance profile also suggests that there are around 500
-instances where the speedup is below 10, whilst using par10 correctly shows that such instances are
-rare. However, par10 is only effective in this regard because the ``typical'' speedup is in the
+do not use the PAR10 scheme, the performance profile also suggests that there are around 500
+instances where the speedup is below 10, whilst using PAR10 correctly shows that such instances are
+rare. However, PAR10 is only effective in this regard because the ``typical'' speedup is in the
 region of 10 (and this is a particular inconvenience because we seek a way of characterising
@@
     number of threads in subsequent plots. The final plot shows 32 threads versus a repeated run
-    also with 32 threads. ?? 8 threaded results still running}\label{figure:cliquescale}
+    also with 32 threads.}\label{figure:cliquescale}
 \end{figure}

### 2017-11-26 paper/paper.tex
Patrick's comments

@@
 strongest on certain classes of instance. The first involves a reduction to the maximum clique
-problem, whilst the other two approaches have a constraint programming feel to them.
+problem, whilst the other two approaches are inspired by constraint programming.
 
@@
 property. It explores essentially the same search tree as the basic forward-checking constraint
-programming model, but using different supporting algorithms and data structures.  Rather than
+programming approach, but using different supporting algorithms and data structures.  Rather than
 storing a domain for each vertex in the pattern graph, equivalence classes of vertices in both
@@
 
-A different take on the constraint programming model is presented by
+A different take on the constraint programming approach is presented by
 Hoffmann et al.\ \cite{DBLP:conf/aaai/HoffmannMR17}. They approach maximum common subgraph via the
@@
 with the proof of optimality being nearly trivial.  Indeed, attempts to parallelise the basic
-constraint programming model by static decomposition have had limited success
+constraint programming approach by static decomposition have had limited success
 \cite{DBLP:conf/ictai/MinotNS15}.

### 2017-11-26 paper/paper.tex
James's comments

@@
 We perform our experiments on systems with dual Intel Xeon E5-2697A v4 processors and 512GBytes RAM,
-running Ubuntu 17.04, with GCC 6.3.0 as the compiler. Each machine has a total of thirty two cores.
+running Ubuntu 17.04, with GCC 6.3.0 as the compiler. Each machine has a total of thirty-two cores.
 We run all our experiments with a one thousand second timeout for each instance. All of our
@@
 With one exception, each plot gives the same conclusion: if we are working with a solving time of at
-least $10^2$ milliseconds, then for any problem family and any sequential algorithm, if given the
+least $100$ milliseconds, then for any problem family and any sequential algorithm, if given the
 option of switching to the corresponding parallel algorithm, then we should do so. For the McSplit
@@
 of one thousand seconds, we get speedups of thirty to forty in the unlabelled, vertex labelled, and
-both labelled, directed cases. In the unlabelled cases, our aggregate speedup are over thirty two,
+both labelled, directed cases. In the unlabelled cases, our aggregate speedup are over thirty-two,
 which is superlinear.  With some detailed knowledge of the underlying sequential algorithm, this
@@
 increased. (Although not shown, we also tried to parallelise $k{\downarrow}$ using randomised
-work-stealing from Intel Cilk Plus---similar to the clique algorithm
-\cite{DBLP:journals/topc/McCreeshP15}, doing so gives generally reasonable results on average, but
-now repeat runtimes can differ by more than an order of magnitude.)
+work-stealing from Intel Cilk Plus. Doing so gives generally reasonable results on average, as it
+does for the clique algorithm \cite{DBLP:journals/topc/McCreeshP15}, but now repeat runtimes can
+differ by more than an order of magnitude.)
 

### 2017-11-27 paper/paper.tex
Christine's comments

@@
 A \emph{subgraph isomorphism} is an injective mapping from a \emph{pattern} graph to a \emph{target}
-graph which \emph{preserves adjacency}---that is, if it maps adjacent vertices to adjacent vertices.
+graph which \emph{preserves adjacency}---that is, it maps adjacent vertices to adjacent vertices.
 The isomorphism is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent
@@
 \begin{description}
-    \item[Unlabelled] instances, by selecting the first ten members of each parameter class where the
-        graphs have up to 50 vertices each---this gives us a total of 4,110 instances.
+    \item[Unlabelled] undirected instances, by selecting the first ten members of each parameter
+        class where the graphs have up to 50 vertices each---this gives us a total of 4,110
+        instances.
 
-    \item[Vertex labelled] instances, by selecting the first ten members of each parameter class
-        (and so graphs have up to 100 vertices each), applying the 33\% labelling scheme for
-        vertices only, and treating edges as undirected. This gives 8,140 instances.
+    \item[Vertex labelled] undirected instances, by selecting the first ten members of each
+        parameter class (and so graphs have up to 100 vertices each), and applying the 33\%
+        labelling scheme for vertices only. This gives 8,140 instances.
 
@@
 
-    \item[Unlabelled, connected] instances, using the same instances as the \emph{unlabelled} case.
+    \item[Unlabelled, connected] instances, as per the \emph{unlabelled} case.
 
@@
         \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-plain-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-plain-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Unlabelled}};
+        \node[anchor=west, right = 0.1 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-plain-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.1 of C2.east] (C3) {\includegraphics*{gen-graph-pp2-plain-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, left = 0.1 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Unlabelled}};
     \end{tikzpicture}
@@
         \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-33v-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33v-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-33v-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Vertex labelled}};
+        \node[anchor=west, right = 0.1 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33v-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.1 of C2.east] (C3) {\includegraphics*{gen-graph-pp2-33v-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, left = 0.1 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Vertex labelled}};
     \end{tikzpicture}
@@
         \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33ved-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-33ved-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Both labelled,}};
+        \node[anchor=west, right = 0.1 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33ved-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.1 of C2.east] (C3) {\includegraphics*{gen-graph-pp2-33ved-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, left = 0.1 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Both labelled,}};
         \node[anchor=west, right = -0.2 of C0.east] (C0) {\rotatebox{90}{\scriptsize directed}};
@@
         \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-plain-connected-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-plain-connected-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-plain-connected-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Unlabelled,}};
+        \node[anchor=west, right = 0.1 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-plain-connected-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.1 of C2.east] (C3) {\includegraphics*{gen-graph-pp2-plain-connected-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, left = 0.1 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Unlabelled,}};
         \node[anchor=west, right = -0.2 of C0.east] (C0) {\rotatebox{90}{\scriptsize connected}};
@@
         \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-33ve-connected-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33ve-connected-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-33ve-connected-clique-vs-clique-par-t32.pdf}};
-        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Vertex labelled,}};
+        \node[anchor=west, right = 0.1 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33ve-connected-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, right = 0.1 of C2.east] (C3) {\includegraphics*{gen-graph-pp2-33ve-connected-clique-vs-clique-par-t32.pdf}};
+        \node[anchor=west, left = 0.1 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Vertex labelled,}};
         \node[anchor=west, right = -0.2 of C0.east] (C0) {\rotatebox{90}{\scriptsize connected}};
@@
 In the third column we show performance profiles \cite{DBLP:journals/mp/DolanM02}. A performance
-profile is a cumulative plot of the relative performance of the parallel algorithm to the sequential
-algorithm---instances occurring to the left of the vertical $x = 1$ line are speedups. Each plot
-shows three options as different lines. The `all' line include easy instances whose sequential
-runtime is below 500 milliseconds, whilst the other two lines include them. The `hard' line treats
-sequential timeouts as having been solved at the time limit, whilst the `PAR10' line treats timeouts
-as taking ten times longer than the timeout (this convention is common in portfolios
-\cite{DBLP:conf/aaai/XuHL10}).
+profile is a cumulative plot of how many times worse the performance of an algorithm is relative to
+the virtual best algorithm. Each plot shows three options as different lines. The `all' lines
+include easy instances whose sequential runtime is below 500 milliseconds, whilst the other two
+lines exclude them. The `hard' line treats sequential timeouts as having been solved at the time
+limit, whilst the `PAR10' line treats timeouts as taking ten times longer than the timeout (this
+convention is common in portfolios \cite{DBLP:conf/aaai/XuHL10}). The solid lines show the
+sequential algorithms, whilst the dotted lines show the parallel algorithms. (There are no dotted
+lines on the top four plots for the `hard' and `PAR10' cases, since the parallel algorithm always
+beats the sequential algorithm in these cases.) We have normalised the $y$-axis to the number of
+counted instances in a given class.
 
 Unfortunately, these three lines can paint very different pictures. For example, for unlabelled
-instances on the top row, if we include easy instances, it appears that parallelism is better on
-just over half of the instances, whilst if we exclude them, parallelism is always beneficial. If we
-do not use the PAR10 scheme, the performance profile also suggests that there are around 500
-instances where the speedup is below 10, whilst using PAR10 correctly shows that such instances are
-rare. However, PAR10 is only effective in this regard because the ``typical'' speedup is in the
+instances on the top row, if we include easy instances, it appears that the parallel algorithm can
+be up to ten times worse, whereas if we exclude them, it is never worse. If we do not use the PAR10
+scheme, the performance profile also suggests that there are around twenty-five percent of the
+hard instances where the speedup is below 10, whilst using PAR10 correctly shows that such instances
+are rare. However, PAR10 is only effective in this regard because the ``typical'' speedup is in the
 region of 10 (and this is a particular inconvenience because we seek a way of characterising
@@
 
-What about scalability and reproducibility? The first plot in \cref{figure:cliquescale} show the
+What about scalability and reproducibility? The first plot in \cref{figure:cliquescale} shows the
 effects of going from sequential to threaded with two cores, and the next four plots show the

### 2017-11-27 paper/paper.tex
32

@@
 
-    \caption{Aggregate speedups from 32 threads on 16 hyper-threaded cores, shown as a function of
-    sequential runtime, for each family supported by the clique algorithm.}\label{figure:cliqueas}
+    \caption{Aggregate speedups from 32 threads, shown as a function of sequential runtime, for each
+    family supported by the clique algorithm.}\label{figure:cliqueas}
 \end{figure}

### 2018-02-28 paper/paper.tex
Little tweaks

@@
     We discuss our experiences adapting three recent algorithms for maximum common (connected)
-    subgraph problems to exploit multi-core parallelism. These algorithms are heavily irregular and
-    very sensitive to search order changes, and so do not easily lend themselves to parallel search.
+    subgraph problems to exploit multi-core parallelism. These algorithms do not easily lend
+    themselves to parallel search, as the search trees are extremely irregular, making balanced work
+    distribution hard, and runtimes are very sensitive to value-ordering heuristic behaviour.
     Nonetheless, our results show that each algorithm can be parallelised successfully, with the
     threaded algorithms we create being clearly better than the sequential ones. We then look in
-    more detail at the results, and discuss how speedups should be measured for this kind of algorithm.
-    Because of the difficulty in quantifying an average speedup when so-called anomalous
+    more detail at the results, and discuss how speedups should be measured for this kind of
+    algorithm.  Because of the difficulty in quantifying an average speedup when so-called anomalous
     speedups (superlinear and sublinear) are common, we propose a new measure called \emph{aggregate
@@
 Two approaches exist for ensuring connectedness: either a conventional global constraint and
-propagator can be used \cite{DBLP:conf/cp/McCreeshNPS16}, or a special branching rule can be used
-to enforce connectedness during search \cite{DBLP:conf/mco/VismaraV08}. The two techniques are
+propagator can be used \cite{DBLP:conf/cp/McCreeshNPS16}, or a special branching rule can
+enforce connectedness during search \cite{DBLP:conf/mco/VismaraV08}. The two techniques are
 broadly comparable performance-wise \cite{DBLP:conf/cp/McCreeshNPS16}, but the branching rule is
-much simpler to implement.
+simpler to implement.
 
@@
     \item[Vertex labelled] undirected instances, by selecting the first ten members of each
-        parameter class (and so graphs have up to 100 vertices each), and applying the 33\%
-        labelling scheme for vertices only. This gives 8,140 instances.
+        parameter class (and so graphs have up to 100 vertices each), using the 33\%
+        labelling scheme \cite{DBLP:journals/prl/SantoFSV03} for vertices only. This gives 8,140 instances.
 
     \item[Both labelled, directed] instances, by selecting the first ten members of each parameter
-        class, and applying the 33\% labelling scheme to both vertices and edges. Again, this gives
-        8,140 instances.
+        class, and applying the 33\% labelling scheme \cite{DBLP:journals/prl/SantoFSV03} to both
+        vertices and edges. Again, this gives 8,140 instances.
 
@@
 strong incumbent is found more quickly). An absolute slowdown (a speedup much less than 1) is also
-possible under some models.
+possible when using some parallelisation techniques.
 
@@
 described by McCreesh et al.\ \cite{DBLP:conf/cp/McCreeshNPS16} is sufficiently similar that it may be parallelised in
-exactly the same way. Based upon preliminary experiments, we use a splitting depth limit of five
-rather than the original three, since maximum common subgraph instances appear to give even more
-irregular search trees than normal clique problem instances.
+exactly the same way. Based upon preliminary experiments, we set the mechanism's splitting depth
+limit parameter to be five rather than the original three, since maximum common subgraph instances
+appear to give even more irregular search trees than normal clique problem instances.
 
@@
 In \cref{figure:kdownscatters} we show per-instance and aggregate speedups for the $k{\downarrow}$
-algorithm. On unlabelled instances, we see a range of speedups between slightly less than one and
-ten, with an aggregate speedup of seven. These results are not as good as with the clique algorithm.
-Profiling suggests memory allocation problems: although the amount of work done would suggest good
+algorithm. On unlabelled instances, we see a range of speedups between 0.9 and ten, with an
+aggregate speedup of seven. These results are not as good as with the clique algorithm.  Profiling
+suggests memory allocation problems: although the amount of work done would suggest good
 parallelism, the time taken to perform each domain copy operation increases as the number of threads

### 2018-06-27 talk/talk.tex
Talk

% vim: set spell spelllang=en tw=100 et sw=4 sts=4 foldmethod=marker foldmarker={{{,}}} :

\documentclass{beamer}

\usepackage{tikz}
\usepackage{xcolor}
\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{amsmath}                   % \operatorname
\usepackage{amsfonts}                  % \mathcal
\usepackage{amssymb}                   % \nexists
\usepackage{gnuplot-lua-tikz}          % graphs
\usepackage[vlined]{algorithm2e} % algorithms
\usepackage{centernot}

\usetikzlibrary{shapes, arrows, shadows, calc, positioning, fit}
\usetikzlibrary{decorations.pathreplacing, decorations.pathmorphing, shapes.misc}
\usetikzlibrary{tikzmark}

\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}
\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.2, 0.129412}
\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\tikzset{vertex/.style={draw, circle, inner sep=0pt, minimum size=0.5cm, font=\small\bfseries}}
\tikzset{notvertex/.style={vertex, color=white, text=black}}
\tikzset{plainvertex/.style={vertex}}
\tikzset{vertexc1/.style={vertex, fill=uofgburgundy, text=white}}
\tikzset{vertexc2/.style={vertex, fill=uofgsandstone, text=white}}
\tikzset{vertexc3/.style={vertex, fill=uofgforest, text=white}}
\tikzset{vertexc4/.style={vertex, fill=uofgheather, text=white}}
\tikzset{edge/.style={color=black!50!white}}
\tikzset{bedge/.style={ultra thick}}

% {{{ theme things
\useoutertheme[footline=authortitle]{miniframes}
\useinnertheme{rectangles}

\setbeamerfont{block title}{size={}}
\setbeamerfont{title}{size=\large,series=\bfseries}
\setbeamerfont{section title}{size=\large,series=\mdseries}
\setbeamerfont{author}{size=\normalsize,series=\mdseries}
\setbeamercolor*{structure}{fg=uofguniversityblue}
\setbeamercolor*{palette primary}{use=structure,fg=black,bg=white}
\setbeamercolor*{palette secondary}{use=structure,fg=white,bg=uofgcobalt}
\setbeamercolor*{palette tertiary}{use=structure,fg=white,bg=uofguniversityblue}
\setbeamercolor*{palette quaternary}{fg=white,bg=black}

\setbeamercolor*{titlelike}{parent=palette primary}

\beamertemplatenavigationsymbolsempty

\setbeamertemplate{title page}
{
    \begin{tikzpicture}[remember picture, overlay]
        \node (shade) at (current page.north west) {
            \begin{tikzpicture}[remember picture, overlay]
                \fill [fill=white, anchor=north west, yshift=-0.5cm, opacity=0.8] (0, 0) rectangle (\paperwidth, -3.2cm);
            \end{tikzpicture}
        };

        \node [anchor=west, xshift=0.4cm, yshift=-2cm] at (shade.west) {
            \begin{minipage}{0.95\paperwidth}\raggedright
                {\usebeamerfont{title}\usebeamercolor[black]{}\inserttitle}\\[0.2cm]
                {\usebeamerfont{author}\usebeamercolor[black]{}\insertauthor}
            \end{minipage}
        };
    \end{tikzpicture}
}

\setbeamertemplate{section page}
{
    \begin{centering}
        \begin{beamercolorbox}[sep=12pt,center]{part title}
            \usebeamerfont{section title}\insertsection\par
        \end{beamercolorbox}
    \end{centering}
}

\newcommand{\frameofframes}{/}
\newcommand{\setframeofframes}[1]{\renewcommand{\frameofframes}{#1}}

\makeatletter
\setbeamertemplate{footline}
{%
    \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
    \end{beamercolorbox}
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
        leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \leavevmode{\usebeamerfont{author in head/foot}\insertshortauthor}%
        \hfill%
        {\usebeamerfont{institute in head/foot}\usebeamercolor[fg]{institute in head/foot}\insertshortinstitute}%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
        leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        {\usebeamerfont{title in head/foot}\insertshorttitle}%
        \hfill%
        {\usebeamerfont{frame number}\usebeamercolor[fg]{frame number}\insertframenumber~\frameofframes~\inserttotalframenumber}
    \end{beamercolorbox}%
    \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
    \end{beamercolorbox}
}

\makeatletter
\newenvironment{nearlyplainframe}[2][]{
    \def\beamer@entrycode{\vspace*{-\headheight}\vspace*{3pt}}
    \setbeamertemplate{headline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line head}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line head}
        \end{beamercolorbox}
        \vspace*{\headheight}
    }

    \setbeamertemplate{footline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
        \end{beamercolorbox}
    }

    \begin{frame}[#1]{#2}
    }{
    \end{frame}
}
\makeatother

\makeatletter
\newenvironment{justborderframe}[2][]{
    \def\beamer@entrycode{\vspace*{-\headheight}}
    \setbeamertemplate{headline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line head}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line head}
        \end{beamercolorbox}
        \vspace*{\headheight}
    }

    \setbeamertemplate{footline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
        \end{beamercolorbox}
    }

    \begin{frame}[#1]{}
    }{
    \end{frame}
}
\makeatother

% }}}

\author{Ruth Hoffmann \and Ciaran McCreesh \and Samba Ndojh Ndiaye \and \\ Patrick Prosser \and Craig Reilly \and Christine Solnon \and James Trimble}
\title{Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms}

\begin{document}

{
    \usebackgroundtemplate{
        \tikz[overlay, remember picture]
        \node[at=(current page.south), anchor=south, inner sep=0pt]{\includegraphics[keepaspectratio=true, height=\paperheight]{background.jpg}};
    }
    \begin{frame}[plain,noframenumbering]
        \titlepage
    \end{frame}
}

\begin{frame}{Maximum Common Induced Subgraph}
    \centering
    \begin{tikzpicture}
        \begin{scope}
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=2pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [color=uofgsandstone!60] (M4) -- (M6);
            \draw [color=uofgsandstone!60] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [color=uofgsandstone!60] (M6) -- (M7);
            \draw [color=uofgsandstone!60] (M6) -- (M8);
            \draw [color=uofgsandstone!60] (M6) -- (M9);
            \draw [very thick] (M7) -- (M9);
            \draw [very thick] (M8) -- (M9);
        \end{scope}

        \begin{scope}[xshift=6cm]
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=2pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=2pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [color=uofgsandstone!60] (M4) -- (M6);
            \draw [color=uofgsandstone!60] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [color=uofgsandstone!60] (M6) -- (M7);
            \draw [very thick] (M7) -- (M8);
            \draw [very thick] (M8) -- (M9);
        \end{scope}
    \end{tikzpicture}
\end{frame}

\begin{frame}{Maximum Common Connected Subgraph}
    \centering
    \begin{tikzpicture}
        \begin{scope}
            \node[draw, circle, fill=uofglawn, inner sep=2pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=2pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=2pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=2pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=2pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=2pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=2pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=2pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=2pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [very thick] (M4) -- (M6);
            \draw [very thick] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [color=uofgsandstone!60] (M6) -- (M7);
            \draw [very thick] (M6) -- (M8);
            \draw [color=uofgsandstone!60] (M6) -- (M9);
            \draw [color=uofgsandstone!60] (M7) -- (M9);
            \draw [color=uofgsandstone!60] (M8) -- (M9);
        \end{scope}

        \begin{scope}[xshift=6cm]
            \node[draw, circle, fill=uofglawn, inner sep=2pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=2pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=2pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=2pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=2pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=2pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=2pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=2pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=2pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [very thick] (M4) -- (M6);
            \draw [very thick] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [very thick] (M6) -- (M7);
            \draw [draw=uofgsandstone!60] (M7) -- (M8);
            \draw [draw=uofgsandstone!60] (M8) -- (M9);
        \end{scope}
    \end{tikzpicture}
\end{frame}

\begin{frame}{Maximum Common Subgraph Problems are Hard}
    \begin{itemize}
        \item State of the art can struggle with:
            \begin{itemize}
                \item 50 vertices on unlabelled graphs (two years ago: 35 vertices).
                \item 100 vertices on labelled graphs (two years ago: 50 vertices).
            \end{itemize}
        \item If the solution is nearly all of the smaller of the two graphs, the problem is easier
            and we can approach ten thousand vertices in the larger graph.
    \end{itemize}
\end{frame}

\begin{frame}{CP Models}
    \begin{itemize}
        \item One variable for each vertex in the smaller input graph.
        \item Domains range over vertices in the other input graph, plus $\bot$.
        \item Objective is to maximise variables not assigned $\bot$.
        \item Adjacent pairs of non-$\bot$ vertices must be mapped to adjacent pairs of vertices.
        \item Non-adjacent pairs of non-$\bot$ vertices must be mapped to non-adjacent pairs of vertices.
        \item All different except $\bot$, with number of allowed $\bot$ occurrences controlled by
            the objective variable.
        \item Smallest domain first, prefer vertices of similar degree.
    \end{itemize}
\end{frame}

\begin{frame}{McSplit and McSplit{$\downarrow$}}
    \begin{itemize}
        \item During search, any two domains are either identical or disjoint, excluding $\bot$.
        \item Backtrackable partition data structures.
        \item Can propagate the bound in linear time.
        \item Can do dual viewpoint variable and value ordering heuristics in linear time and no
            extra space.
        \item Two variants: conventional branch and bound, or descending.
    \end{itemize}
\end{frame}

\begin{frame}{k{$\downarrow$}}
    \begin{itemize}
        \item Start by trying to find an induced subgraph isomorphism.
        \item If that fails, try again allowing one wildcard.
        \item Then two wildcards, and so on.
        \item CP-like algorithm with special data structures.
    \end{itemize}
\end{frame}

\begin{frame}{Clique Models and Microstructure}
    \begin{tikzpicture}[scale=0.33]%{{{
        \begin{scope}
            \node[draw, circle, fill=white, draw=uofgsandstone!60, inner sep=0.5pt, font=\normalsize] (Ma) at (2, 8) {\phantom{0}};
            \node[draw, circle, fill=uofgpumpkin, inner sep=0.5pt, font=\normalsize] (Mb) at (0, 6) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=0.5pt, font=\normalsize] (Mc) at (2, 6) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (Md) at (4, 6) {\phantom{0}};

            \node [right = 0 of Ma, font=\small] { \vphantom{0}a };
            \node [below = 0 of Mb, font=\small] { \vphantom{0}b };
            \node [below = 0 of Mc, font=\small] { \vphantom{0}c };
            \node [below = 0 of Md, font=\small] { \vphantom{0}d };

            \draw [very thick] (Mb) -- (Mc);
            \draw [very thick] (Mc) -- (Md);
            \draw [color=uofgsandstone!60] (Ma) -- (Mc);
            \draw [color=uofgsandstone!60] (Ma) to [out=150, in=80, looseness=8] (Ma);
            \draw [very thick] (Mb) to [out=150, in=80, looseness=8] (Mb);
        \end{scope}

        \begin{scope}[yshift=-2cm]
            \node[draw, circle, fill=uofgpumpkin, inner sep=0.5pt, font=\normalsize] (M1) at (1, 2) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=0.5pt, font=\normalsize] (M2) at (3, 2) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!60, inner sep=0.5pt, font=\normalsize] (M3) at (1, 0) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M4) at (3, 0) {\phantom{0}};

            \node [left = 0 of M1, font=\small] { \vphantom{0}1 };
            \node [right = 0 of M2, font=\small] { \vphantom{0}2 };
            \node [left = 0 of M3, font=\small] { \vphantom{0}3 };
            \node [right = 0 of M4, font=\small] { \vphantom{0}4 };

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [color=uofgsandstone!60] (M3) -- (M4);
            \draw [color=uofgsandstone!60] (M1) -- (M3);

            \draw [color=uofgsandstone!60] (M3) to [out=280, in=210, looseness=8] (M3);
            \draw [very thick] (M1) to [out=150, in=80, looseness=8] (M1);
        \end{scope}

        \begin{scope}[xshift=15cm, yshift=-2cm]
            \node[draw, circle, fill=white, draw=uofgsandstone!60, inner sep=0.5pt, font=\normalsize] (Ma1) at (2.6, 9) {\phantom{0}};
            \node[draw, circle, fill=white, draw=white, inner sep=0.5pt, font=\normalsize] (Ma2) at (4.2, 9) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!60, inner sep=0.5pt, font=\normalsize] (Ma3) at (5.8, 9) {\phantom{0}};
            \node[draw, circle, fill=white, draw=white, inner sep=0.5pt, font=\normalsize] (Ma4) at (7.4, 9) {\phantom{0}};
            \node (La1) [above = 0.2cm of Ma1, font=\small] { \vphantom{0}1 };
            \node (La2) [above = 0.2cm of Ma2, font=\small, color=uofgsandstone!60] { \vphantom{0}2 };
            \node (La3) [above = 0.2cm of Ma3, font=\small] { \vphantom{0}3 };
            \node (La4) [above = 0.2cm of Ma4, font=\small, color=uofgsandstone!60] { \vphantom{0}4 };

            \node [left=0 of La1] { $a \mapsto \{$ };
            \node [right=0 of La4] { $\}$ };

            \node[draw, circle, fill=uofgpumpkin, inner sep=0.5pt, font=\normalsize] (Mb1) at (1, 2.6) {\phantom{0}};
            \node[draw, circle, fill=white, draw=white, inner sep=0.5pt, font=\normalsize] (Mb2) at (1, 4.2) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!60, inner sep=0.5pt, font=\normalsize] (Mb3) at (1, 5.8) {\phantom{0}};
            \node[draw, circle, fill=white, draw=white, inner sep=0.5pt, font=\normalsize] (Mb4) at (1, 7.4) {\phantom{0}};
            \node [left = 0.2cm of Mb1, font=\small] { \vphantom{0}1 };
            \node [left = 0.2cm of Mb2, font=\small, color=uofgsandstone!60] { \vphantom{0}2 };
            \node [left = 0.2cm of Mb3, font=\small] { \vphantom{0}3 };
            \node [left = 0.2cm of Mb4, font=\small, color=uofgsandstone!60] { \vphantom{0}4 };

            \draw [decorate, decoration={brace, raise=0.8cm}, thick] (Mb1.south west) -- (Mb4.north west) node [midway, left=1cm] { $b \mapsto$ };

            \node[draw, circle, fill=white, draw=white, inner sep=0.5pt, font=\normalsize] (Mc1) at (2.6, 1) {\phantom{0}};
            \node[draw, circle, fill=uofglawn, inner sep=0.5pt, font=\normalsize] (Mc2) at (4.2, 1) {\phantom{0}};
            \node[draw, circle, fill=white, draw=white, inner sep=0.5pt, font=\normalsize] (Mc3) at (5.8, 1) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!60, inner sep=0.5pt, font=\normalsize] (Mc4) at (7.4, 1) {\phantom{0}};
            \node (Lc1) [below = 0.2cm of Mc1, font=\small, color=uofgsandstone!60] { \vphantom{0}1 };
            \node (Lc2) [below = 0.2cm of Mc2, font=\small] { \vphantom{0}2 };
            \node (Lc3) [below = 0.2cm of Mc3, font=\small, color=uofgsandstone!60] { \vphantom{0}3 };
            \node (Lc4) [below = 0.2cm of Mc4, font=\small] { \vphantom{0}4 };

            \node [left=0 of Lc1] { $c \mapsto \{$ };
            \node [right=0 of Lc4] { $\}$ };

            \node[draw, circle, fill=white, draw=white, inner sep=0.5pt, font=\normalsize] (Md1) at (9, 2.6) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!60, inner sep=0.5pt, font=\normalsize] (Md2) at (9, 4.2) {\phantom{0}};
            \node[draw, circle, fill=white, draw=white, inner sep=0.5pt, font=\normalsize] (Md3) at (9, 5.8) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (Md4) at (9, 7.4) {\phantom{0}};
            \node [right = 0.2cm of Md1, font=\small, color=uofgsandstone!60] { \vphantom{0}1 };
            \node [right = 0.2cm of Md2, font=\small] { \vphantom{0}2 };
            \node [right = 0.2cm of Md3, font=\small, color=uofgsandstone!60] { \vphantom{0}3 };
            \node [right = 0.2cm of Md4, font=\small] { \vphantom{0}4 };

            \draw [decorate, decoration={brace, raise=0.8cm}, thick] (Md4.north east) -- (Md1.south east) node [midway, right=1cm] { $\reflectbox{\ensuremath{\mapsto}}\, d$ };

            \begin{scope}[on background layer]
                \draw [draw=uofgsandstone!60] ($(Ma1.south)!0.5!(Ma1)$) -- ($(Md4.west)!0.5!(Md4)$);
                \draw [draw=uofgsandstone!60] ($(Ma1.south)!0.5!(Ma1)$) -- ($(Mc2.north)!0.5!(Mc2)$);
                \draw [draw=uofgsandstone!60] ($(Ma3.south)!0.5!(Ma3)$) -- ($(Mc4.north)!0.5!(Mc4)$);
                \draw [draw=uofgsandstone!60] ($(Mb3.east)!0.5!(Mb3)$) -- ($(Mc4.north)!0.5!(Mc4)$);
                \draw [draw=uofgsandstone!60] ($(Md2.west)!0.5!(Md2)$) -- ($(Mc4.north)!0.5!(Mc4)$);
                \draw [draw=uofgsandstone!60] ($(Md2.west)!0.5!(Md2)$) -- ($(Ma3.south)!0.5!(Ma3)$);
                \draw [draw=uofgsandstone!60] ($(Md2.west)!0.5!(Md2)$) -- ($(Mb3.east)!0.5!(Mb3)$);
                \draw [very thick] ($(Mb1.east)!0.5!(Mb1)$) -- ($(Mc2.north)!0.5!(Mc2)$);
                \draw [very thick] ($(Mb1.east)!0.5!(Mb1)$) -- ($(Md4.west)!0.5!(Md4)$);
                \draw [very thick] ($(Mc2.north)!0.5!(Mc2)$) -- ($(Md4.west)!0.5!(Md4)$);
        \end{scope}
        \end{scope}

    \end{tikzpicture}
\end{frame}

\begin{frame}{Connected?}
    \begin{itemize}
        \item We can express connectivity as a constraint, or by a special branching rule.
        \item This also works inside the clique algorithm.
    \end{itemize}
\end{frame}

\begin{frame}{Parallel Tree-Search}
    \begin{itemize}
        \item All three state of the art approaches use backtracking search in some form.
        \item The search trees are particularly nasty, from a parallelism perspective:
            \begin{itemize}
                \item We have a very good value-ordering heuristic.
                \item No pruning at the top of search, and awkward changes to domain sizes during
                    search.
                \item If we have found a large solution, the optimality proof is very short.
            \end{itemize}
        \item Various attempts at static decomposition have failed miserably.
    \end{itemize}
\end{frame}

\begin{frame}{Deterministic Heuristic-Aware Work-Stealing}
    \begin{itemize}
        \item We can adapt a parallel work-stealing strategy used by the clique algorithm:
            \begin{itemize}
                \item Queue the first level of the search tree, preserving left to right ordering
                    (respecting the value-ordering heuristic).
                \item When we run out of work, re-queue remaining work at depth two.
                \item And so on.
            \end{itemize}
        \item Various nice properties, in theory:
            \begin{itemize}
                \item Respects value-ordering heuristic.
                \item Parallel can never be worse than sequential, and adding threads cannot hurt.
                \item Reproducible.
                \item Queue cannot get too big.
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}{Experiments}
    \begin{itemize}
        \item Dual Intel Xeon E5-2697A v4, 32 cores in total, 512GBytes RAM, Ubuntu 17.04.
        \item C++ native threads, manual work-stealing.
        \item Speedups over unmodified sequential algorithms.
        \item Standard benchmark instances:
            \begin{itemize}
                \item 4,110 small unlabelled instances of up to 50 vertices.
                \item 8,140 labelled instances of up to 100 vertices.
                \item 5,725 large unlabelled instances with differing pattern and target sizes.
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{justborderframe}[t]{Cumulative Plots?}
    \includegraphics{gen-graph-cumulative-plain.pdf}\hfill\includegraphics{gen-graph-cumulative-sip.pdf}\\[10pt]
    \includegraphics{gen-graph-cumulative-33ved.pdf}\hfill\includegraphics{gen-graph-cumulative-33ve-connected.pdf}
\end{justborderframe}

\begin{justborderframe}[t]{Scatter Plots?}
    \vspace*{10pt}
    \includegraphics{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}\hfill%
    \includegraphics{gen-graph-scatter-sip-kdown-vs-kdown-par-t32.pdf}\hfill%
    \includegraphics{gen-graph-scatter-sip-mcsplit-vs-mcsplit-par-t32.pdf} \\[10pt]
    \includegraphics{gen-graph-scatter-33ved-clique-vs-clique-par-t32.pdf}\hfill%
    \includegraphics{gen-graph-scatter-33ve-connected-clique-vs-clique-par-t32.pdf}\hfill%
    \includegraphics{gen-graph-scatter-33ve-connected-mcsplit-vs-mcsplit-par-t32.pdf}
\end{justborderframe}

\begin{nearlyplainframe}[t]{What's The Speedup?}
    \includegraphics{gen-graph-histogram-33v-clique-vs-clique-par-t32.pdf}
    \includegraphics{gen-graph-histogram-33ve-connected-clique-vs-clique-par-t32.pdf}
    \includegraphics{gen-graph-histogram-33ved-clique-vs-clique-par-t32.pdf}
    \includegraphics{gen-graph-histogram-plain-clique-vs-clique-par-t32.pdf}
    \includegraphics{gen-graph-histogram-plain-connected-clique-vs-clique-par-t32.pdf}

    \begin{itemize}
        \item Only including items with a sequential runtime of at least 500 milliseconds, that did
            not timeout\ldots
    \end{itemize}
\end{nearlyplainframe}

\begin{nearlyplainframe}[t]{Performance Profiles?}
    \includegraphics{gen-graph-pp2-33v-clique-vs-clique-par-t32.pdf}
    \includegraphics{gen-graph-pp2-33ve-connected-clique-vs-clique-par-t32.pdf}
    \includegraphics{gen-graph-pp2-33ved-clique-vs-clique-par-t32.pdf}
    \includegraphics{gen-graph-pp2-plain-clique-vs-clique-par-t32.pdf}
    \includegraphics{gen-graph-pp2-plain-connected-clique-vs-clique-par-t32.pdf}

    \begin{itemize}
        \item Heavily dependent upon how we treat sequential timeouts, and on whether we include
            trivial instances.
    \end{itemize}
\end{nearlyplainframe}

\begin{nearlyplainframe}[t]{Aggregate Speedups?}
    \includegraphics<1>{gen-graph-as-clique.pdf}
    \includegraphics<2>{gen-graph-as-kdown.pdf}
    \includegraphics<3>{gen-graph-as-mcsplit.pdf}

    \begin{itemize}
        \item Looking at the cumulative plots, how long did it take sequentially to solve the number
            of instances we can solve in this time in parallel?
    \end{itemize}
\end{nearlyplainframe}

\begin{frame}[t]{Reproducibility}
    \includegraphics{gen-graph-scatter-33ved-clique-par-t32r-vs-clique-par-t32.pdf}\hfill
    \includegraphics{gen-graph-scatter-plain-kdown-par-t32r-vs-kdown-par-t32.pdf}\hfill
    \includegraphics{gen-graph-scatter-plain-mcsplit-par-t32r-vs-mcsplit-par-t32.pdf}

    \begin{itemize}
        \item In theory, if we run the same instance twice on the same hardware, the runtimes should
            be identical.
        \item This does not happen in practice if we use conventional randomised work stealing.
    \end{itemize}
\end{frame}

\begin{frame}[t]{Scalability}
    \includegraphics{gen-graph-scatter-33ved-clique-par-t2-vs-clique-par-t4.pdf}\hfill%
    \includegraphics{gen-graph-scatter-33ved-clique-par-t4-vs-clique-par-t8.pdf}\hfill%
    \includegraphics{gen-graph-scatter-33ved-clique-par-t8-vs-clique-par-t16.pdf}

    \begin{itemize}
        \item In theory, increasing the number of threads cannot make things worse.
    \end{itemize}
\end{frame}

\begin{frame}{Conclusion}
    \begin{itemize}
        \item Parallel is better than sequential.
        \item But it is very sensitive to both instance and algorithm choice.
        \item Evaluating parallel algorithms is hard.
    \end{itemize}
\end{frame}

\begin{frame}[plain,noframenumbering]
    \begin{center}
        \vspace*{4em}
        \url{http://www.dcs.gla.ac.uk/~ciaran} \\
        \vspace{1em}
        \href{mailto:ciaran.mccreesh@glasgow.ac.uk}{\nolinkurl{ciaran.mccreesh@glasgow.ac.uk}} \\
    \end{center}

    \begin{tikzpicture}[remember picture, overlay]
        \node at (current page.north west) {
            \begin{tikzpicture}[remember picture, overlay]
                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -3.2cm);
            \end{tikzpicture}
        };
    \end{tikzpicture}
\end{frame}

\end{document}
